---
title: "DA_IBD_LM"
author: "Stefano Romano, adapted by Rebecca Ansorge"
date: "13 November 2023"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: inline
---
# Install load packages
```{r setup, include=FALSE}

list.of.packages <- c("dplyr", "reshape2", "ggplot2", "lattice")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq","microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)


library(ggplot2)
library(phyloseq)
library(dplyr)
library(microbiome)
library(reshape2)
library(vegan)
library(ggExtra)

```

```{r}
PATHanalNov2023<-"~/Documents/Rebecca/onedrive/analysisNov2023/"
```

# Load and clean data
Load the abundance file and the metadata

```{r, message=F}

spec<-read.delim(paste0(PATHanalNov2023,"PrepData/IBD.MGSabundance.txt"), header = T, row.names = 1)
dim(spec)


rownames(spec) %>% tail
```

There are 1547 Species and 4839 samples
The last row is "other". This will be used to calcualte the relative abundances but will then be removed
```{r, message =F, warning =F}

spec.tss<-as.data.frame(prop.table(as.matrix(spec), 2))


# now remove other
spec.tss<-spec.tss[-1*which(rownames(spec.tss) == "other"),]
dim(spec.tss)


# check NAs
spec.na<-apply(spec,2, function(x) is.na(x))
all(spec.na == FALSE) 
xl<-apply(is.na(spec.tss), 2, which)
na.samples<-gsub("[.].*", "", names(unlist(xl))) %>% unique()
# these samples need to be removed

if(length(na.samples)>0){
  spec.tss<-spec.tss[, -1*which(names(spec.tss) %in% na.samples )]
} else {
  cat("no NAs found")
}
  
# are there empty rows
spec.tss[rowSums(spec.tss) == 0,] %>% rownames()

```
There are 1546 Species and 4839 samples


# Load and format the metadata

```{r, message =F, warning =F}
meta<-read.delim(paste0(PATHanalNov2023,"PrepData/IBDmeta.cur.filt.mc.matched.1TimeP.txt"), header = T, sep = "\t") # load a single time point metadata
dim(meta)
sapply(meta, class)

# check how many UIBD and remove to make downstream analysis easier
meta %>%
  group_by(IBD2)%>%
  count()
meta<-meta[which(meta$IBD2!="IBDU"),]

meta$IBD<-factor(meta$IBD, levels = c("healthy", "IBD"))
meta$IBD2<-factor(meta$IBD2, levels = c("healthy", "UC", "CD"))

# filter the species according to samples in the metadata
spec.tss<-spec.tss[, which(names(spec.tss) %in% meta$SmplID)]
dim(spec.tss)

spec.tss[rowSums(spec.tss) == 0,] %>% rownames() # now there are species with 0. Remove them

spec.tss<-spec.tss[rowSums(spec.tss) != 0,] 
dim(spec.tss)

spec.tss[,colSums(spec.tss) == 0] %>% colnames()
spec.tss<-spec.tss[,colSums(spec.tss) != 0] # now there are samples with 0. Remove them
dim(spec.tss)

# I need to remove the same from the metadata
meta<-meta[meta$SmplID %in% names(spec.tss),]
meta<-meta[order(meta$SmplID),]
spec.tss<-spec.tss[, order(names(spec.tss))]
all(names(spec.tss) == meta$SmplID)
dim(spec.tss)
dim(meta)
```

## Chec if all samples hava the right metadata

Check if the cohort is specified for all
```{r, message =F, warning =F}
meta$Cohort_precise %>% unique()
meta$Cohort %>% unique()

```

All have both cohort and cohort precise

## Check all samples have IBD assigned
```{r, message =F, warning =F}
meta$IBD %>% unique()
meta$IBD2 %>% unique()

```

## Check case controls per cohort
```{r, message =F, warning =F}
meta %>%
  dplyr::group_by(Cohort_precise) %>%
  summarise(unique(IBD)) %>%
  print(n = 40)

```

There are two cohorts with no controls:
Fang_UCSD_IBD_Biobank,USA


# Calculate BC

## Make a plot using the single timepoints
```{r}
# just check that metadata and samples are the same
dir.create(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/Figures"), recursive = T)

# make ordination using phyloseq
rownames(meta)<-meta$SmplID
phy<-phyloseq(otu_table(spec.tss, taxa_are_rows = T), sample_data(meta))
bc.spec.tss<-distance(phy, method = "bray")
ord.bs<-ordinate(bc.spec.tss, physeq = phy, method = "PCoA")

sc<-ord.bs$vectors[,1:2]  %>% as.data.frame() # extract the coordinates
all(rownames(sc) == meta$SmplID)

sc.m<-cbind.data.frame(sc, meta)
sc.m$IBD2<-as.factor(sc.m$IBD2)
(o.pcoa<-ggplot(sc.m, aes(Axis.1,   Axis.2, colour = IBD2)) +
    geom_point(size = 4, alpha = 0.5) +
    theme_minimal() +
    scale_colour_manual(values = c("#50B4C8", "#ff6666", "#DACB01", "#A7A7A7")) +
    xlab("PCo1 [14.5%]") +
    ylab("PCo2 [6.2%]") +
  theme(
        plot.caption  = element_text(size = 10),
        plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size =14),
        panel.grid = element_blank(),
        axis.text = element_blank(),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1)) +
    labs(title = "PCoA based on Bray-Curtis dissimilarities"))
(og.cpoa<-ggMarginal(o.pcoa, groupFill = TRUE, size =8))

saveRDS(sc.m,paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.pcoa.data.rds"))

# perform dbRDA with conditioning per study
meta$Cohort_precise<-as.factor(meta$Cohort_precise)
ord.RDA<-ordinate(bc.spec.tss, physeq = phy, 
                  method = "CAP", 
                  formula = ~ Condition(Cohort_precise)) 

sc<-scores(ord.RDA, display = "sites") %>% as.data.frame() # extract the coordinates
all(rownames(sc) == meta$SmplID)

sc.m<-cbind.data.frame(sc, meta)
sc.m$IBD2<-as.factor(sc.m$IBD2)
(o<-ggplot(sc.m, aes(MDS2, MDS1, colour = IBD2)) +
    geom_point(size = 4, alpha = 0.5) +
    theme_minimal() +
    scale_colour_manual(values = c("#50B4C8", "#ff6666", "#DACB01", "#A7A7A7")) +
    xlab("MDS1 [5.4%]") +
    ylab("MDS2 [4.8%]") +
  theme(
        plot.caption  = element_text(size = 10),
        plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size =14),
        panel.grid = element_blank(),
        axis.text = element_blank(),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1))+
        labs(title = "dbRDA based on Bray-Curtis dissimilarities"))

(og<-ggMarginal(o, groupFill = TRUE, size =8))
saveRDS(sc.m,paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.dbrda.data.rds"))
```

# Permanova on first time-point data

```{r}
set.seed(222)
perm<-how(blocks = meta$Cohort_precise, nperm = 999)
ad.bc<-adonis2(formula = bc.spec.tss ~ meta$IBD, data = meta, parallel = T, permutations = perm)
ad.bc
```

2.9% of variance explained (IBD-HC) when conditioned by Cohort_precise.


# Pairwise permanova for CD/UC/HC comparisons

```{r}
library(pairwiseAdonis)
set.seed(222)
ad.bc.ibd2<-pairwise.adonis2(bc.spec.tss ~ IBD2, data = meta, strata = 'Cohort_precise')
ad.bc.ibd2
```

All pairwise comparisons are significant with Cohort_precise as strata.
CD vs UC account for 1.0% of variation
CD vs healthy account for 3.9% of variation
UC vs healthy account for 1.7% of variation

```{r}
set.seed(222)
ad.bc.cohort<-adonis2(formula = bc.spec.tss ~ meta$Cohort_precise, data = meta, parallel = T)
ad.bc.cohort

```

Cohort_precise explains 12.4% of the variation

```{r}
# Save permanova results in RDS
saveRDS(ad.bc,paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.pmva.bc.ibd.rds"))
saveRDS(ad.bc.ibd2,paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.pmvaPair.bc.ibd2.rds"))
saveRDS(ad.bc.cohort,paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.pmva.bc.cohort.rds"))
```

# Calculate distance from centroids
# Reload and clean data
Load the abundance file and the metadata

```{r, message=F}

spec<-read.delim(paste0(PATHanalNov2023,"PrepData/IBD.MGSabundance.txt"), header = T, row.names = 1)
dim(spec)

```

```{r, message =F, warning =F}

spec.tss<-as.data.frame(prop.table(as.matrix(spec), 2))


# now remove other
spec.tss<-spec.tss[-1*which(rownames(spec.tss) == "other"),]
dim(spec.tss)


# check NAs
spec.na<-apply(spec,2, function(x) is.na(x))
all(spec.na == FALSE) 
xl<-apply(is.na(spec.tss), 2, which)
na.samples<-gsub("[.].*", "", names(unlist(xl))) %>% unique()
# this samples need to be removed

if(length(na.samples)>0){
  spec.tss<-spec.tss[, -1*which(names(spec.tss) %in% na.samples )]
} else {
  cat("no NAs")
}

# are there empty rows
spec.tss[rowSums(spec.tss) == 0,] %>% rownames()

spec.tss[,colSums(spec.tss) == 0] %>% colnames()
spec.tss<-spec.tss[,colSums(spec.tss) != 0] # now there are samples with 0. Remove them
dim(spec.tss)
```

# Load and format the metadata

```{r, message =F, warning =F}
meta<-read.delim(paste0(PATHanalNov2023,"PrepData/IBDmeta.cur.filt.mc.matched.txt"), header = T, sep = "\t") # load a single time point metadata
dim(meta)
sapply(meta, class)

# check how many UIBD and remove 
meta %>%
  group_by(IBD2)%>%
  count()
meta<-meta[which(meta$IBD2!="IBDU"),]

meta$IBD<-factor(meta$IBD, levels = c("healthy", "IBD"))
meta$IBD2<-factor(meta$IBD2, levels = c("healthy", "UC", "CD"))

# filter the species according to samples in the meta
spec.tss<-spec.tss[, which(names(spec.tss) %in% meta$SmplID)]
dim(spec.tss)

# I need to remove the same from the metadata
meta<-meta[meta$SmplID %in% names(spec.tss),]
meta<-meta[order(meta$SmplID),]
spec.tss<-spec.tss[, order(names(spec.tss))]
all(names(spec.tss) == meta$SmplID)

```


# Remove samples with only one condition
```{r}
meta$Cohort_precise <-as.character(meta$Cohort_precise)
meta<-subset(meta, Cohort_precise != "Fang_UCSD_IBD_Biobank" | Cohort_precise != "USA")
spect.tss<-spec.tss[, names(spec.tss) %in% meta$SmplID]

```


# Split data in cross-sectional and long
```{r, message =F, warning =F}
meta.cross<-subset(meta, Cohort_setup == "crosssectional")
meta.long<-subset(meta, Cohort_setup == "longitudinal")


spec.cross<-spec.tss[,names(spec.tss) %in% meta.cross$SmplID]
spec.long<-spec.tss[,names(spec.tss) %in% meta.long$SmplID]

```

## Split the data for cohort
```{r, message =F, warning =F}
############
## CROSS ##
##########
spec.cross<-as.data.frame(t(spec.cross))
all(spec.cross %>% rownames() == meta.cross$SmplID)

# add the cohort column to split the data
spec.cross$cohort<-meta.cross$Cohort_precise

# split
l_spec.cross<-split(spec.cross, f = as.factor(spec.cross$cohort))

############
## Long ##
###########

spec.long<-as.data.frame(t(spec.long))
all(spec.long%>% rownames() == meta.long$SmplID)

# add the cohort column to split the data
spec.long$cohort<-meta.long$Cohort_precise

# split
l_spec.long<-split(spec.long, f = as.factor(spec.long$cohort))

```


## cross-sectional
```{r}
l_spec.cross.bc<-lapply(l_spec.cross, function(x) vegdist(x[,-1*which(names(x) == "cohort")], method = "bray"))
# use betadisp
l_spec.cross.betad<-l_spec.cross.bc
l_spec.cross.distances<-l_spec.cross.betad
for(i in 1:length(l_spec.cross.bc)){
  meta.tmps<-meta.cross[meta.cross$SmplID %in% rownames(l_spec.cross[[i]]),]
  l_spec.cross.betad[[i]]<-betadisper(l_spec.cross.bc[[i]], group = meta.tmps$IBD %>% as.factor(), type = "centroid")

  df<-data.frame(samples = names(l_spec.cross.betad[[i]]$distances),
                 distances_ccentr = l_spec.cross.betad[[i]]$distances)
  df<-df[order(df$samples),]
  meta.tmps<-meta.tmps[order(meta.tmps$SmplID),]
  if(all(meta.tmps$SmplID != df$samples)){
    stop("Meta and samples are not in same order")
  }
  df$IBD<-meta.tmps$IBD
  df$IBD2<-meta.tmps$IBD2
  df$cohort<-meta.tmps$Cohort_precise
  l_spec.cross.distances[[i]]<-df
}

# merge
df_all_cross<-do.call(rbind, l_spec.cross.distances)
saveRDS(df_all_cross,paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.df_all_cross.rds"))
```

## longitudinal

Edit: Use the resampling dataset from the DA analysis for the beta-div

```{r}
# import the dataframe with re-sampling samples IBD/healthy
resam.ibd.df<-read.delim2(paste0(PATHanalNov2023,"PrepData/resampling.20.txt"))
unique(resam.ibd.df$Cohort)
resam.ibd2.df<-read.delim2(paste0(PATHanalNov2023,"PrepData/resampling.CDUC.20.txt"))
unique(resam.ibd2.df$Cohort)
```

#### IBD 

```{r}
# remove the ones with only IBD for now
l_spec.long.ibd<-l_spec.long
l_spec.long.ibd<-l_spec.long.ibd[names(l_spec.long.ibd) %in% c("USA","Fang_UCSD_IBD_Biobank") == FALSE] 

# calculate distance from centroid for each resampling
l_spec.long.distances.df.20<-lapply(names(resam.ibd.df[1:20]), function(n){
  l_spec.long.bc<-lapply(l_spec.long.ibd, function(x) {
    y<-x[which(rownames(x)%in%resam.ibd.df[,n]),]
    vegdist(y[,-1*which(names(y) == "cohort")], method = "bray")
    })
  
  # use betadisp
  l_spec.long.betad<-l_spec.long.bc
  l_spec.long.distances<-l_spec.long.betad
  for(i in 1:length(l_spec.long.bc)){
    meta.tmps<-meta.long[meta.long$SmplID %in% rownames(l_spec.long.ibd[[i]]),]%>%
      filter(SmplID%in%resam.ibd.df[,n])
    l_spec.long.betad[[i]]<-betadisper(l_spec.long.bc[[i]], group = meta.tmps$IBD %>% as.factor(), type = "centroid")
  
    df<-data.frame(samples = names(l_spec.long.betad[[i]]$distances),
                   distances_ccentr = l_spec.long.betad[[i]]$distances)
    df<-df[order(df$samples),]
    meta.tmps<-meta.tmps[order(meta.tmps$SmplID),]
    if(all(meta.tmps$SmplID != df$samples)){
      stop("Meta and samples are not in same order")
    }
    df$IBD<-meta.tmps$IBD
    df$IBD2<-meta.tmps$IBD2
    df$cohort<-meta.tmps$Cohort_precise
    df$iteration<-n
    l_spec.long.distances[[i]]<-df
  }
  l_spec.long.distances.df<-do.call(rbind, l_spec.long.distances)
  l_spec.long.distances.df
})

saveRDS(l_spec.long.distances.df.20,paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS//beta.l_spec.long.distances.df.20.rds"))
```

#### IBD2

```{r}
l_spec.long.ibd2<-l_spec.long

# calculate distance from centroid for each resampling
l_spec.long.distances.df.20ibd2<-lapply(names(resam.ibd2.df[1:20]), function(n){
  l_spec.long.bc<-lapply(l_spec.long.ibd2, function(x) {
    y<-x[which(rownames(x)%in%resam.ibd2.df[,n]),]
    vegdist(y[,-1*which(names(y) == "cohort")], method = "bray")
    })
  
  # use betadisp
  l_spec.long.betad<-l_spec.long.bc
  l_spec.long.distances<-l_spec.long.betad
  for(i in 1:length(l_spec.long.bc)){
    meta.tmps<-meta.long[meta.long$SmplID %in% rownames(l_spec.long.ibd2[[i]]),]%>%
      filter(SmplID%in%resam.ibd2.df[,n])
    l_spec.long.betad[[i]]<-betadisper(l_spec.long.bc[[i]], group = meta.tmps$IBD %>% as.factor(), type = "centroid")
  
    df<-data.frame(samples = names(l_spec.long.betad[[i]]$distances),
                   distances_ccentr = l_spec.long.betad[[i]]$distances)
    df<-df[order(df$samples),]
    meta.tmps<-meta.tmps[order(meta.tmps$SmplID),]
    if(all(meta.tmps$SmplID != df$samples)){
      stop("Meta and samples are not in same order")
    }
    df$IBD<-meta.tmps$IBD
    df$IBD2<-meta.tmps$IBD2
    df$cohort<-meta.tmps$Cohort_precise
    df$iteration<-n
    l_spec.long.distances[[i]]<-df
  }
  l_spec.long.distances.df<-do.call(rbind, l_spec.long.distances)
  l_spec.long.distances.df
})

saveRDS(l_spec.long.distances.df.20ibd2,paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.l_spec.long.distances.df.20ibd2.rds"))
```

Distances from control centroids versus inflammation index

```{r}
# all samples included per cohort
# get control centroids

# cross
l_spec.cross.betad2<-list(l_spec.cross.bc)
l_spec.cross.CTRLdistances<-l_spec.cross.betad2

for(i in 1:length(l_spec.cross.bc)){
  print(i)
  meta.tmps<-meta.cross[meta.cross$SmplID %in% rownames(l_spec.cross[[i]]),]
  l_spec.cross.betad2[[i]]<-betadisper(l_spec.cross.bc[[i]], group = meta.tmps$IBD %>% as.factor(), type = "centroid")

  
  centroidsCTRL <- l_spec.cross.betad2[[1]]$centroids["healthy",] # get control centroid coordinates

  # Use principal coordinates of all samples
  sample_coords <- l_spec.cross.betad2[[i]]$vectors
  
  # Function to calculate Euclidean distance to BF centroid
  dist_to_CTRL <- apply(sample_coords, 1, function(x) sqrt(sum((x - centroidsCTRL)^2)))

  df<-data.frame(samples = names(dist_to_CTRL),
                 distances_ccentrCTRL = dist_to_CTRL)
  df<-df[order(df$samples),]
  meta.tmps<-meta.tmps[order(meta.tmps$SmplID),]
  if(all(meta.tmps$SmplID != df$samples)){
    stop("Meta and samples are not in same order")
  }
  df$IBD<-meta.tmps$IBD
  df$IBD2<-meta.tmps$IBD2
  df$cohort<-meta.tmps$Cohort_precise
  l_spec.cross.CTRLdistances[[i]]<-df
}

# merge
CTRLdistances.cross<-do.call(rbind, l_spec.cross.CTRLdistances)

#long
l_spec.long.bc20<-lapply(names(resam.ibd.df[1:20]), function(n){
  l_spec.long.bc<-lapply(l_spec.long.ibd, function(x) {
    y<-x[which(rownames(x)%in%resam.ibd.df[,n]),]
    vegdist(y[,-1*which(names(y) == "cohort")], method = "bray")
    })
  
  # use betadisp
  l_spec.long.betad2<-l_spec.long.bc
  l_spec.long.CTRLdistances<-l_spec.long.betad2
  
  for(i in 1:length(l_spec.long.bc)){
    meta.tmps<-meta.long[meta.long$SmplID %in% rownames(l_spec.long.ibd[[i]]),]%>%
      filter(SmplID%in%resam.ibd.df[,n])
    l_spec.long.betad2[[i]]<-betadisper(l_spec.long.bc[[i]], group = meta.tmps$IBD %>% as.factor(), type = "centroid")
  
  centroidsCTRL <- l_spec.long.betad2[[1]]$centroids["healthy",] # get control centroid coordinates

  # Use principal coordinates of all samples
  sample_coords <- l_spec.long.betad2[[i]]$vectors
  
  # Function to calculate Euclidean distance to BF centroid
  dist_to_CTRL <- apply(sample_coords, 1, function(x) sqrt(sum((x - centroidsCTRL)^2)))
    
    df<-data.frame(samples = names(dist_to_CTRL),
                 distances_ccentrCTRL = dist_to_CTRL)
  df<-df[order(df$samples),]
  meta.tmps<-meta.tmps[order(meta.tmps$SmplID),]
  if(all(meta.tmps$SmplID != df$samples)){
    stop("Meta and samples are not in same order")
  }
  df$IBD<-meta.tmps$IBD
  df$IBD2<-meta.tmps$IBD2
  df$cohort<-meta.tmps$Cohort_precise
  df$iteration<-n
  l_spec.long.CTRLdistances[[i]]<-df
  }
  l_spec.long.CTRLdistances.df<-do.call(rbind, l_spec.long.CTRLdistances)
  l_spec.long.CTRLdistances.df
})

```

```{r}
#load inflammation index
humanReads<-read.delim2(paste0(PATHanalNov2023,"humanReads/IBD2022.4.map"), comment.char = "#", header=FALSE)%>%
  select(1,36)
colnames(humanReads)<-c("SmplID","HumanRds")
head(humanReads)

meta_hr<-inner_join(meta,humanReads,by="SmplID")
meta_hr$HumanRds<-as.numeric(meta_hr$HumanRds)

# calculate human read ranks
meta_hr_ranked <- meta_hr %>%
  group_by(Cohort_precise) %>%
  mutate(rankHumanRds = rank(HumanRds)) %>%
  ungroup()
```

```{r}
# join human read ranks with distance from control centroids
CTRLdistances<-rbind(CTRLdistances.cross%>%mutate(iteration="cross"), l_spec.long.bc20[[1]])
CTRLdistances.crossLongS1_hrRanks<-CTRLdistances%>%
  rename(SmplID=samples)%>%
  left_join(meta_hr_ranked, by="SmplID")%>%
  mutate(IBD=as.character(IBD.x))%>%
  mutate(IBD=as.character(ifelse(IBD=="healthy","control",IBD)))

col_IBD2<-c("gold2","#E31A1C","darkorange","#1F78B4","#1F78B4")
names(col_IBD2)<-c("IBD","CD","UC","healthy","control")

```


```{r}
# use actual inflammation values logged
# plot of first resampling
p_ctrlDitst_HRlog<-ggplot(CTRLdistances.crossLongS1_hrRanks,aes(x=log2(HumanRds), y=log2(distances_ccentrCTRL)))+
  geom_point(aes(color=IBD), alpha=0.7)+  
  geom_smooth(method = "lm", color="black", se = FALSE)+
  scale_color_manual(values=col_IBD2)+
  ylab("log2(Distances to control centroids)")+
  xlab("log2(Inflammation index)")+
  facet_wrap(.~Cohort_precise, scales = "free")+
  theme_bw()
p_ctrlDitst_HRlog
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_ctrlDist_HRlog.png"),p_ctrlDitst_HRlog, width=8, height=6)
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_ctrlDist_HRlog.pdf"),p_ctrlDitst_HRlog, width=8, height=6)
```

```{r}
# test with linear model random effect cohort
ctrlDist.vec<-CTRLdistances.crossLongS1_hrRanks$distances_ccentrCTRL
hr.vec<-CTRLdistances.crossLongS1_hrRanks$HumanRds
cohort.vec<-as.factor(CTRLdistances.crossLongS1_hrRanks$Cohort_precise)

lmm_ctrlDitst_HRlog2<-lmerTest::lmer(log2(ctrlDist.vec) ~ log2(hr.vec) +(1|cohort.vec))
summary_lmm_ctrlDitst_HRlog2<-summary(lmm_ctrlDitst_HRlog2)
summary_lmm_ctrlDitst_HRlog2
```

```{r}
# test with linear model random effect cohort
hr.ranks.vec<-CTRLdistances.crossLongS1_hrRanks$rankHumanRds

lmm_ctrlDitst_HRlog2_ranks<-lmerTest::lmer(ctrlDist.vec ~ hr.ranks.vec +(1|cohort.vec))
summary_lmm_ctrlDitst_HRlog2_ranks<-summary(lmm_ctrlDitst_HRlog2_ranks)
summary_lmm_ctrlDitst_HRlog2_ranks
```

#Save workspace
```{r, message =F, warning =F}
save.image(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RData/Beta_IBD_time.RData"))
```