---
title: "DA_IBD_LM"
author: "Stefano Romano; adapted by R Ansorge"
date: "May 15, 2022"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: console
---
# Install load packages
```{r setup, include=FALSE}

list.of.packages <- c("dplyr", "reshape2", "propr", "meta", "ggplot2", "lattice", "DHARMa", "broomExtra")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq", "microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)

library(meta)
library(DHARMa)
library(nlme)
library(ggplot2)
library(phyloseq)
library(dplyr)
library(microbiome)
library(reshape2)
library(meta)
library(lattice)
#library(broomExtra)

```

```{r}
PATHanalNov2023<-"~/Documents/Rebecca/onedrive/analysisNov2023/"
```
# Load and clean data
Load the abundance file and the metadata

```{r, message=F}

spec<-read.delim(paste0(PATHanalNov2023,"PrepData/IBD.MGSabundance.txt"), header = T, row.names = 1)


dim(spec)

rownames(spec) %>% tail
```

The last row is "other". This will be used to calcualte the relative abundances but will then be removed
```{r, message =F, warning =F}

spec.tss<-as.data.frame(prop.table(as.matrix(spec), 2))
# now remove other

spec.tss<-spec.tss[-1*which(rownames(spec.tss) == "other"),]
```


# Load and format the metadata

```{r, message =F, warning =F}
meta<-read.delim(paste0(PATHanalNov2023,"PrepData/IBDmeta.cur.filt.mc.matched.txt"), header = T, sep = "\t")
dim(meta)
sapply(meta, class)

meta$IBD<-factor(meta$IBD, levels = c("healthy", "IBD"))
meta$IBD2<-factor(meta$IBD2, levels = c("healthy", "UC", "CD"))

# the one below are not made aas factor, they will be made as a factor in the DA function
# meta$age_category<-factor(meta$age_category, levels = c("child", "adult"))
# meta$sex<-factor(meta$sex, levels = c("Female", "Male"))
# load metasumamry
meta.summary<-read.delim(paste0(PATHanalNov2023,"PrepData/IBDmeta.cur.filt.mc.overview.txt"), header = T, sep = "\t")

```

## Chec if all samples hava the right metadata

Check if the cohort is specified for all
```{r, message =F, warning =F}
meta$Cohort_precise %>% unique()
meta$Cohort %>% unique()

```

All have both cohort and cohort precise

## Check all samples have IBD assigned
```{r, message =F, warning =F}
meta$IBD %>% unique()
meta$IBD2 %>% unique()

```
In IBD2 there is IBDU as I converted them into factors I ignored this entry and is treated as an NA now

# Log Tranform data 
We will log the data
```{r, message =F, warning =F}
# Transform in log10
min<-sort(as.vector(as.matrix(spec.tss))[as.vector(as.matrix(spec.tss)) != 0])[1]
spec.tss.log<-log2(spec.tss + min/100)
```

# Log Tranform data 
We will log the data
```{r, message =F, warning =F}
# Transform in log10
min<-sort(as.vector(as.matrix(spec.tss))[as.vector(as.matrix(spec.tss)) != 0])[1]
spec.tss.log<-log2(spec.tss + min/100)
```

# Split data in cross-sectiona and long
```{r, message =F, warning =F}
meta.cross<-subset(meta, Cohort_setup == "crosssectional")
meta.long<-subset(meta, Cohort_setup == "longitudinal")


spec.cross<-spec.tss.log[,names(spec.tss.log) %in% meta.cross$SmplID]
spec.long<-spec.tss.log[,names(spec.tss.log) %in% meta.long$SmplID]

```

## Split the data for cohort
```{r, message =F, warning =F}
############
## CROSS ##
##########
spec.cross<-as.data.frame(t(spec.cross))
all(spec.cross %>% rownames() == meta.cross$SmplID)

# add the cohort column to split the data
spec.cross$cohort<-meta.cross$Cohort_precise

# split
l_spec.cross<-split(spec.cross, f = as.factor(spec.cross$cohort))

############
## Long ##
###########

spec.long<-as.data.frame(t(spec.long))
all(spec.long%>% rownames() == meta.long$SmplID)

# add the cohort column to split the data
spec.long$cohort<-meta.long$Cohort_precise

# split
l_spec.long<-split(spec.long, f = as.factor(spec.long$cohort))

```

# Build lm
## Build linea models for cross-sectional 
```{r, message=FALSE, warning=FALSE, error=FALSE}
# source the function
source(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/DA_function_UCvsCD.R"))

l_spec.cross<-lapply(l_spec.cross, function(x) x[,-1*which(names(x) == "cohort")])

df.cross<-calcualte_da_lm(list.df = l_spec.cross,
                meta = meta.cross,
                treat = "IBD2",
                meta.summary = meta.summary)

saveRDS(df.cross, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/df.cross.uc_cd.rds"))

```


# Build LM for longitudinal with resampling
Load resamplings

```{r, message=FALSE, warning=FALSE, error=FALSE}

df.row<-read.delim2(paste0(PATHanalNov2023,"PrepData/resampling.CDUC.20.txt",sep="\t"))

# now, loop over the 20 repetition and perform the DA 20 times
l.meta.boot<-vector(length = 20, mode = "list")
for(i in 1:20) {
  l.meta.boot[[i]]<- meta.long[which(meta.long$SmplID %in% df.row[,i]),] 
}


# subsample the spec abund matrix 20 times including only the samples in the list of metadata

l_spec.long.boot<-l.meta.boot
for(i in 1:20){
  l_spec.long.boot[[i]]<-lapply(l_spec.long, function(x) x[rownames(x) %in% 
                                                             l.meta.boot[[i]]$SmplID,])

}

# now to run this I have to run the analyses 20 times, I have basically to loop over the list above
l.df.long.boost<-l_spec.long.boot
l_spec.long.boot<-lapply(l_spec.long.boot, function(d) 
  lapply(d, function(x) x[, -1*which(names(x) == "cohort")]))
for(i in 1:20){
  print(paste0("This is repetition number: ", i))
  l.df.long.boost[[i]]<-calcualte_da_lm(list.df = l_spec.long.boot[[i]],
                meta = l.meta.boot[[i]],
                treat = "IBD2", 
                meta.summary = meta.summary)

}

saveRDS(l.df.long.boost, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.df.long.boost.uc_cd.rds"))

```
# Select only the same comparisons

```{r}
uc_cd<-subset(df.cross@df.da, contrast == "UC - CD")
uc_hc<-subset(df.cross@df.da, contrast == "healthy - UC")
cd_hc<-subset(df.cross@df.da, contrast == "healthy - CD")
```


# Perform 20 meta-analyses
############
#  UC vs CD
##########

```{r error=FALSE, message=FALSE, warning=FALSE}

l.meta_uccd<-l.meta.boot

for(i in 1:20){
  tmp<-rbind.data.frame(subset(l.df.long.boost[[i]]@df.da, contrast == "UC - CD"), uc_cd) #adapted to UC -CD from healthy - CD
  bins<-unique(tmp$bin)
  df<-data.frame(TE.random = vector(),
                 seTE.random = vector(),
                 lower.random = vector(),
                 upper.random = vector(),
                 pval.random = vector(),
                 bins = vector())
  for(b in 1:length(bins)){
    tmp.bin<-tmp[tmp$bin == bins[b],]
    met<-try(metagen(TE=estimate, seTE=SE,data=tmp.bin, method.tau = "REML",title =bins[b], studlab = Cohort)) # adapted RAN try command
    if(class(met)[1] != "try-error"){
    met.df<-as.data.frame(c(met[c("TE.random","seTE.random", "lower.random", "upper.random", "pval.random")],bins= bins[b])) 
    df<-rbind.data.frame(df, met.df)  
    } else {df<-rbind.data.frame(df, c(NA, NA, NA, NA, NA, bins= bins[b]))}
  }
  
  df$p.adjust<-p.adjust(df$pval.random, method = "fdr")
  l.meta_uccd[[i]]<-df
}
#saveRDS(l.meta_uccd, "~/Documents/IBD/March_2023/R_analysis/RDS/l.meta_uccd.rds")
saveRDS(l.meta_uccd, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_uccd.rds"))


l.meta_uccd.sign<-lapply(l.meta_uccd, function(x) subset(x,  p.adjust < 0.05))
# create the datafram with the bins
saveRDS(l.meta_uccd.sign, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_uccd.sign.rds"))

df.all<-do.call(rbind, l.meta_uccd.sign)
df.all$in20<-vector(length = nrow(df.all))
for(i in 1: nrow(df.all)){
  df.all$in20[i]<-ifelse(sum(df.all$bins == df.all$bins[i]) == 20, "TRUE", "FALSE")
}
df.all<-df.all[!duplicated(df.all$bins),] # select only duplicated, meaning one per mGS
tax.sign<-df.all[df.all$in20 == TRUE,]$bins

length(tax.sign)
l.meta_uccd.sign.20<-lapply(l.meta_uccd.sign, function(x) subset(x, bins %in% tax.sign))
#dir.create("~/Documents/IBD/March_2023/R_analysis/RDS/", recursive = T)
dir.create(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/", recursive = T))

#saveRDS(l.meta_uccd.sign.20, "~/Documents/IBD/March_2023/R_analysis/RDS/l.meta_uccd.sign.20.rds")
saveRDS(l.meta_uccd.sign.20, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_uccd.sign.20.rds"))

#tax<-read.delim("~/Documents/IBD/March_2023/Raw_data/IBD.MGStaxa_March2023.txt", header = T, row.names = 1)
tax<-read.delim(paste0(PATHanalNov2023,"PrepData/IBD.MGStaxa.txt"), header = T, row.names = 1)

tax[tax.sign,]
#saveRDS(tax[tax.sign,], "~/Documents/IBD/March_2023/R_analysis/RDS/l.meta_uccd.sign.20.taxa.rds")
saveRDS(tax[tax.sign,], paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_uccd.sign.20.taxa.rds"))

```

############
#  UC vs controls
##########

```{r error=FALSE, message=FALSE, warning=FALSE}

l.meta_uccontrols<-l.meta.boot

for(i in 1:20){
  tmp<-rbind.data.frame(subset(l.df.long.boost[[i]]@df.da, contrast == "healthy - UC"), uc_hc)
  bins<-unique(tmp$bin)
  df<-data.frame(TE.random = vector(),
                 seTE.random = vector(),
                 lower.random = vector(),
                 upper.random = vector(),
                 pval.random = vector(),
                 bins = vector())
  for(b in 1:length(bins)){
    tmp.bin<-tmp[tmp$bin == bins[b],]
    met<-try(metagen(TE=estimate, seTE=SE,data=tmp.bin, method.tau = "REML",title =bins[b], studlab = Cohort))# adapted RAN try command
    if(class(met)[1] != "try-error"){
    met.df<-as.data.frame(c(met[c("TE.random","seTE.random", "lower.random", "upper.random", "pval.random")],bins= bins[b])) 
    df<-rbind.data.frame(df, met.df)  
    } else {df<-rbind.data.frame(df, c(NA, NA, NA, NA, NA, bins= bins[b]))}
  }
  
  df$p.adjust<-p.adjust(df$pval.random, method = "fdr")
  l.meta_uccontrols[[i]]<-df
}
saveRDS(l.meta_uccontrols, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_uccontrols.rds"))


l.meta_uccontrols.sign<-lapply(l.meta_uccontrols, function(x) subset(x,  p.adjust < 0.05))
# create the datafram with the bins
saveRDS(l.meta_uccontrols.sign, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_uccontrols.sign.rds"))

df.all<-do.call(rbind, l.meta_uccontrols.sign)
df.all$in20<-vector(length = nrow(df.all))
for(i in 1: nrow(df.all)){
  df.all$in20[i]<-ifelse(sum(df.all$bins == df.all$bins[i]) == 20, "TRUE", "FALSE")
}
df.all<-df.all[!duplicated(df.all$bins),] # select only duplicated
tax.sign<-df.all[df.all$in20 == TRUE,]$bins

length(tax.sign)
l.meta_uccontrols.sign.20<-lapply(l.meta_uccontrols.sign, function(x) subset(x, bins %in% tax.sign))

saveRDS(l.meta_uccontrols.sign.20, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_uccontrols.sign.20.rds"))

tax[tax.sign,]
saveRDS(tax[tax.sign,], paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_uccontrols.sign.20.taxa.rds"))

```

############
#  CD vs controls
##########

```{r error=FALSE, message=FALSE, warning=FALSE}

l.meta_cdcontrols<-l.meta.boot

for(i in 1:20){
  tmp<-rbind.data.frame(subset(l.df.long.boost[[i]]@df.da,  contrast == "healthy - CD"), cd_hc)
  bins<-unique(tmp$bin)
  df<-data.frame(TE.random = vector(),
                 seTE.random = vector(),
                 lower.random = vector(),
                 upper.random = vector(),
                 pval.random = vector(),
                 bins = vector())
  for(b in 1:length(bins)){
    tmp.bin<-tmp[tmp$bin == bins[b],]
    met<-try(metagen(TE=estimate, seTE=SE,data=tmp.bin, method.tau = "REML",title =bins[b], studlab = Cohort))# adapted RAN try command
    if(class(met)[1] != "try-error"){
    met.df<-as.data.frame(c(met[c("TE.random","seTE.random", "lower.random", "upper.random", "pval.random")],bins= bins[b])) 
    df<-rbind.data.frame(df, met.df)  
    } else {df<-rbind.data.frame(df, c(NA, NA, NA, NA, NA, bins= bins[b]))}
  }
  
  df$p.adjust<-p.adjust(df$pval.random, method = "fdr")
  l.meta_cdcontrols[[i]]<-df
}
saveRDS(l.meta_cdcontrols, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_cdcontrols.rds"))


l.meta_cdcontrols.sign<-lapply(l.meta_cdcontrols, function(x) subset(x,  p.adjust < 0.05))
# create the datafram with the bins
saveRDS(l.meta_cdcontrols.sign, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_cdcontrols.sign.rds"))

df.all<-do.call(rbind, l.meta_cdcontrols.sign)
df.all$in20<-vector(length = nrow(df.all))
for(i in 1: nrow(df.all)){
  df.all$in20[i]<-ifelse(sum(df.all$bins == df.all$bins[i]) == 20, "TRUE", "FALSE")
}
df.all<-df.all[!duplicated(df.all$bins),] # select only duplicated
tax.sign<-df.all[df.all$in20 == TRUE,]$bins

length(tax.sign)
l.meta_cdcontrols.sign.20<-lapply(l.meta_cdcontrols.sign, function(x) subset(x, bins %in% tax.sign))
#dir.create("~/Documents/IBD/March_2023/R_analysis/RDS/", recursive = T)

saveRDS(l.meta_cdcontrols.sign.20, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_cdcontrols.sign.20.rds"))

tax[tax.sign,]
saveRDS(tax[tax.sign,], paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_cdcontrols.sign.20.taxa.rds"))

```

#Save workspace
```{r, message =F, warning =F}
dir.create(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RData/", recursive = T))
save.image(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/DA_UC_CD_Nov2023.RData"))
```