---
title: "Correlate human reads"
author: "Rebecca Ansorge"
date: "14/06/2024"
output: 
  html_document:
    df_print: paged
---

Calculate correlations of human read percent with species abundances and with differential abundance results

```{r}
library(tidyverse)
library(lme4)
library(cowplot)
```

Set parameters

```{r}
qval=0.1 # q-value cutoff

# set colors
col_IBD2<-c("gold2","#E31A1C","darkorange","#1F78B4","#1F78B4","#1F78B4", "grey40","#E8A400")
names(col_IBD2)<-c("IBD","CD","UC","healthy","control", "Healthy","variable","disease")

```

### define functions

```{r}
# function to correlate taxa abundances with human read percent
# per cohort
corr_all_chrts<-function(chrt, list_taxa_abun, metaVar, var){
  #chrt<-"Fang_UCSD_IBD_Biobank" # debug
  #list_taxa_abun <- l_spec.long.subsamps[[s]]
  #metaVar <- meta_hr
  #var <- "HumanRds"
    taxab.df<-list_taxa_abun[[chrt]]
    testmgs.vec<-names(taxab.df)
    #testmgs.vec<-names(taxab.df)[1:5] # debug
    taxmeta.df<-taxab.df%>%
      rownames_to_column("SmplID")%>%
      left_join(metaVar, by="SmplID")%>%
      filter(!is.na(!!sym(var)))

    cor_cohort<-lapply(testmgs.vec, function(m) {
     #m<-"Bacteria;Bacteroidota;Bacteroidia;Bacteroidales;Rikenellaceae;Alistipes;Alistipes avistercoris;MGS.424" # debug
      taxab.vec<-taxmeta.df[,m]
      hr.vec<-as.numeric(c(taxmeta.df$HumanRds))

      class(taxab.vec)
      class(hr.vec)

      sd_taxab.vec<-sd(taxab.vec,na.rm = TRUE)
      sd_hr.vec<-sd(hr.vec, na.rm = TRUE)


      if(sd_taxab.vec == 0 | sd_hr.vec == 0){
        return(NULL)
      } else {
        tryCatch(
        {
          cor_taxab_hr<-cor.test(taxab.vec, hr.vec, use = "complete.obs", method="spearman")
          #return(cor_taxab_hr)
          # Extract correlation coefficient and p-value
          corr_rho <- cor_taxab_hr$estimate
          corr_p <- cor_taxab_hr$p.value
        },
        error = function(e) {
          message("Error: ", e$message)
          #return(NA)
          corr_rho <- NA
          corr_p <- NA
        })
        cor_cohort_tmp<-data.frame(
          MGS=m,
          pvalue=corr_p,
          rho=corr_rho
        )
        return(cor_cohort_tmp)
      }
    })
    names(cor_cohort)<-testmgs.vec
    cor_cohort<-cor_cohort[sapply(cor_cohort, function(x) !is.null(x))]
    cor_cohort_df<-do.call(rbind, cor_cohort)%>%
      mutate(cohort=paste(chrt))
    print(paste(chrt, "done"))
    return(cor_cohort_df)
}
```


```{r}
# function to correlate taxa abundances with human read percent
# linear mixed effects model with random effect cohort
corr_blocked<-function(taxa_abun, metaVar, var){
  #taxa_abun <- spec.cross
  #metaVar <- meta_hr
  #var <- "HumanRds"
    taxab.df<-taxa_abun
    testmgs.vec<-names(taxab.df%>%select(-cohort))
    #testmgs.vec<-names(taxab.df)[1:5] # debug
    taxmeta.df<-taxab.df%>%
      rownames_to_column("SmplID")%>%
      left_join(metaVar, by="SmplID")%>%
      filter(!is.na(!!sym(var)))

    cor_mgs<-lapply(testmgs.vec, function(m) {
     #m<-"Bacteria;Firmicutes_C;Negativicutes;Veillonellales;Megasphaeraceae;Megasphaera;Megasphaera sp000417505;MGS.966" # debug
      #m<-"Bacteria;Firmicutes_A;Clostridia;Lachnospirales;Lachnospiraceae;Faecalimonas;Faecalimonas sp900546325;MGS.1523 cohort"
      #print(m)
      taxab.vec<-taxmeta.df[,m]
      hr.vec<-as.numeric(c(taxmeta.df$HumanRds))
      cohort<-taxmeta.df$cohort
      
      class(taxab.vec)
      class(hr.vec)
      cohort<-as.factor(cohort)
      class(cohort)

      sd_taxab.vec<-sd(taxab.vec)
      sd_hr.vec<-sd(hr.vec)


      if(sd_taxab.vec == 0 | sd_hr.vec == 0){
        return(NULL)
      } else {
        tryCatch(
        {
          #cor_taxab_hr<-cor.test(formula= ~ taxab.vec + hr.vec | cohort, use = "complete.obs", method="spearman")
          cor_taxab_hr<-lmerTest::lmer(taxab.vec ~ hr.vec +(1|cohort))
          cor_taxab_hr_summary<-summary(cor_taxab_hr)
          
          #return(cor_taxab_hr)
          # Extract correlation coefficient and p-value
          corr_est <- cor_taxab_hr_summary$coefficients[2,1]
          corr_p <- cor_taxab_hr_summary$coefficients[2,5]
        },
        error = function(e) {
          message("Error: ", e$message)
          #return(NA)
          corr_rho <- NA
          corr_p <- NA
        })
        cor_cohort_tmp<-data.frame(
          MGS=m,
          pvalue=corr_p,
          estimate=corr_est
        )
        return(cor_cohort_tmp)
      }
    })
    names(cor_mgs)<-testmgs.vec
    cor_mgs<-cor_mgs[sapply(cor_mgs, function(x) !is.null(x))]
    cor_mgs_df<-do.call(rbind, cor_mgs)
    return(cor_mgs_df)
}

```

```{r}
# function to correlate taxa abundances with human read percent
# spearman correlation blocked by cohort
corr_spear_blocked<-function(taxa_abun, metaVar, var){
  #taxa_abun <- l_spec.long.subs[[s]] spec.cross
  #metaVar <- meta_ibd_hr
  #var <- "HumanRds"
    taxab.df<-taxa_abun
    testmgs.vec<-names(taxab.df%>%select(-cohort))
    #testmgs.vec<-names(taxab.df)[1:5] # debug
    taxmeta.df<-taxab.df%>%
      rownames_to_column("SmplID")%>%
      left_join(metaVar, by="SmplID")%>%
      filter(!is.na(!!sym(var)))

    cor_mgs<-lapply(testmgs.vec, function(m) {
     #m<-"Bacteria;Firmicutes_C;Negativicutes;Veillonellales;Megasphaeraceae;Megasphaera;Megasphaera sp000417505;MGS.966" # debug
      #m<-"Bacteria;Firmicutes_A;Clostridia;Lachnospirales;Lachnospiraceae;Faecalimonas;Faecalimonas sp900546325;MGS.1523 cohort"
      #print(m)
      taxab.vec<-taxmeta.df[,m]
      hr.vec<-as.numeric(c(taxmeta.df$HumanRds))
      cohort<-taxmeta.df$cohort
      
      class(taxab.vec)
      class(hr.vec)
      cohort<-as.factor(cohort)
      class(cohort)

      sd_taxab.vec<-sd(taxab.vec, na.rm = TRUE)
      sd_hr.vec<-sd(hr.vec, na.rm = TRUE)


      if(sd_taxab.vec == 0 | sd_hr.vec == 0){
        return(NULL)
      } else {
        tryCatch(
        {
          #cor_taxab_hr<-cor.test(formula= ~ taxab.vec + hr.vec | cohort, use = "complete.obs", method="spearman")
          #linear model
          cor_taxab_hr<-lmerTest::lmer(taxab.vec ~ hr.vec +(1|cohort))
          cor_taxab_hr_summary<-summary(cor_taxab_hr)
          
          # spearman correlation
          sp_taxab_hr<-coin::spearman_test(taxab.vec ~ hr.vec | cohort)
          
          #return(cor_taxab_hr)
          # Extract correlation coefficient and p-value
          corr_est <- cor_taxab_hr_summary$coefficients[2,1]
          corr_p <- cor_taxab_hr_summary$coefficients[2,5]
          
          # spearman
          sp_rho <- coin::statistic(sp_taxab_hr)
          sp_p <- coin::pvalue(sp_taxab_hr)
        },
        error = function(e) {
          message("Error: ", e$message)
          #return(NA)
          corr_rho <- NA
          corr_p <- NA
        })
        cor_cohort_tmp<-data.frame(
          MGS=m,
          pvalue=corr_p,
          estimate=corr_est,
          spear_pvalue=sp_p,
          spear_estimate=sp_rho
        )
        return(cor_cohort_tmp)
      }
    })
    names(cor_mgs)<-testmgs.vec
    cor_mgs<-cor_mgs[sapply(cor_mgs, function(x) !is.null(x))]
    cor_mgs_df<-do.call(rbind, cor_mgs)
    return(cor_mgs_df)
}

```

### set main path

```{r}
PATHanalNov2023<-"~/Documents/Rebecca/onedrive/analysisNov2023/"
PATH="~/Documents/Rebecca/onedrive/submission/natcom/revision/checks/"
```

### import metadata and human reads

```{r}
meta<-read.delim2(paste0(PATHanalNov2023,"PrepData/IBDmeta.cur.filt.mc.matched.txt"),sep="\t")
humanReads<-read.delim2(paste0(PATHanalNov2023,"humanReads/IBD2022.4.map"), comment.char = "#", header=FALSE)%>%
  select(1,36)
colnames(humanReads)<-c("SmplID","HumanRds")
head(humanReads)

meta_hr<-inner_join(meta,humanReads,by="SmplID")
meta_hr$HumanRds<-as.numeric(meta_hr$HumanRds)
```



### import significantly DA MGS

```{r}
ibd_ctrl.sig<-read.delim2(paste0(PATHanalNov2023,"Downstream/Results/DAmeans.IBDHC.0.1.txt"))%>%mutate(enrichedIn=ifelse(robustDA=="non-sig",NA,ifelse(meanTE<0,"disease","control")))
(ibd_ctrl.sig%>%filter(grepl("gnavus",MGS)))$meanTE<0

cd_ctrl.sig<-read.delim2(paste0(PATHanalNov2023,"Downstream/Results/DAmeans.CDHC.0.1.txt"))%>%mutate(enrichedIn=ifelse(robustDA=="non-sig",NA,ifelse(meanTE<0,"disease","control")))
(cd_ctrl.sig%>%filter(grepl("gnavus",MGS)))$meanTE<0

uc_ctrl.sig<-read.delim2(paste0(PATHanalNov2023,"Downstream/Results/DAmeans.UCHC.0.1.txt"))%>%mutate(enrichedIn=ifelse(robustDA=="non-sig",NA,ifelse(meanTE<0,"disease","control")))
(uc_ctrl.sig%>%filter(grepl("gnavus",MGS)))$meanTE<0

cd_uc.sig<-read.delim2(paste0(PATHanalNov2023,"Downstream/Results/DAmeans.CDUC.0.1.txt"))%>%mutate(enrichedIn=ifelse(robustDA=="non-sig",NA,ifelse(meanTE<0,"CD","UC")))
(cd_uc.sig%>%filter(grepl("gnavus",MGS)))$meanTE<0

da_all.sig<-read.delim2(paste0(PATHanalNov2023,"Downstream/Results/DAmeans.ALL.0.1.txt"))



# make effect size numeric
sapply(ibd_ctrl.sig, class)
ibd_ctrl.sig[["meanTE"]] <- sapply(ibd_ctrl.sig[["meanTE"]],as.numeric)
sapply(cd_ctrl.sig, class)
cd_ctrl.sig[["meanTE"]] <- sapply(cd_ctrl.sig[["meanTE"]],as.numeric)
sapply(uc_ctrl.sig, class)
uc_ctrl.sig[["meanTE"]] <- sapply(uc_ctrl.sig[["meanTE"]],as.numeric)
sapply(cd_uc.sig, class)
cd_uc.sig[["meanTE"]] <- sapply(cd_uc.sig[["meanTE"]],as.numeric)
```

### load abundances

```{r, message=F}
spec<-read.delim(paste0(PATHanalNov2023,"PrepData/IBD.MGSabundance.txt"), header = T, row.names = 1)

spec.tss<-as.data.frame(prop.table(as.matrix(spec), 2))
# now remove other

spec.tss<-spec.tss[-1*which(rownames(spec.tss) == "other"),]
dim(spec.tss)

prevMGS<-read.table(paste0(PATHanalNov2023,"Downstream/Results/prev.filt.MGS.txt"),check.names = FALSE, header=TRUE)
spec.tss<-spec.tss[prevMGS$x,]

meta$IBD<-factor(meta$IBD, levels = c("healthy", "IBD"))
meta$IBD2<-factor(meta$IBD2, levels = c("healthy", "UC", "CD"))

spec.tss.x<-spec.tss
```

### Split data in cross-sectional and long

```{r, message =F, warning =F}
spec.tss.x<-spec.tss.x[, names(spec.tss.x) %in% meta$SmplID]

meta.cross<-subset(meta, Cohort_setup == "crosssectional")
meta.long<-subset(meta, Cohort_setup == "longitudinal")


spec.cross<-spec.tss.x[,names(spec.tss.x) %in% meta.cross$SmplID]
spec.long<-spec.tss.x[,names(spec.tss.x) %in% meta.long$SmplID]

```

## Split the data for cohort
```{r, message =F, warning =F}
############
## CROSS ##
##########
spec.cross<-as.data.frame(t(spec.cross))
all(spec.cross %>% rownames() == meta.cross$SmplID)

# add the cohort column to split the data
spec.cross$cohort<-meta.cross$Cohort_precise

# split
l_spec.cross<-split(spec.cross, f = as.factor(spec.cross$cohort))

############
## Long ##
###########

spec.long<-as.data.frame(t(spec.long))
all(spec.long%>% rownames() == meta.long$SmplID)

# add the cohort column to split the data
spec.long$cohort<-meta.long$Cohort_precise

# split
l_spec.long<-split(spec.long, f = as.factor(spec.long$cohort))

```

# Plot taxa abundances against inflammation index

```{r}
all(colnames(spec.cross)==colnames(spec.long))
spec_cross_long<-rbind(spec.cross,spec.long)%>%
  rownames_to_column("SmplID")

specAb_meta<-spec_cross_long%>%left_join(meta_hr, by="SmplID")%>%
  gather(species,abundance,c(colnames(spec_cross_long[-1])))
specAb_meta$HumanRds%>%class()
specAb_meta$abundance<-as.numeric(specAb_meta$abundance)
specAb_meta$abundance%>%class()
specAb_meta$species%>%class()
```

# Spearman correlation of taxa abundances and human reads

## crosssectional - per cohort

```{r, warning=FALSE}
# Calculate correlation crossectional
l_spec.cross<-lapply(l_spec.cross, function(x) x[,-1*which(names(x) == "cohort")])

l_cor_cross<-lapply(names(l_spec.cross), function(a){
  corr.df<-corr_all_chrts(a, l_spec.cross, meta_hr, "HumanRds")
  corr.df$padjust<-p.adjust(corr.df$pvalue, method="BH")
  corr.df
})
names(l_cor_cross)<-names(l_spec.cross)
l_cor_cross$FedericiGermany

```

## crosssectional - per cohort - IBD only

```{r, warning=FALSE}
#Function to calculate correlation crossectional - IBD indiv only
meta_ibd<-meta%>%
  filter(IBD=="IBD")
meta_ibd_hr<-inner_join(meta_ibd,humanReads,by="SmplID")

l_cor_cross_IBD<-lapply(names(l_spec.cross), function(a){
  corr.df<-corr_all_chrts(a, l_spec.cross, meta_ibd_hr, "HumanRds")
  corr.df$padjust<-p.adjust(corr.df$pvalue, method="BH")
  corr.df
})
names(l_cor_cross_IBD)<-names(l_spec.cross)
l_cor_cross_IBD$FedericiGermany

```

## crosssectional - per cohort - controls only

```{r, warning=FALSE}
#Function to calculate correlation crossectional - control indiv only
meta_ctrl<-meta%>%
  filter(IBD=="healthy")
meta_ctrl_hr<-inner_join(meta_ctrl,humanReads,by="SmplID")

l_cor_cross_CTRL<-lapply(names(l_spec.cross), function(a){
  corr.df<-corr_all_chrts(a, l_spec.cross, meta_ctrl_hr, "HumanRds")
  corr.df$padjust<-p.adjust(corr.df$pvalue, method="BH")
  corr.df
})
names(l_cor_cross_CTRL)<-names(l_spec.cross)
l_cor_cross_CTRL$FedericiGermany

```

## longitudinal

### load resamplings


```{r, warning=FALSE}
# load resampling
df.row<-read.delim2(paste0(PATHanalNov2023,"PrepData/resampling.CDUC.20.txt"),sep="\t")
# remove cohort column
l_spec.long<-lapply(l_spec.long, function(x) x[,-1*which(names(x) == "cohort")])

# make list with dataframe subset according to resampling
l_spec.long.subsamps<-lapply(names(df.row[1:20]), function(s){
  l_spec.long.samp<-lapply(names(l_spec.long), function(x){
    subsamps<-(df.row%>%filter(Cohort==x))[[s]]
    df_spec.samp<-l_spec.long[[x]][subsamps,]
    return(df_spec.samp)
  })
  names(l_spec.long.samp)<-names(l_spec.long)
  return(l_spec.long.samp)
})
names(l_spec.long.subsamps)<-names(df.row[1:20])
```

## longitudinal - per cohort and 20 resamplings

```{r}
l_cor_long.resamp<-lapply(names(l_spec.long.subsamps), function(s){
  l_cor_long<-lapply(names(l_spec.long.subsamps[[s]]), function(a){
     corr_long<-corr_all_chrts(a, l_spec.long.subsamps[[s]], meta_hr, "HumanRds") # l_spec.long
    corr_long$padjust<-p.adjust(corr_long$pvalue, method="BH")
     return(corr_long)
  })
  names(l_cor_long)<-names(l_spec.long.subsamps[[s]])
  return(l_cor_long)
})
names(l_cor_long.resamp)<-names(l_spec.long.subsamps)
l_cor_long.resamp$Sample3$HMP2
```
## longitudinal - per cohort and 20 resamplings - only IBD

```{r, warning=FALSE}
# only IBD

l_cor_long.resamp_IBD<-lapply(names(l_spec.long.subsamps), function(s){
  l_cor_long<-lapply(names(l_spec.long.subsamps[[s]]), function(a){
     corr_long<-corr_all_chrts(a, l_spec.long.subsamps[[s]], meta_ibd_hr, "HumanRds") # l_spec.long
    corr_long$padjust<-p.adjust(corr_long$pvalue, method="BH")
     return(corr_long)
  })
  names(l_cor_long)<-names(l_spec.long.subsamps[[s]])
  return(l_cor_long)
})
names(l_cor_long.resamp_IBD)<-names(l_spec.long.subsamps)
l_cor_long.resamp_IBD$Sample3$HMP2
```
## longitudinal - per cohort and 20 resamplings - only controls

```{r, warning=FALSE}
# only CTRL
cohorts_remove_cduc<-(meta_ctrl_hr%>%filter(Cohort_setup=="longitudinal"))$Cohort_precise%>%unique()

l_cor_long.resamp_CTRL<-lapply(names(l_spec.long.subsamps), function(s){
  l_cor_long<-lapply(cohorts_remove_cduc, function(a){
    corr_long<-corr_all_chrts(a, l_spec.long.subsamps[[s]], meta_ctrl_hr, "HumanRds") # l_spec.long
    corr_long$padjust<-p.adjust(corr_long$pvalue, method="BH")
    return(corr_long)
  })
  names(l_cor_long)<-cohorts_remove_cduc
  return(l_cor_long)
})
names(l_cor_long.resamp_CTRL)<-names(l_spec.long.subsamps)
l_cor_long.resamp_CTRL$Sample3$HMP2
```

## combine crosssectional and longitudinal lists

### all

```{r}
l_cor_all<-list(l_cor_cross,l_cor_long.resamp)
names(l_cor_all)<-c("cross","long")
saveRDS(l_cor_all,paste0(PATHanalNov2023,"humanReads/humRds_cor_MGS_list.rds"))
#l_cor_all<-readRDS(paste0(PATHanalNov2023,"humanReads/humRds_cor_MGS_list.rds"))

l_cor_all$long$Sample4$HMP

```

### IBD only

```{r}
l_cor_all_IBD<-list(l_cor_cross_IBD,l_cor_long.resamp_IBD)
names(l_cor_all_IBD)<-c("cross","long")
saveRDS(l_cor_all_IBD,paste0(PATHanalNov2023,"humanReads/humRds_cor_MGS_list_IBD.rds"))
#l_cor_all<-readRDS(paste0(PATHanalNov2023,"humanReads/humRds_cor_MGS_list.rds"))

l_cor_all_IBD$long$Sample4$HMP

```
### controls only

```{r}
l_cor_all_CTRL<-list(l_cor_cross_CTRL,l_cor_long.resamp_CTRL)
names(l_cor_all_CTRL)<-c("cross","long")
saveRDS(l_cor_all_CTRL,paste0(PATHanalNov2023,"humanReads/humRds_cor_MGS_list_CTRL.rds"))
#l_cor_all<-readRDS(paste0(PATHanalNov2023,"humanReads/humRds_cor_MGS_list.rds"))

l_cor_all_CTRL$long$Sample4$HMP

```

# combine correlation results with differential abundance results

## crosssectional

```{r}

l_cor_cross_adjust<-l_cor_cross
df_cor_cross<-do.call(rbind, l_cor_cross_adjust)
df_cor_cross_DAibd<-df_cor_cross%>%
  full_join(ibd_ctrl.sig, by="MGS")

```

## longitudinal

```{r}
l_cor_long_adjust<-lapply(names(l_cor_long.resamp), function(s){
    l_cor_long_adjtmp<-lapply(names(l_cor_long.resamp[[s]]), function(x){
    corstmp<-l_cor_long.resamp[[s]][[x]]
    #corstmp$padjust<-p.adjust(corstmp$pvalue, method = "BH") # already adjusted before
    return(corstmp)
  })
  names(l_cor_long_adjtmp)<-names(l_cor_long.resamp[[s]])
  df_cor_long_samp<-do.call(rbind, l_cor_long_adjtmp)
  return(df_cor_long_samp)
})
names(l_cor_long_adjust)<-names(l_cor_long.resamp)

siglist<-list(ibd_ctrl.sig,cd_ctrl.sig,uc_ctrl.sig,cd_uc.sig)
names(siglist)<-c("ibd_ctrl","cd_ctrl","uc_ctrl","cd_uc")
  
corDA_df_long<-lapply(names(l_cor_long_adjust), function(x){
  df_cor_long<-l_cor_long_adjust[[x]]
  corDA_plot_long_samp<-lapply(siglist, function(s){
  
    df.filt<-df_cor_long%>%
      full_join(s, by="MGS")%>%
      filter(padjust<qval)
    
  
    return(df.filt)
  })
  names(corDA_plot_long_samp)<-names(siglist)
  return(corDA_plot_long_samp)
})
names(corDA_df_long)<-names(l_cor_long_adjust)
corDA_df_long$Sample1

```


## Evaluate numbers of species with significant correlation of abundance and human reads and significant DA

### crosssectional 

```{r}
# count how many times MGS significant across cohorts
# select all that are significant in 1 or more cohorts

numSigCorr<-df_cor_cross_DAibd%>%
  mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
  group_by(MGS) %>%
  summarise(num_sigcorr = sum(sigcorr == "significant"))%>%
  left_join(df_cor_cross_DAibd, by="MGS")

cross_corsig<-numSigCorr%>%filter(num_sigcorr>0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
cross_corsig_pos<-numSigCorr%>%filter(num_sigcorr>0)%>%filter(padjust<qval)%>%filter(rho>0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
cross_corsig_neg<-numSigCorr%>%filter(num_sigcorr>0)%>%filter(padjust<qval)%>%filter(rho<0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS

cross_corsig_robDA<-numSigCorr%>%filter(num_sigcorr>0 & robustDA=="robust")%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and robust DA 
cross_corsig_varDA<-numSigCorr%>%filter(num_sigcorr>0 & robustDA=="variable")%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and variable DA
cross_corsig_notDA<-numSigCorr%>%filter(num_sigcorr>0 & is.na(robustDA))%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and not DA

cross_notcorr<-numSigCorr%>%filter(num_sigcorr==0)%>%select(MGS)%>%unique()%>%nrow() # nonsignificant corr MGS
cross_allmgs<-numSigCorr%>%select(MGS)%>%unique()%>%nrow() # all MGS

cross_compare_CORvsDA<-data.frame(
  Cor=cross_corsig,
  Cor_robDA=cross_corsig_robDA,
  Cor_varDA=cross_corsig_varDA,
  Cor_notDA=cross_corsig_notDA,
  NoCor=cross_notcorr,
  numMGS=cross_allmgs
)
cross_compare_CORvsDA
```

### export dataframes with results

```{r}
df_cor_cross_DAibd_sigs<-df_cor_cross_DAibd%>%
  mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
  filter(sigcorr=="significant")%>%  
  mutate(rho2=rho)%>%
  gather(measure,effect,c(rho,meanTE))%>%
  arrange(MGS, rho2)
write.csv(df_cor_cross_DAibd_sigs, paste0(PATHanalNov2023,"Downstream/Results/MGScorvsDAibd.q0.1.csv"))

# save table with all info 
df_cor_cross_DAibd_all<-df_cor_cross_DAibd%>%
  full_join(da_all.sig, by="MGS")
df_cor_cross_DAibd_all_sigs<-df_cor_cross_DAibd_all%>%
  mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
  filter(sigcorr=="significant")%>%  
  mutate(rho2=rho)%>%
  gather(measure,effect,c(rho,meanTE.IBDHC,meanTE.CDHC,meanTE.UCHC,meanTE.CDUC))%>%
  arrange(MGS, rho2)
```

### longitudinal

```{r}
# count how many times MGS significant across cohorts
# select all that are significant in 1 or more cohorts
# list longitudinal
numSigCorr.l_long<-lapply(corDA_df_long, function (x){
  numSigCorr<-x[["ibd_ctrl"]]%>%
    mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
    group_by(MGS) %>%
    summarise(num_sigcorr = sum(sigcorr == "significant"))%>%
    left_join(x[["ibd_ctrl"]], by="MGS")

  numSigCorr
})
names(numSigCorr.l_long)<-names(corDA_df_long)
#numSigCorr.l_long$Sample2

long_compare_CORvsDA.l_long<-lapply(numSigCorr.l_long, function (x){
  #x<-numSigCorr.l_long[[a]]
  long_corsig<-x%>%filter(num_sigcorr>0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
  long_corsig_pos<-x%>%filter(num_sigcorr>0)%>%filter(padjust<qval)%>%filter(rho>0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
  long_corsig_neg<-x%>%filter(num_sigcorr>0)%>%filter(padjust<qval)%>%filter(rho<0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
  
  long_corsig_robDA<-x%>%filter(num_sigcorr>0 & robustDA=="robust")%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and robust DA 
  long_corsig_varDA<-x%>%filter(num_sigcorr>0 & robustDA=="variable")%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and variable DA
  long_corsig_notDA<-x%>%filter(num_sigcorr>0 & is.na(robustDA))%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and not DA
  
  long_notcorr<-x%>%filter(num_sigcorr==0)%>%select(MGS)%>%unique()%>%nrow() # nonsignificant corr MGS
  long_allmgs<-x%>%select(MGS)%>%unique()%>%nrow() # all MGS
  
  long_compare_CORvsDA<-data.frame(
    Cor=long_corsig,
    Cor_robDA=long_corsig_robDA,
    Cor_varDA=long_corsig_varDA,
    Cor_notDA=long_corsig_notDA,
    NoCor=long_notcorr,
    numMGS=long_allmgs
  )
  long_compare_CORvsDA
  })
names(long_compare_CORvsDA.l_long)<-names(numSigCorr.l_long)
long_compare_CORvsDA.df<-do.call(rbind,long_compare_CORvsDA.l_long)
long_compare_CORvsDA.df

df_cor_long_DAibd_sigs_l_long<-lapply(numSigCorr.l_long, function (x){
  df_cor_long_DAibd_sigs<-x%>%
  mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
  filter(sigcorr=="significant")%>%  
  mutate(rho2=rho)%>%
  gather(measure,effect,c(rho,meanTE))%>%
  arrange(MGS, rho2)
  
  df_cor_long_DAibd_sigs

})
names(df_cor_long_DAibd_sigs_l_long)<-names(numSigCorr.l_long)

df_cor_long_DAibd_sigs2_l_long<-lapply(numSigCorr.l_long, function (x){
  df_cor_long_DAibd_sigs2<-x%>%
    mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
    filter(sigcorr=="significant")%>%  
    filter(num_sigcorr>1)%>%  
    mutate(rho2=rho)%>%
    gather(measure,effect,c(rho,meanTE))%>%
    arrange(MGS, rho2)
  
  df_cor_long_DAibd_sigs2
})
names(df_cor_long_DAibd_sigs2_l_long)<-names(numSigCorr.l_long)

# save table with all info
df_cor_long_DAibd_all_l_long<-lapply(numSigCorr.l_long, function (x){
  df_cor_long_DAibd_all<-x%>%
    full_join(da_all.sig, by="MGS")
  df_cor_long_DAibd_all_sigs<-df_cor_long_DAibd_all%>%
    mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
    filter(sigcorr=="significant")%>%  
    mutate(rho2=rho)%>%
    gather(measure,effect,c(rho,meanTE.IBDHC,meanTE.CDHC,meanTE.UCHC,meanTE.CDUC))%>%
    arrange(MGS, rho2)
  
  df_cor_long_DAibd_all_sigs
})
names(df_cor_long_DAibd_all_l_long)<-names(numSigCorr.l_long)

```


```{r fig.width=8, fig.height=7}

cor_long_DAibd_all_sigs.list<-lapply(names(df_cor_long_DAibd_all_l_long), function (x){
  x.df<-df_cor_long_DAibd_all_l_long[[x]]%>%
    mutate(meanTE=ifelse(measure=="rho",NA,effect))%>%
    mutate(effect=ifelse(IBDHC=="non-sig" & measure=="meanTE.IBDHC",NA,effect))%>%
    mutate(effect=ifelse(CDHC=="non-sig" & measure=="meanTE.CDHC",NA,effect))%>%
    mutate(effect=ifelse(UCHC=="non-sig" & measure=="meanTE.UCHC",NA,effect))%>%
    mutate(effect=ifelse(CDUC=="non-sig" & measure=="meanTE.CDUC",NA,effect))%>%
    mutate(robustDA=ifelse(measure=="meanTE.IBDHC",IBDHC,NA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.CDHC",CDHC,robustDA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.UCHC",UCHC,robustDA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.CDUC",CDUC,robustDA))%>%
    mutate(robustDA=ifelse(is.na(robustDA),"corr",robustDA))%>%
    mutate(sample=x)
  
  x.df$effect<-as.numeric(x.df$effect)
  x.df
})
names(cor_long_DAibd_all_sigs.list)<-names(df_cor_long_DAibd_all_l_long)

cor_long_DAibd_all_sigs.df<-do.call(rbind,cor_long_DAibd_all_sigs.list)

```

### combine crosssectional and longitudinal results

```{r}
all(names(df_cor_cross_DAibd_all_sigs%>%mutate(sample="cross"))==names(cor_long_DAibd_all_sigs.df%>%select(-num_sigcorr)))

df_cor_all_DAibd_all_sigs<-rbind(df_cor_cross_DAibd_all_sigs%>%mutate(sample="cross"),cor_long_DAibd_all_sigs.df%>%select(-num_sigcorr))

length(unique(df_cor_all_DAibd_all_sigs$MGS)) 
```

### Export dataframes

```{r}
write.csv(df_cor_all_DAibd_all_sigs, paste0(PATHanalNov2023,"Downstream/Results/MGScorvsDAibd.q0.1.crosslong.csv"))
saveRDS(df_cor_all_DAibd_all_sigs, paste0(PATHanalNov2023,"humanReads/MGScorvsDAibd.crosslong.q0.1.rds"))

# Get some overview numbers
testdf<-df_cor_all_DAibd_all_sigs%>%
  filter(robustDA=="corr")%>%
  group_by(MGS,cohort)%>%
  count()

# how many MGS more than 1 cohort
testdf%>%group_by(MGS)%>%count()%>%filter(n>1)%>%nrow()

# direction
testdf2<-df_cor_all_DAibd_all_sigs%>%
  filter(robustDA=="corr")%>%
  mutate(direction=ifelse(rho2<0,"negative","positive"))%>%
  select(MGS,direction)%>%
  group_by(MGS,direction)%>%
  count()
# positive
testdf2%>%filter(direction=="positive")%>%nrow()
# negative
testdf2%>%filter(direction=="negative")%>%nrow()
# which MGS have discrepancy
testdf2%>%group_by(MGS)%>%count()%>%filter(n>1)
discrMGS<-(testdf2%>%group_by(MGS)%>%count()%>%filter(n>1))$MGS

# which MGS have no discrepancy
testdf2%>%group_by(MGS)%>%count()%>%filter(n==1)%>%nrow()

# MGS significant in DA and correlation
(df_cor_all_DAibd_all_sigs%>%filter(!is.na(effect))%>%filter(measure!="rho"))$MGS%>%unique()%>%length()
sum(discrMGS%in%(df_cor_all_DAibd_all_sigs%>%filter(!is.na(effect))%>%filter(measure!="rho"))$MGS)
```

### create plot with combined results

```{r fig.width=8, fig.height=15}
df_cor_all_DAibd_all_sigs<-df_cor_all_DAibd_all_sigs%>%
  mutate(effect=as.numeric(effect))

df_cor_all_DAibd_all_sigs2<-df_cor_all_DAibd_all_sigs%>%
  mutate(effect=ifelse(measure=="meanTE.IBDHC" & IBDHC=="non-sig",NA,effect),
         effect=ifelse(measure=="meanTE.CDHC" & CDHC=="non-sig",NA,effect),
         effect=ifelse(measure=="meanTE.UCHC" & UCHC=="non-sig",NA,effect),
         effect=ifelse(measure=="meanTE.CDUC" & CDUC=="non-sig",NA,effect),
         significance=ifelse(
           measure=="meanTE.IBDHC",paste0(robustDA," IBD-associated"), ifelse(
             measure=="meanTE.CDHC",paste0(CDHC," IBD-associated"), ifelse(
               measure=="meanTE.UCHC",paste0(UCHC," IBD-associated"),ifelse(
                 measure=="meanTE.CDUC",paste0(CDUC," IBD-associated"),ifelse(
                   measure=="rho","sig. correlated","x"
                 )
               )
             )
           )
         ),
         significance=ifelse(significance=="non-sig IBD-associated",NA,significance),
         measure=ifelse(measure=="meanTE.IBDHC","meanTE IBD - control",measure),
         measure=ifelse(measure=="meanTE.CDHC","meanTE CD - control",measure),
         measure=ifelse(measure=="meanTE.UCHC","meanTE UC - control",measure),
         measure=ifelse(measure=="meanTE.CDUC","meanTE CD - UC",measure))


p_cor_all_DAibd_all_sigs<-ggplot(df_cor_all_DAibd_all_sigs2,
   aes(effect,reorder(gsub(".*;","",gsub(";MGS."," MGS.",MGS)),rho2),color=cohort,shape=significance))+
    geom_vline(xintercept=0, linetype=2, color="grey50")+
    geom_point(size=3, alpha=0.7) +
    scale_shape_manual(values=shape_sig)+
    facet_grid(.~measure, scales = "free_x") +
    ylab("MGS") +
    scale_x_reverse() +
    theme_bw()
p_cor_all_DAibd_all_sigs
legened_p_cor_all_DAibd_all_sigs<-get_legend(p_cor_all_DAibd_all_sigs)

p_cor_all_DAibd_all_sigs_DAonly<-ggplot(df_cor_all_DAibd_all_sigs2%>%filter(measure!="rho"),
   aes(effect,reorder(gsub(".*;","",gsub(";MGS."," MGS.",MGS)),rho2),shape=significance))+
    geom_vline(xintercept=0, linetype=2, color="grey50")+
    geom_point(size=3) +
    scale_shape_manual(values=shape_sig)+
    facet_grid(.~measure, scales = "free_x") +
    ylab("MGS") +
    scale_x_reverse() +
    theme_bw()+
    theme(legend.position = "none")
p_cor_all_DAibd_all_sigs_DAonly

p_cor_all_DAibd_all_sigs_rhoonly<-ggplot(df_cor_all_DAibd_all_sigs2%>%filter(measure=="rho"),
   aes(effect,reorder(gsub(".*;","",gsub(";MGS."," MGS.",MGS)),rho2),color=cohort,shape=significance))+
    geom_vline(xintercept=0, linetype=2, color="grey50")+
    geom_point(size=3, alpha=0.7) +
    scale_shape_manual(values=shape_sig)+
    facet_grid(.~measure, scales = "free_x") +
    ylab("MGS") +
    theme_bw()+
    theme(axis.text.y=element_blank(),
          axis.title.y = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = "none")
p_cor_all_DAibd_all_sigs_rhoonly

p_cor_all_DAibd_all_sigs_combi<-plot_grid(p_cor_all_DAibd_all_sigs_DAonly,p_cor_all_DAibd_all_sigs_rhoonly,legened_p_cor_all_DAibd_all_sigs,rel_widths = c(5,1,1), nrow=1)
```

### export plot with combined results

```{r, results=FALSE,message=FALSE,warning=FALSE}
pdf(paste0(PATHanalNov2023,"Downstream/Results/Figures/MGScorvsDAibd2_long_cross.q0.1.pdf"),height=20,width=14)
p_cor_all_DAibd_all_sigs_combi
dev.off()
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/MGScorvsDAibd2_long_cross.q0.1.png"),p_cor_all_DAibd_all_sigs_combi,height=20,width=14)
```

```{r fig.width=7, fig.height=6}
df_cor_all_DAibd_rhoVSmeanTE<-df_cor_all_DAibd_all_sigs%>%
  filter(!is.na(effect))%>%
  select(c(MGS,cohort,measure,effect,sample))%>%
  spread(measure,effect)%>%
  gather(comparison,meanTE,c("meanTE.CDUC","meanTE.IBDHC","meanTE.UCHC","meanTE.CDHC"))%>%
  mutate(comparison=gsub("meanTE.","",comparison))%>%
  mutate(comparison=gsub("HC"," vs. control",comparison))%>%
  mutate(comparison=gsub("CDUC","CD vs. UC",comparison))

(df_cor_all_DAibd_rhoVSmeanTE%>%filter(!is.na(meanTE)))$MGS%>%unique()%>%length()
```


```{r}
# get correlation between mean TE and rho
df_cor_all_DAibd_rhoVSmeanTE$comparison%>%unique()

cduc_dfcorr<-df_cor_all_DAibd_rhoVSmeanTE%>%filter(comparison=="CD vs. UC")
cdhc_dfcorr<-df_cor_all_DAibd_rhoVSmeanTE%>%filter(comparison=="CD vs. control")
uchc_dfcorr<-df_cor_all_DAibd_rhoVSmeanTE%>%filter(comparison=="UC vs. control")
ibd_dfcorr<-df_cor_all_DAibd_rhoVSmeanTE%>%filter(comparison=="IBD vs. control")

lm_cduc<-lmerTest::lmer(meanTE ~ rho + (1|cohort), data=cduc_dfcorr)
lm_cdhc<-lmerTest::lmer(meanTE ~ rho + (1|cohort), data=cdhc_dfcorr)
lm_uchc<-lmerTest::lmer(meanTE ~ rho + (1|cohort), data=uchc_dfcorr)
lm_ibd<-lmerTest::lmer(meanTE ~ rho + (1|cohort), data=ibd_dfcorr)

sum_lm_cduc<-summary(lm_cduc)
sum_lm_cdhc<-summary(lm_cdhc)
sum_lm_uchc<-summary(lm_uchc)
sum_lm_ibd<-summary(lm_ibd)

lm_meanTE_rho<-data.frame(
  comparison=c("CD vs. UC","IBD vs. control","UC vs. control","CD vs. control"),
  p=c(sum_lm_cduc$coefficients[2,5],sum_lm_ibd$coefficients[2,5],sum_lm_uchc$coefficients[2,5],sum_lm_cdhc$coefficients[2,5]),
  rho=c(sum_lm_cduc$coefficients[2,1],sum_lm_ibd$coefficients[2,1],sum_lm_uchc$coefficients[2,1],sum_lm_cdhc$coefficients[2,1])
)
lm_meanTE_rho$padjust<-p.adjust(lm_meanTE_rho$p)
lm_meanTE_rho

cor_cduc<-cor.test((df_cor_all_DAibd_rhoVSmeanTE%>%filter(comparison=="CD vs. UC"))$meanTE, (df_cor_all_DAibd_rhoVSmeanTE%>%filter(comparison=="CD vs. UC"))$rho, use = "complete.obs", method="spearman")

cor_ibd<-cor.test((df_cor_all_DAibd_rhoVSmeanTE%>%filter(comparison=="IBD vs. control"))$meanTE, (df_cor_all_DAibd_rhoVSmeanTE%>%filter(comparison=="IBD vs. control"))$rho, use = "complete.obs", method="spearman")

cor_uchc<-cor.test((df_cor_all_DAibd_rhoVSmeanTE%>%filter(comparison=="UC vs. control"))$meanTE, (df_cor_all_DAibd_rhoVSmeanTE%>%filter(comparison=="UC vs. control"))$rho, use = "complete.obs", method="spearman")

cor_cdhc<-cor.test((df_cor_all_DAibd_rhoVSmeanTE%>%filter(comparison=="CD vs. control"))$meanTE, (df_cor_all_DAibd_rhoVSmeanTE%>%filter(comparison=="CD vs. control"))$rho, use = "complete.obs", method="spearman")

cor_meanTE_rho<-data.frame(
  comparison=c("CD vs. UC","IBD vs. control","UC vs. control","CD vs. control"),
  p=c(cor_cduc$p.value,cor_ibd$p.value,cor_uchc$p.value,cor_cdhc$p.value),
  rho=c(cor_cduc$estimate,cor_ibd$estimate,cor_uchc$estimate,cor_cdhc$es)
)
cor_meanTE_rho$padjust<-p.adjust(cor_meanTE_rho$p)
cor_meanTE_rho
```

## check impact between human reads and IBD

```{r}
log2(min((meta_hr%>%filter(HumanRds>0))$HumanRds))

meta_hr_log2<-meta_hr%>%
  mutate(log2_HumanRds=log2(HumanRds))%>%
  mutate(log2_HumanRds=ifelse(log2_HumanRds=="-Inf",-16,log2_HumanRds))%>%
  mutate(IBD2=gsub("healthy","control",IBD2))
meta_hr_log2$IBD2<-factor(meta_hr_log2$IBD2,levels=c("control","UC","CD"))
p_hr_ibd2<-ggplot(meta_hr_log2%>%filter(IBD2!="IBDU"), aes(x=IBD2,y=log2_HumanRds))+
  geom_violin(aes(fill=IBD2))+
  geom_boxplot(width=0.05)+
  theme_bw()+
  scale_fill_manual(values=col_IBD2, name="")+
  ylab("log2(Human read %)")+
  xlab("")
p_hr_ibd2
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_hr_ibd2.png"),p_hr_ibd2,width=4,height=3)
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_hr_ibd2.pdf"),p_hr_ibd2,width=4,height=3)
```

## fig correlate calprotectin and hr

```{r}
meta_hr_clp<-meta_hr%>%
  mutate(IBD2=gsub("healthy","control",IBD2))%>%
  filter(!is.na(fecal_calprotectin))
meta_hr_clp$IBD2<-factor(meta_hr_clp$IBD2,levels=c("control","UC","CD"))
meta_hr_clp$fecal_calprotectin<-as.numeric(meta_hr_clp$fecal_calprotectin)
p_hr_clp<-ggplot(meta_hr_clp%>%filter(IBD2!="IBDU"), aes(x=log2(HumanRds),y=log2(fecal_calprotectin)))+
  geom_point(aes(color=IBD2))+
  geom_smooth(se=FALSE, method = 'lm', aes(color=IBD2))+
  theme_bw()+
  scale_color_manual(values=col_IBD2, name="")+
  xlab("log2(Human read %)")+
  ylab("log2(Calprotectin)")
p_hr_clp
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_hr_clp.png"),p_hr_clp,width=4,height=3)
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_hr_clp.pdf"),p_hr_clp,width=4,height=3)
```

```{r}
p_hr_clpIBD2<-plot_grid(p_hr_clp, p_hr_ibd2, rel_widths=c(1.5,1), labels = c("A","B"))
p_hr_clpIBD2
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_hr_clp-ibd2.png"),p_hr_clpIBD2,width=8,height=4)
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_hr_clp-ibd2.pdf"),p_hr_clpIBD2,width=8,height=4)
```

```{r}
# test differences with wilcoxon test
meta_hr_log2$Cohort_precise<-as.factor(meta_hr_log2$Cohort_precise)
wt_hr_CDHC<-coin::wilcox_test(HumanRds ~ IBD2 | Cohort_precise,data=meta_hr_log2%>%filter(IBD2%in%c("control","CD")))
wt_hr_UCHC<-coin::wilcox_test(HumanRds ~ IBD2 | Cohort_precise,data=meta_hr_log2%>%filter(IBD2%in%c("control","UC")))
wt_hr_CDUC<-coin::wilcox_test(HumanRds ~ IBD2 | Cohort_precise,data=meta_hr_log2%>%filter(IBD2%in%c("UC","CD")))

wt_hr_CDHC
wt_hr_UCHC
wt_hr_CDUC

coin::statistic(wt_hr_CDHC)/sqrt(meta_hr_log2%>%filter(IBD2%in%c("control","CD"))%>%nrow())
coin::statistic(wt_hr_UCHC)/sqrt(meta_hr_log2%>%filter(IBD2%in%c("control","UC"))%>%nrow())
coin::statistic(wt_hr_CDUC)/sqrt(meta_hr_log2%>%filter(IBD2%in%c("UC","CD"))%>%nrow())

```

```{r}
#crosssectional

glm_spec_hr_blocked2<-corr_spear_blocked(spec.cross,meta_hr,"HumanRds")
glm_spec_hr_blocked2$padjust<-p.adjust(glm_spec_hr_blocked2$pvalue, method="BH")
glm_spec_hr_blocked2$spear_padjust<-p.adjust(glm_spec_hr_blocked2$spear_pvalue, method="BH")

glm_spec_hr_blocked2%>%
  filter(spear_padjust<0.1)%>%
  nrow()

glm_spec_hr_blocked2%>%
  filter(padjust<0.1)%>%
  nrow()
```

```{r}
#longitudinal
# make list with dataframe subset according to resampling
l_spec.long.subs<-lapply(names(df.row[1:20]), function(s){
  subsamps<-df.row[[s]]
  df_spec.samp<-spec.long[subsamps,]
  return(df_spec.samp)
})
#names(l_spec.long.subs)<-names(df.row[1:20])
names(l_spec.long.subs)<-names(df.row[1:20])

l_glm_blocked2_long.resamp<-lapply(names(l_spec.long.subs), function(s){
  corr_long<-corr_spear_blocked(l_spec.long.subs[[s]],meta_hr,"HumanRds")
  corr_long$padjust<-p.adjust(corr_long$pvalue, method="BH")
  corr_long$spear_padjust<-p.adjust(corr_long$spear_pvalue, method="BH")

  return(corr_long)
})
names(l_glm_blocked2_long.resamp)<-names(l_spec.long.subs)

l_glm_blocked2_long.resamp$Sample1%>%filter(padjust<qval)%>%nrow()
l_glm_blocked2_long.resamp$Sample1%>%filter(spear_padjust<qval)%>%nrow()
```

Correlate IBD-CTRL DA meanTE vs. estimate MGS correl HR

```{r}
# combine cross and long
# make list with dataframe subset according to resampling
l_spec.crosslong.subs<-lapply(names(df.row[1:20]), function(s){
  subsamps<-df.row[[s]]
  df_spec.samp<-spec.long[subsamps,]
  if(all(names(spec.cross)==names(df_spec.samp))){
    df_spec.samp.all<-rbind(df_spec.samp,spec.cross)
    return(df_spec.samp.all)
  }
})
names(l_spec.crosslong.subs)<-names(df.row[1:20])

l_glm_blocked2.resamp.all<-lapply(names(l_spec.crosslong.subs), function(s){
  corr_long<-corr_spear_blocked(l_spec.crosslong.subs[[s]],meta_hr,"HumanRds")
  corr_long$padjust<-p.adjust(corr_long$pvalue, method="BH")
  corr_long$spear_padjust<-p.adjust(corr_long$spear_pvalue, method="BH")
  return(corr_long)
})
names(l_glm_blocked2.resamp.all)<-names(l_spec.crosslong.subs)

l_glm_blocked2.resamp.all$Sample1%>%filter(padjust<0.1)%>%nrow()
l_glm_blocked2.resamp.all$Sample1%>%filter(spear_padjust<0.1)%>%nrow()
saveRDS(l_glm_blocked2.resamp.all,paste0(PATHanalNov2023,"humanReads/MGScor.crosslong.blocked.q0.1.rds"))
```

```{r}
df_glm_blocked2.resamp.all<-do.call(
  rbind,
  lapply(names(l_glm_blocked2.resamp.all), function(s){
    adds<-l_glm_blocked2.resamp.all[[s]]%>%
      mutate(sample=s)
    return(adds)
  })
)

df_glm_blocked2.resamp.all.sig<-df_glm_blocked2.resamp.all%>%
  filter(spear_padjust<0.1)%>%
  group_by(MGS)%>%
  summarize(mean_estimate=mean(estimate), mean_stdev=sd(estimate), n=length(sample))

df_glm_blocked2.resamp.all.allMGS<-df_glm_blocked2.resamp.all%>%
  group_by(MGS)%>%
  summarize(mean_estimate=mean(estimate), mean_stdev=sd(estimate))%>%
  left_join(df_glm_blocked2.resamp.all.sig%>%select(c(MGS,n)),by="MGS")%>%
  mutate(n=ifelse(is.na(n),0,n))%>%
  mutate(significance=ifelse(n==0,"non-sig",ifelse(n==20,"robust","variable")))
```

```{r}
# combine cross and long
# IBD only
# make list with dataframe subset according to resampling
l_glm_blocked2.resamp.all_IBD<-lapply(names(l_spec.crosslong.subs), function(s){
  corr_long<-corr_spear_blocked(l_spec.crosslong.subs[[s]],meta_ibd_hr,"HumanRds")
  corr_long$padjust<-p.adjust(corr_long$pvalue, method="BH")
  corr_long$spear_padjust<-p.adjust(corr_long$spear_pvalue, method="BH")
  return(corr_long)
})
names(l_glm_blocked2.resamp.all_IBD)<-names(l_spec.crosslong.subs)

l_glm_blocked2.resamp.all_IBD$Sample1%>%filter(padjust<qval)%>%nrow()
l_glm_blocked2.resamp.all_IBD$Sample1%>%filter(spear_padjust<qval)%>%nrow()
saveRDS(l_glm_blocked2.resamp.all_IBD,paste0(PATHanalNov2023,"humanReads/MGScor_onlyIBD.crosslong.blocked.q0.1.rds"))
```


```{r}
df_glm_blocked2.resamp.all_IBD<-do.call(
  rbind,
  lapply(names(l_glm_blocked2.resamp.all_IBD), function(s){
    adds<-l_glm_blocked2.resamp.all_IBD[[s]]%>%
      mutate(sample=s)
    return(adds)
  })
)

df_glm_blocked2.resamp.all_IBD.sig<-df_glm_blocked2.resamp.all_IBD%>%
  filter(spear_padjust<0.1)%>%
  group_by(MGS)%>%
  summarize(mean_estimate=mean(estimate), mean_stdev=sd(estimate), n=length(sample))

df_glm_blocked2.resamp.all_IBD.allMGS<-df_glm_blocked2.resamp.all_IBD%>%
  group_by(MGS)%>%
  summarize(mean_estimate=mean(estimate), mean_stdev=sd(estimate))%>%
  left_join(df_glm_blocked2.resamp.all_IBD.sig%>%select(c(MGS,n)),by="MGS")%>%
  mutate(n=ifelse(is.na(n),0,n))%>%
  mutate(significance=ifelse(n==0,"non-sig",ifelse(n==20,"robust","variable")))
```

```{r}
# combine cross and long
# controls only
# make list with dataframe subset according to resampling
l_glm_blocked2.resamp.all_CTRL<-lapply(names(l_spec.crosslong.subs), function(s){
  corr_long<-corr_spear_blocked(l_spec.crosslong.subs[[s]],meta_ctrl_hr,"HumanRds")
  corr_long$padjust<-p.adjust(corr_long$pvalue, method="BH")
  corr_long$spear_padjust<-p.adjust(corr_long$spear_pvalue, method="BH")
  return(corr_long)
})
names(l_glm_blocked2.resamp.all_CTRL)<-names(l_spec.crosslong.subs)

l_glm_blocked2.resamp.all_CTRL$Sample1%>%filter(padjust<qval)%>%nrow()
l_glm_blocked2.resamp.all_CTRL$Sample2%>%filter(spear_padjust<qval)%>%nrow()
saveRDS(l_glm_blocked2.resamp.all_CTRL,paste0(PATHanalNov2023,"humanReads/MGScor_onlyCTRL.crosslong.blocked.0.1.rds"))
```

```{r}
df_glm_blocked2.resamp.all_CTRL<-do.call(
  rbind,
  lapply(names(l_glm_blocked2.resamp.all_CTRL), function(s){
    adds<-l_glm_blocked2.resamp.all_CTRL[[s]]%>%
      mutate(sample=s)
    return(adds)
  })
)

df_glm_blocked2.resamp.all_CTRL.sig<-df_glm_blocked2.resamp.all_CTRL%>%
  filter(spear_padjust<0.1)%>%
  group_by(MGS)%>%
  summarize(mean_estimate=mean(estimate), mean_stdev=sd(estimate), n=length(sample))

df_glm_blocked2.resamp.all_CTRL.allMGS<-df_glm_blocked2.resamp.all_CTRL%>%
  group_by(MGS)%>%
  summarize(mean_estimate=mean(estimate), mean_stdev=sd(estimate))%>%
  left_join(df_glm_blocked2.resamp.all_CTRL.sig%>%select(c(MGS,n)),by="MGS")%>%
  mutate(n=ifelse(is.na(n),0,n))%>%
  mutate(significance=ifelse(n==0,"non-sig",ifelse(n==20,"robust","variable")))
```

```{r}
# combine with DA
DA_glm_blocked2.all<-da_all.sig%>%
  full_join(df_glm_blocked2.resamp.all.allMGS,by="MGS")%>%
  rename(meanEstimate.all=mean_estimate)%>%
  rename(meanStdev.all=mean_stdev)%>%
  rename(n.all=n)%>%
  rename(significance.all=significance)%>%
  full_join(df_glm_blocked2.resamp.all_IBD.allMGS,by="MGS")%>%
  rename(meanEstimate.IBD=mean_estimate)%>%
  rename(meanStdev.IBD=mean_stdev)%>%
  rename(n.IBD=n)%>%
  rename(significance.IBD=significance)%>%
  full_join(df_glm_blocked2.resamp.all_CTRL.allMGS,by="MGS")%>%
  rename(meanEstimate.CTRL=mean_estimate)%>%
  rename(meanStdev.CTRL=mean_stdev)%>%
  rename(n.CTRL=n)%>%
  rename(significance.CTRL=significance)

DA_glm_blocked2.all<-DA_glm_blocked2.all%>%
  mutate(MGS=gsub(";MGS"," (MGS",MGS))%>%
  mutate(MGS=gsub(".*;","",MGS))%>%
  mutate(MGS=paste0(MGS,")"))

DA_glm_blocked2.all$meanTE.IBDHC<-as.numeric(DA_glm_blocked2.all$meanTE.IBDHC)
DA_glm_blocked2.all$meanTE.CDHC<-as.numeric(DA_glm_blocked2.all$meanTE.CDHC)
DA_glm_blocked2.all$meanTE.UCHC<-as.numeric(DA_glm_blocked2.all$meanTE.UCHC)
DA_glm_blocked2.all$meanTE.CDUC<-as.numeric(DA_glm_blocked2.all$meanTE.CDUC)

```

```{r}
# create plot
ggplot(DA_glm_blocked2.all%>%
         filter(significance.IBD!="non-sig" | significance.CTRL!="non-sig")%>%
         mutate(significance=paste0(significance.IBD,"_",significance.CTRL)),
       aes(meanEstimate.IBD,meanEstimate.CTRL))+
  geom_hline(yintercept = 0, linetype=2, color="grey50")+
  geom_vline(xintercept = 0, linetype=2, color="grey50")+
  geom_point(aes(color=significance))+
  geom_smooth(method="lm")+
  theme_bw()
```

```{r}
# IBD-CTRL DA meanTE vs. estimate MGS correl HR
ggplot(DA_glm_blocked2.all%>%
         filter(IBDHC!="non-sig" | significance.all!="non-sig"),#%>%
         #mutate(significance=paste0(significance.IBD,"_",significance.CTRL)),
       aes(meanEstimate.all,meanTE.IBDHC))+
  geom_hline(yintercept = 0, linetype=2, color="grey50")+
  geom_vline(xintercept = 0, linetype=2, color="grey50")+
  geom_point()+#aes(color=significance))+
  geom_smooth(method="lm", se = FALSE)+
  theme_bw()
```



```{r}
# IBD-CTRL DA meanTE vs. estimate MGS correl HR in IBD only
ggplot(DA_glm_blocked2.all%>%
         filter(IBDHC!="non-sig" | significance.IBD!="non-sig"),#%>%
         #mutate(significance=paste0(significance.IBD,"_",significance.CTRL)),
       aes(meanEstimate.IBD,meanTE.IBDHC))+
  geom_hline(yintercept = 0, linetype=2, color="grey50")+
  geom_vline(xintercept = 0, linetype=2, color="grey50")+
  geom_point()+#aes(color=significance))+
  geom_smooth(method="lm", se = FALSE)+
  theme_bw()
```

```{r}
# IBD-CTRL DA meanTE vs. estimate MGS correl HR in CTRL only
ggplot(DA_glm_blocked2.all%>%
         filter(IBDHC!="non-sig" | significance.CTRL!="non-sig"),#%>%
         #mutate(significance=paste0(significance.IBD,"_",significance.CTRL)),
       aes(meanEstimate.CTRL,meanTE.IBDHC))+
  geom_hline(yintercept = 0, linetype=2, color="grey50")+
  geom_vline(xintercept = 0, linetype=2, color="grey50")+
  geom_point()+#aes(color=significance))+
  geom_smooth(method="lm", se = FALSE)+
  theme_bw()
```

```{r fig.width=15, fig.height=12}

max(DA_glm_blocked2.all$meanEstimate.all)
min(DA_glm_blocked2.all$meanEstimate.all)

DA_glm_blocked2.all_gat<-DA_glm_blocked2.all%>%
  filter(significance.all!="non-sig" | significance.IBD!="non-sig" | significance.CTRL!="non-sig")%>%
  filter(abs(meanEstimate.all)>0.002)%>%
  mutate(meanEstimate.all=ifelse(significance.all=="non-sig",NA,meanEstimate.all))%>%
  mutate(meanEstimate.IBD=ifelse(significance.IBD=="non-sig",NA,meanEstimate.IBD))%>%
  mutate(meanEstimate.CTRL=ifelse(significance.CTRL=="non-sig",NA,meanEstimate.CTRL))%>%
  mutate(meanTE.IBDHC=ifelse(IBDHC=="non-sig",NA,meanTE.IBDHC))%>%
  mutate(meanTE.CDHC=ifelse(IBDHC=="non-sig",NA,meanTE.CDHC))%>%
  mutate(meanTE.UCHC=ifelse(IBDHC=="non-sig",NA,meanTE.UCHC))%>%
  mutate(meanTE.CDUC=ifelse(IBDHC=="non-sig",NA,meanTE.CDUC))%>%
  gather(test,effect, c(meanTE.IBDHC,meanTE.CDHC,meanTE.UCHC,meanTE.CDUC,meanEstimate.all,meanEstimate.IBD,meanEstimate.CTRL))


DA_glm_blocked2.all_gat$MGS<-factor(DA_glm_blocked2.all_gat$MGS,levels = (DA_glm_blocked2.all%>%arrange(desc(meanEstimate.all)))$MGS)
DA_glm_blocked2.all_gat$test<-factor(DA_glm_blocked2.all_gat$test,levels=c("meanEstimate.all","meanEstimate.IBD","meanEstimate.CTRL","meanTE.IBDHC","meanTE.CDHC","meanTE.UCHC","meanTE.CDUC"))

p_da_blocked.all_est0.002<-ggplot(DA_glm_blocked2.all_gat,#%>%
         #mutate(significance=paste0(significance.IBD,"_",significance.CTRL)),
       aes(effect,MGS))+
  geom_hline(yintercept = 0, linetype=2, color="grey50")+
  geom_vline(xintercept = 0, linetype=2, color="grey50")+
  geom_point()+#aes(color=significance))+
  theme_bw()+
  facet_grid(.~test, scales = "free_x")
p_da_blocked.all_est0.002
```

```{r}
DA_glm_blocked2.all_IBDHC<-DA_glm_blocked2.all%>%filter(IBDHC!="non-sig" | significance.all!="non-sig")
DA_glm_blocked2.all_CDHC<-DA_glm_blocked2.all%>%filter(CDHC!="non-sig" | significance.all!="non-sig")
DA_glm_blocked2.all_UCHC<-DA_glm_blocked2.all%>%filter(UCHC!="non-sig" | significance.all!="non-sig")
DA_glm_blocked2.all_CDUC<-DA_glm_blocked2.all%>%filter(CDUC!="non-sig" | significance.all!="non-sig")

DA_glm_blocked2.all_IBDHC_IBD<-DA_glm_blocked2.all%>%filter(IBDHC!="non-sig" | significance.IBD!="non-sig")
DA_glm_blocked2.all_CDHC_IBD<-DA_glm_blocked2.all%>%filter(CDHC!="non-sig" | significance.IBD!="non-sig")
DA_glm_blocked2.all_UCHC_IBD<-DA_glm_blocked2.all%>%filter(UCHC!="non-sig" | significance.IBD!="non-sig")
DA_glm_blocked2.all_CDUC_IBD<-DA_glm_blocked2.all%>%filter(CDUC!="non-sig" | significance.IBD!="non-sig")

DA_glm_blocked2.all_IBDHC_CTRL<-DA_glm_blocked2.all%>%filter(IBDHC!="non-sig" | significance.CTRL!="non-sig")
DA_glm_blocked2.all_CDHC_CTRL<-DA_glm_blocked2.all%>%filter(CDHC!="non-sig" | significance.CTRL!="non-sig")
DA_glm_blocked2.all_UCHC_CTRL<-DA_glm_blocked2.all%>%filter(UCHC!="non-sig" | significance.CTRL!="non-sig")
DA_glm_blocked2.all_CDUC_CTRL<-DA_glm_blocked2.all%>%filter(CDUC!="non-sig" | significance.CTRL!="non-sig")

lm_ibdhc.all<-lm(meanTE.IBDHC ~ meanEstimate.all, data=DA_glm_blocked2.all_IBDHC)
lm_cdhc.all<-lm(meanTE.CDHC ~ meanEstimate.all, data=DA_glm_blocked2.all_CDHC)
lm_uchc.all<-lm(meanTE.UCHC ~ meanEstimate.all, data=DA_glm_blocked2.all_UCHC)
lm_cduc.all<-lm(meanTE.CDUC ~ meanEstimate.all, data=DA_glm_blocked2.all_CDUC)
sum_lm_ibdhc.all<-summary(lm_ibdhc.all)
sum_lm_cdhc.all<-summary(lm_cdhc.all)
sum_lm_uchc.all<-summary(lm_uchc.all)
sum_lm_cduc.all<-summary(lm_cduc.all)

lm_ibdhc.all_IBD<-lm(meanTE.IBDHC ~ meanEstimate.IBD, data=DA_glm_blocked2.all_IBDHC_IBD)
lm_cdhc.all_IBD<-lm(meanTE.CDHC ~ meanEstimate.IBD, data=DA_glm_blocked2.all_CDHC_IBD)
lm_uchc.all_IBD<-lm(meanTE.UCHC ~ meanEstimate.IBD, data=DA_glm_blocked2.all_UCHC_IBD)
lm_cduc.all_IBD<-lm(meanTE.CDUC ~ meanEstimate.IBD, data=DA_glm_blocked2.all_CDUC_IBD)
sum_lm_ibdhc.all_IBD<-summary(lm_ibdhc.all_IBD)
sum_lm_cdhc.all_IBD<-summary(lm_cdhc.all_IBD)
sum_lm_uchc.all_IBD<-summary(lm_uchc.all_IBD)
sum_lm_cduc.all_IBD<-summary(lm_cduc.all_IBD)

lm_ibdhc.all_CTRL<-lm(meanTE.IBDHC ~ meanEstimate.CTRL, data=DA_glm_blocked2.all_IBDHC_CTRL)
lm_cdhc.all_CTRL<-lm(meanTE.CDHC ~ meanEstimate.CTRL, data=DA_glm_blocked2.all_CDHC_CTRL)
lm_uchc.all_CTRL<-lm(meanTE.UCHC ~ meanEstimate.CTRL, data=DA_glm_blocked2.all_UCHC_CTRL)
lm_cduc.all_CTRL<-lm(meanTE.CDUC ~ meanEstimate.CTRL, data=DA_glm_blocked2.all_CDUC_CTRL)
sum_lm_ibdhc.all_CTRL<-summary(lm_ibdhc.all_CTRL)
sum_lm_cdhc.all_CTRL<-summary(lm_cdhc.all_CTRL)
sum_lm_uchc.all_CTRL<-summary(lm_uchc.all_CTRL)
sum_lm_cduc.all_CTRL<-summary(lm_cduc.all_CTRL)

df_lm_blocked2<-data.frame(
  subset=c("all","all","all","all","IBD","IBD","IBD","IBD","control","control","control","control"),
  DA_comparison=c("IBD vs. control","CD vs. UC","UC vs. control","CD vs. control","IBD vs. control","CD vs. UC","UC vs. control","CD vs. control","IBD vs. control","CD vs. UC","UC vs. control","CD vs. control"),
  p=c(
    sum_lm_ibdhc.all$coefficients[2,4],
    sum_lm_cdhc.all$coefficients[2,4],
    sum_lm_uchc.all$coefficients[2,4],
    sum_lm_cduc.all$coefficients[2,4],
    sum_lm_ibdhc.all_IBD$coefficients[2,4],
    sum_lm_cdhc.all_IBD$coefficients[2,4],
    sum_lm_uchc.all_IBD$coefficients[2,4],
    sum_lm_cduc.all_IBD$coefficients[2,4],
    sum_lm_ibdhc.all_CTRL$coefficients[2,4],
    sum_lm_cdhc.all_CTRL$coefficients[2,4],
    sum_lm_uchc.all_CTRL$coefficients[2,4],
    sum_lm_cduc.all_CTRL$coefficients[2,4]
  ),
  estimate=c(
    sum_lm_ibdhc.all$coefficients[2,1],
    sum_lm_cdhc.all$coefficients[2,1],
    sum_lm_uchc.all$coefficients[2,1],
    sum_lm_cduc.all$coefficients[2,1],
    sum_lm_ibdhc.all_IBD$coefficients[2,1],
    sum_lm_cdhc.all_IBD$coefficients[2,1],
    sum_lm_uchc.all_IBD$coefficients[2,1],
    sum_lm_cduc.all_IBD$coefficients[2,1],
    sum_lm_ibdhc.all_CTRL$coefficients[2,1],
    sum_lm_cdhc.all_CTRL$coefficients[2,1],
    sum_lm_uchc.all_CTRL$coefficients[2,1],
    sum_lm_cduc.all_CTRL$coefficients[2,1]
  )
)
df_lm_blocked2
```

```{r}
DA_glm_blocked2.all$IBDHC<-as.factor(DA_glm_blocked2.all$IBDHC)
DA_glm_blocked2.all$CDHC<-as.factor(DA_glm_blocked2.all$CDHC)
DA_glm_blocked2.all$UCHC<-as.factor(DA_glm_blocked2.all$UCHC)

coin::wilcox_test(abs(meanEstimate.all)~IBDHC, data=DA_glm_blocked2.all%>%filter(IBDHC!="non-sig"))
coin::wilcox_test(abs(meanEstimate.all)~CDHC, data=DA_glm_blocked2.all%>%filter(CDHC!="non-sig"))
coin::wilcox_test(abs(meanEstimate.all)~UCHC, data=DA_glm_blocked2.all%>%filter(UCHC!="non-sig"))
```

###

```{r}
# save table with all info
cor_spearblocked_DA<-l_glm_blocked2.resamp.all%>%
  full_join(da_all.sig, by="MGS")

cor_spearblocked_DA_sigs<-cor_spearblocked_DA%>%
  mutate(sigcorr=ifelse(spear_padjust<qval,"significant","notSignificant"))%>%
  filter(sigcorr=="significant")%>%  
  mutate(rho2=rho)%>%
  gather(measure,effect,c(rho,meanTE.IBDHC,meanTE.CDHC,meanTE.UCHC,meanTE.CDUC))%>%
  arrange(MGS, rho2)

all(names(df_cor_cross_DAibd_all_sigs%>%mutate(sample="cross"))==names(cor_long_DAibd_all_sigs.df%>%select(-num_sigcorr)))
df_cor_all_DAibd_all_sigs<-rbind(df_cor_cross_DAibd_all_sigs%>%mutate(sample="cross"),cor_long_DAibd_all_sigs.df%>%select(-num_sigcorr))
length(unique(df_cor_all_DAibd_all_sigs$MGS)) 
write.csv(df_cor_all_DAibd_all_sigs, paste0(PATHanalNov2023,"Downstream/Results/MGScorvsDAibd.crosslong.q0.1.csv"))
saveRDS(df_cor_all_DAibd_all_sigs, paste0(PATHanalNov2023,"humanReads/MGScorvsDAibd.crosslong.q0.1.rds"))


testdf<-df_cor_all_DAibd_all_sigs%>%
  filter(robustDA=="corr")%>%
  group_by(MGS,cohort)%>%
  count()

# how many MGS more than 1 cohort
testdf%>%group_by(MGS)%>%count()%>%filter(n>1)%>%nrow()

# direction
testdf2<-df_cor_all_DAibd_all_sigs%>%
  filter(robustDA=="corr")%>%
  mutate(direction=ifelse(rho2<0,"negative","positive"))%>%
  select(MGS,direction)%>%
  group_by(MGS,direction)%>%
  count()
# positive
testdf2%>%filter(direction=="positive")%>%nrow()
# negative
testdf2%>%filter(direction=="negative")%>%nrow()
# which MGS have discrepancy
testdf2%>%group_by(MGS)%>%count()%>%filter(n>1)
```

## IBD only

### crosssectional

```{r}
# count how many times MGS significant across cohorts
# select all that are significant in 1 or more cohorts

l_cor_cross_adjust_IBD<-l_cor_cross_IBD
df_cor_cross_IBD<-do.call(rbind, l_cor_cross_adjust_IBD)
df_cor_cross_DAibd_IBD<-df_cor_cross_IBD%>%
  full_join(ibd_ctrl.sig, by="MGS")

numSigCorr_IBD<-df_cor_cross_DAibd_IBD%>%
  mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
  group_by(MGS) %>%
  summarise(num_sigcorr = sum(sigcorr == "significant"))%>%
  left_join(df_cor_cross_DAibd_IBD, by="MGS")

cross_corsig_IBD<-numSigCorr_IBD%>%filter(num_sigcorr>0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
cross_corsig_pos_IBD<-numSigCorr_IBD%>%filter(num_sigcorr>0)%>%filter(padjust<qval)%>%filter(rho>0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
cross_corsig_neg_IBD<-numSigCorr_IBD%>%filter(num_sigcorr>0)%>%filter(padjust<qval)%>%filter(rho<0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS

cross_corsig_robDA_IBD<-numSigCorr_IBD%>%filter(num_sigcorr>0 & robustDA=="robust")%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and robust DA 
cross_corsig_varDA_IBD<-numSigCorr_IBD%>%filter(num_sigcorr>0 & robustDA=="variable")%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and variable DA
cross_corsig_notDA_IBD<-numSigCorr_IBD%>%filter(num_sigcorr>0 & is.na(robustDA))%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and not DA

cross_notcorr_IBD<-numSigCorr_IBD%>%filter(num_sigcorr==0)%>%select(MGS)%>%unique()%>%nrow() # nonsignificant corr MGS
cross_allmgs_IBD<-numSigCorr_IBD%>%select(MGS)%>%unique()%>%nrow() # all MGS

cross_compare_CORvsDA_IBD<-data.frame(
  Cor=cross_corsig_IBD,
  Cor_robDA=cross_corsig_robDA_IBD,
  Cor_varDA=cross_corsig_varDA_IBD,
  Cor_notDA=cross_corsig_notDA_IBD,
  NoCor=cross_notcorr_IBD,
  numMGS=cross_allmgs_IBD
)
cross_compare_CORvsDA_IBD

df_cor_cross_DAibd_sigs_IBD<-df_cor_cross_DAibd_IBD%>%
  mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
  filter(sigcorr=="significant")%>%  
  mutate(rho2=rho)%>%
  gather(measure,effect,c(rho,meanTE))%>%
  arrange(MGS, rho2)
write.csv(df_cor_cross_DAibd_sigs_IBD, paste0(PATHanalNov2023,"Downstream/Results/MGScorvsDAibd_IBD.q0.1.csv"))


df_cor_cross_DAibd_sigs2_IBD<-numSigCorr_IBD%>%
  mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
  filter(sigcorr=="significant")%>%  
  filter(num_sigcorr>1)%>%  
  mutate(rho2=rho)%>%
  gather(measure,effect,c(rho,meanTE))%>%
  arrange(MGS, rho2)

# save table with all info
df_cor_cross_DAibd_all_IBD<-df_cor_cross_DAibd_IBD%>%
  full_join(da_all.sig, by="MGS")
df_cor_cross_DAibd_all_sigs_IBD<-df_cor_cross_DAibd_all_IBD%>%
  mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
  filter(sigcorr=="significant")%>%  
  mutate(rho2=rho)%>%
  gather(measure,effect,c(rho,meanTE.IBDHC,meanTE.CDHC,meanTE.UCHC,meanTE.CDUC))%>%
  arrange(MGS, rho2)
```


```{r fig.width=8, fig.height=7}
df_cor_cross_DAibd_all_sigs_IBD<-df_cor_cross_DAibd_all_sigs_IBD%>%
   mutate(meanTE=ifelse(measure=="rho",NA,effect))%>%
   mutate(effect=ifelse(IBDHC=="non-sig" & measure=="meanTE.IBDHC",NA,effect))%>%
   mutate(effect=ifelse(CDHC=="non-sig" & measure=="meanTE.CDHC",NA,effect))%>%
   mutate(effect=ifelse(UCHC=="non-sig" & measure=="meanTE.UCHC",NA,effect))%>%
   mutate(effect=ifelse(CDUC=="non-sig" & measure=="meanTE.CDUC",NA,effect))%>%
   mutate(robustDA=ifelse(measure=="meanTE.IBDHC",IBDHC,NA))%>%
   mutate(robustDA=ifelse(measure=="meanTE.CDHC",CDHC,robustDA))%>%
   mutate(robustDA=ifelse(measure=="meanTE.UCHC",UCHC,robustDA))%>%
   mutate(robustDA=ifelse(measure=="meanTE.CDUC",CDUC,robustDA))%>%
   mutate(robustDA=ifelse(is.na(robustDA),"corr",robustDA))

df_cor_cross_DAibd_all_sigs_IBD$effect<-as.numeric(df_cor_cross_DAibd_all_sigs_IBD$effect)
class(df_cor_cross_DAibd_all_sigs_IBD$effect)

p_cor_cross_DAibd_all_sigs_IBD<-ggplot(df_cor_cross_DAibd_all_sigs_IBD%>%
   mutate(measure=ifelse(measure=="meanTE.IBDHC","meanTE IBD - control",measure))%>%
   mutate(measure=ifelse(measure=="meanTE.CDHC","meanTE CD - control",measure))%>%
   mutate(measure=ifelse(measure=="meanTE.UCHC","meanTE UC - control",measure))%>%
   mutate(measure=ifelse(measure=="meanTE.CDUC","meanTE CD - UC",measure)),
   aes(effect,reorder(gsub(".*;","",MGS),rho2),color=cohort,shape=robustDA))+
  geom_point(size=3, alpha=0.7) +
  facet_grid(.~measure, scales = "free_x") +
  ylab("MGS") +
  theme_bw()
p_cor_cross_DAibd_all_sigs_IBD
```


### longitudinal

```{r}
#corDA_df_long$Sample2

l_cor_long_adjust_IBD<-lapply(names(l_cor_long.resamp_IBD), function(s){
    l_cor_long_adjtmp<-lapply(names(l_cor_long.resamp_IBD[[s]]), function(x){
    corstmp<-l_cor_long.resamp_IBD[[s]][[x]]
    #corstmp$padjust<-p.adjust(corstmp$pvalue, method = "BH") # already adjusted before
    return(corstmp)
  })
  names(l_cor_long_adjtmp)<-names(l_cor_long.resamp_IBD[[s]])
  df_cor_long_samp<-do.call(rbind, l_cor_long_adjtmp)
  return(df_cor_long_samp)
})
names(l_cor_long_adjust_IBD)<-names(l_cor_long.resamp_IBD)

siglist<-list(ibd_ctrl.sig,cd_ctrl.sig,uc_ctrl.sig,cd_uc.sig)
names(siglist)<-c("ibd_ctrl","cd_ctrl","uc_ctrl","cd_uc")
  
corDA_df_long_IBD<-lapply(names(l_cor_long_adjust_IBD), function(x){
  df_cor_long<-l_cor_long_adjust_IBD[[x]]
  corDA_plot_long_samp<-lapply(siglist, function(s){
  
    df.filt<-df_cor_long%>%
      full_join(s, by="MGS")%>%
      filter(padjust<qval)
    
  
    return(df.filt)
  })
  names(corDA_plot_long_samp)<-names(siglist)
  return(corDA_plot_long_samp)
})
names(corDA_df_long_IBD)<-names(l_cor_long_adjust_IBD)
corDA_df_long_IBD$Sample1

# count how many times MGS significant across cohorts
# select all that are significant in 1 or more cohorts
# list longitudinal
numSigCorr.l_long_IBD<-lapply(corDA_df_long_IBD, function (x){
  numSigCorr<-x[["ibd_ctrl"]]%>%
    mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
    group_by(MGS) %>%
    summarise(num_sigcorr = sum(sigcorr == "significant"))%>%
    left_join(x[["ibd_ctrl"]], by="MGS")

  numSigCorr
})
names(numSigCorr.l_long_IBD)<-names(corDA_df_long_IBD)
#numSigCorr.l_long$Sample2

long_compare_CORvsDA.l_long_IBD<-lapply(numSigCorr.l_long_IBD, function (x){
  #x<-numSigCorr.l_long[[a]]
  long_corsig<-x%>%filter(num_sigcorr>0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
  long_corsig_pos<-x%>%filter(num_sigcorr>0)%>%filter(padjust<qval)%>%filter(rho>0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
  long_corsig_neg<-x%>%filter(num_sigcorr>0)%>%filter(padjust<qval)%>%filter(rho<0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
  
  long_corsig_robDA<-x%>%filter(num_sigcorr>0 & robustDA=="robust")%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and robust DA 
  long_corsig_varDA<-x%>%filter(num_sigcorr>0 & robustDA=="variable")%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and variable DA
  long_corsig_notDA<-x%>%filter(num_sigcorr>0 & is.na(robustDA))%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and not DA
  
  long_notcorr<-x%>%filter(num_sigcorr==0)%>%select(MGS)%>%unique()%>%nrow() # nonsignificant corr MGS
  long_allmgs<-x%>%select(MGS)%>%unique()%>%nrow() # all MGS
  
  long_compare_CORvsDA<-data.frame(
    Cor=long_corsig,
    Cor_robDA=long_corsig_robDA,
    Cor_varDA=long_corsig_varDA,
    Cor_notDA=long_corsig_notDA,
    NoCor=long_notcorr,
    numMGS=long_allmgs
  )
  long_compare_CORvsDA
  })
names(long_compare_CORvsDA.l_long_IBD)<-names(numSigCorr.l_long_IBD)
long_compare_CORvsDA.df_IBD<-do.call(rbind,long_compare_CORvsDA.l_long_IBD)
long_compare_CORvsDA.df_IBD

df_cor_long_DAibd_sigs_l_long_IBD<-lapply(numSigCorr.l_long_IBD, function (x){
  df_cor_long_DAibd_sigs<-x%>%
  mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
  filter(sigcorr=="significant")%>%  
  mutate(rho2=rho)%>%
  gather(measure,effect,c(rho,meanTE))%>%
  arrange(MGS, rho2)
  
  df_cor_long_DAibd_sigs
#write.csv(df_cor_long_DAibd_sigs, paste0(PATHanalNov2023,"Downstream/Results/MGScorvsDAibd.csv"))
})
names(df_cor_long_DAibd_sigs_l_long_IBD)<-names(numSigCorr.l_long_IBD)

df_cor_long_DAibd_sigs2_l_long_IBD<-lapply(numSigCorr.l_long_IBD, function (x){
  df_cor_long_DAibd_sigs2<-x%>%
    mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
    filter(sigcorr=="significant")%>%  
    filter(num_sigcorr>1)%>%  
    mutate(rho2=rho)%>%
    gather(measure,effect,c(rho,meanTE))%>%
    arrange(MGS, rho2)
  
  df_cor_long_DAibd_sigs2
})
names(df_cor_long_DAibd_sigs2_l_long_IBD)<-names(numSigCorr.l_long_IBD)

# save table with all info
df_cor_long_DAibd_all_l_long_IBD<-lapply(numSigCorr.l_long_IBD, function (x){
  df_cor_long_DAibd_all<-x%>%
    full_join(da_all.sig, by="MGS")
  df_cor_long_DAibd_all_sigs<-df_cor_long_DAibd_all%>%
    mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
    filter(sigcorr=="significant")%>%  
    mutate(rho2=rho)%>%
    gather(measure,effect,c(rho,meanTE.IBDHC,meanTE.CDHC,meanTE.UCHC,meanTE.CDUC))%>%
    arrange(MGS, rho2)
  
  df_cor_long_DAibd_all_sigs
})
names(df_cor_long_DAibd_all_l_long_IBD)<-names(numSigCorr.l_long_IBD)

```

```{r fig.width=8, fig.height=2}
#x<-df_cor_long_DAibd_all_l_long$Sample6

p_cor_long_DAibd_all_sigs.list_IBD<-lapply(df_cor_long_DAibd_all_l_long_IBD, function (x){
  x.df<-x%>%
    mutate(meanTE=ifelse(measure=="rho",NA,effect))%>%
    mutate(effect=ifelse(IBDHC=="non-sig" & measure=="meanTE.IBDHC",NA,effect))%>%
    mutate(effect=ifelse(CDHC=="non-sig" & measure=="meanTE.CDHC",NA,effect))%>%
    mutate(effect=ifelse(UCHC=="non-sig" & measure=="meanTE.UCHC",NA,effect))%>%
    mutate(effect=ifelse(CDUC=="non-sig" & measure=="meanTE.CDUC",NA,effect))%>%
    mutate(robustDA=ifelse(measure=="meanTE.IBDHC",IBDHC,NA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.CDHC",CDHC,robustDA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.UCHC",UCHC,robustDA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.CDUC",CDUC,robustDA))%>%
    mutate(robustDA=ifelse(is.na(robustDA),"corr",robustDA))
  
  x.df$effect<-as.numeric(x.df$effect)
  class(x.df$effect)
  
  p_cor_long_DAibd_all_sigs<-ggplot(x.df,aes(effect,reorder(gsub(".*;","",MGS),rho2),color=cohort,shape=robustDA))+
    geom_point(size=3, alpha=0.7) +
    facet_grid(.~measure, scales = "free_x") +
    ylab("MGS") +
    theme_bw()
  p_cor_long_DAibd_all_sigs
})

#p_cor_long_DAibd_all_sigs.list_IBD$Sample1

```

```{r fig.width=8, fig.height=7}
#x<-df_cor_long_DAibd_all_l_long$Sample6

cor_long_DAibd_all_sigs.list_IBD<-lapply(names(df_cor_long_DAibd_all_l_long_IBD), function (x){
  x.df<-df_cor_long_DAibd_all_l_long[[x]]%>%
    mutate(meanTE=ifelse(measure=="rho",NA,effect))%>%
    mutate(effect=ifelse(IBDHC=="non-sig" & measure=="meanTE.IBDHC",NA,effect))%>%
    mutate(effect=ifelse(CDHC=="non-sig" & measure=="meanTE.CDHC",NA,effect))%>%
    mutate(effect=ifelse(UCHC=="non-sig" & measure=="meanTE.UCHC",NA,effect))%>%
    mutate(effect=ifelse(CDUC=="non-sig" & measure=="meanTE.CDUC",NA,effect))%>%
    mutate(robustDA=ifelse(measure=="meanTE.IBDHC",IBDHC,NA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.CDHC",CDHC,robustDA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.UCHC",UCHC,robustDA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.CDUC",CDUC,robustDA))%>%
    mutate(robustDA=ifelse(is.na(robustDA),"corr",robustDA))%>%
    mutate(sample=x)
  
  x.df$effect<-as.numeric(x.df$effect)
  x.df
})
names(cor_long_DAibd_all_sigs.list_IBD)<-names(df_cor_long_DAibd_all_l_long_IBD)

cor_long_DAibd_all_sigs.df_IBD<-do.call(rbind,cor_long_DAibd_all_sigs.list_IBD)


p_cor_long_DAibd_all_sigs<-ggplot(cor_long_DAibd_all_sigs.df,aes(effect,reorder(gsub(".*;","",MGS),rho2),color=cohort,shape=robustDA))+
    geom_point(size=3, alpha=0.7) +
    facet_grid(.~measure, scales = "free_x") +
    ylab("MGS") +
    theme_bw()
p_cor_long_DAibd_all_sigs
```

```{r}
all(names(df_cor_cross_DAibd_all_sigs_IBD%>%mutate(sample="cross"))==names(cor_long_DAibd_all_sigs.df_IBD%>%select(-num_sigcorr)))
df_cor_all_DAibd_all_sigs_IBD<-rbind(df_cor_cross_DAibd_all_sigs_IBD%>%mutate(sample="cross"),cor_long_DAibd_all_sigs.df_IBD%>%select(-num_sigcorr))
length(unique(df_cor_all_DAibd_all_sigs_IBD$MGS)) 
write.csv(df_cor_all_DAibd_all_sigs_IBD, paste0(PATHanalNov2023,"Downstream/Results/MGScorvsDAibd.crosslong.IBDonly.q0.1.csv"))
saveRDS(df_cor_all_DAibd_all_sigs_IBD, paste0(PATHanalNov2023,"humanReads/MGScorvsDAibd.crosslong.IBDonly.q0.1.rds"))


testdf_IBD<-df_cor_all_DAibd_all_sigs_IBD%>%
  filter(robustDA=="corr")%>%
  group_by(MGS,cohort)%>%
  count()

# how many MGS more than 1 cohort
testdf_IBD%>%group_by(MGS)%>%count()%>%filter(n>1)%>%nrow()

# direction
testdf2_IBD<-df_cor_all_DAibd_all_sigs_IBD%>%
  filter(robustDA=="corr")%>%
  mutate(direction=ifelse(rho2<0,"negative","positive"))%>%
  select(MGS,direction)%>%
  group_by(MGS,direction)%>%
  count()
# positive
testdf2_IBD%>%filter(direction=="positive")%>%nrow()
# negative
testdf2_IBD%>%filter(direction=="negative")%>%nrow()
# which MGS have discrepancy
testdf2_IBD%>%group_by(MGS)%>%count()%>%filter(n>1)
```

```{r fig.width=8, fig.height=15}
p_cor_all_DAibd_all_sigs_IBD<-ggplot(df_cor_all_DAibd_all_sigs_IBD%>%
   mutate(measure=ifelse(measure=="meanTE.IBDHC","meanTE IBD - control",measure))%>%
   mutate(measure=ifelse(measure=="meanTE.CDHC","meanTE CD - control",measure))%>%
   mutate(measure=ifelse(measure=="meanTE.UCHC","meanTE UC - control",measure))%>%
   mutate(measure=ifelse(measure=="meanTE.CDUC","meanTE CD - UC",measure)),
   aes(effect,reorder(gsub(".*;","",gsub(";MGS."," MGS.",MGS)),rho2),color=cohort,shape=robustDA))+
    geom_point(size=3, alpha=0.7) +
    facet_grid(.~measure, scales = "free_x") +
    ylab("MGS") +
    theme_bw()
p_cor_all_DAibd_all_sigs_IBD
```


```{r fig.width=7, fig.height=6}
df_cor_all_DAibd_rhoVSmeanTE_IBD<-df_cor_all_DAibd_all_sigs_IBD%>%
  filter(!is.na(effect))%>%
  select(c(MGS,cohort,measure,effect,sample))%>%
  spread(measure,effect)%>%
  gather(comparison,meanTE,c("meanTE.CDUC","meanTE.IBDHC","meanTE.UCHC","meanTE.CDHC"))%>%
  mutate(comparison=gsub("meanTE.","",comparison))%>%
  mutate(comparison=gsub("HC"," vs. control",comparison))%>%
  mutate(comparison=gsub("CDUC","CD vs. UC",comparison))

(df_cor_all_DAibd_rhoVSmeanTE_IBD%>%filter(!is.na(meanTE)))$MGS%>%unique()%>%length()

p2_cor_all_DAibd_all_sigs_IBD<-ggplot(df_cor_all_DAibd_rhoVSmeanTE_IBD,aes(rho,meanTE))+
    geom_smooth(method="lm", se = FALSE, color="grey70", size=1)+
    geom_point(size=1, alpha=0.7, aes(color=cohort)) +
    facet_wrap(.~comparison, scales = "free_x") +
    ylab("Mean differential abundance effect size (mean TE)") +
    theme_bw()
p2_cor_all_DAibd_all_sigs_IBD

```



```{r}
df_cor_all_DAibd_rhoVSmeanTE_IBD$comparison%>%unique()


cduc_dfcorr_IBD<-df_cor_all_DAibd_rhoVSmeanTE_IBD%>%filter(comparison=="CD vs. UC")
cdhc_dfcorr_IBD<-df_cor_all_DAibd_rhoVSmeanTE_IBD%>%filter(comparison=="CD vs. control")
uchc_dfcorr_IBD<-df_cor_all_DAibd_rhoVSmeanTE_IBD%>%filter(comparison=="UC vs. control")
ibd_dfcorr_IBD<-df_cor_all_DAibd_rhoVSmeanTE_IBD%>%filter(comparison=="IBD vs. control")

lm_cduc_IBD<-lmerTest::lmer(meanTE ~ rho + (1|cohort), data=cduc_dfcorr_IBD)
lm_cdhc_IBD<-lmerTest::lmer(meanTE ~ rho + (1|cohort), data=cdhc_dfcorr_IBD)
lm_uchc_IBD<-lmerTest::lmer(meanTE ~ rho + (1|cohort), data=uchc_dfcorr_IBD)
lm_ibd_IBD<-lmerTest::lmer(meanTE ~ rho + (1|cohort), data=ibd_dfcorr_IBD)

sum_lm_cduc_IBD<-summary(lm_cduc_IBD)
sum_lm_cdhc_IBD<-summary(lm_cdhc_IBD)
sum_lm_uchc_IBD<-summary(lm_uchc_IBD)
sum_lm_ibd_IBD<-summary(lm_ibd_IBD)

lm_meanTE_rho_IBD<-data.frame(
  comparison=c("CD vs. UC","IBD vs. control","UC vs. control","CD vs. control"),
  p=c(sum_lm_cduc_IBD$coefficients[2,5],sum_lm_ibd_IBD$coefficients[2,5],sum_lm_uchc_IBD$coefficients[2,5],sum_lm_cdhc_IBD$coefficients[2,5]),
  rho=c(sum_lm_cduc_IBD$coefficients[2,1],sum_lm_ibd_IBD$coefficients[2,1],sum_lm_uchc_IBD$coefficients[2,1],sum_lm_cdhc_IBD$coefficients[2,1])
)
lm_meanTE_rho_IBD$padjust<-p.adjust(lm_meanTE_rho_IBD$p)
lm_meanTE_rho_IBD



cor_cduc_IBD<-cor.test((df_cor_all_DAibd_rhoVSmeanTE_IBD%>%filter(comparison=="CD vs. UC"))$meanTE, (df_cor_all_DAibd_rhoVSmeanTE_IBD%>%filter(comparison=="CD vs. UC"))$rho, use = "complete.obs", method="spearman")

cor_ibd_IBD<-cor.test((df_cor_all_DAibd_rhoVSmeanTE_IBD%>%filter(comparison=="IBD vs. control"))$meanTE, (df_cor_all_DAibd_rhoVSmeanTE_IBD%>%filter(comparison=="IBD vs. control"))$rho, use = "complete.obs", method="spearman")

cor_uchc_IBD<-cor.test((df_cor_all_DAibd_rhoVSmeanTE_IBD%>%filter(comparison=="UC vs. control"))$meanTE, (df_cor_all_DAibd_rhoVSmeanTE_IBD%>%filter(comparison=="UC vs. control"))$rho, use = "complete.obs", method="spearman")

cor_cdhc_IBD<-cor.test((df_cor_all_DAibd_rhoVSmeanTE_IBD%>%filter(comparison=="CD vs. control"))$meanTE, (df_cor_all_DAibd_rhoVSmeanTE_IBD%>%filter(comparison=="CD vs. control"))$rho, use = "complete.obs", method="spearman")

cor_meanTE_rho_IBD<-data.frame(
  comparison=c("CD vs. UC","IBD vs. control","UC vs. control","CD vs. control"),
  p=c(cor_cduc_IBD$p.value,cor_ibd_IBD$p.value,cor_uchc_IBD$p.value,cor_cdhc_IBD$p.value),
  rho=c(cor_cduc_IBD$estimate,cor_ibd_IBD$estimate,cor_uchc_IBD$estimate,cor_cdhc_IBD$es)
)
cor_meanTE_rho_IBD$padjust<-p.adjust(cor_meanTE_rho_IBD$p)
cor_meanTE_rho_IBD
```


```{r}
df_cor_all_DAibd_rhoVSmeanTE2_IBD<-df_cor_all_DAibd_all_sigs_IBD%>%
  filter(!is.na(effect))%>%
  select(c(MGS,cohort,measure,effect,sample,robustDA))

var_rob.cdhc_IBD<-df_cor_all_DAibd_all_sigs_IBD%>%
  filter(!is.na(effect))%>%
  filter(measure%in%c("rho","meanTE.CDHC"))%>%
  filter(CDHC%in%c("variable","robust"))%>%
  select(c(MGS,cohort,measure,effect,sample,CDHC))%>%
  spread(measure,effect)%>%
  rename("meanTE"="meanTE.CDHC")%>%
  rename("robustDA"="CDHC")

ggplot(var_rob.cdhc_IBD, aes(x=robustDA, y=abs(rho)))+
  geom_boxplot()+
  geom_jitter()

var_rob.cdhc_IBD$cohort<-as.factor(var_rob.cdhc_IBD$cohort)
var_rob.cdhc_IBD$robustDA<-as.factor(var_rob.cdhc_IBD$robustDA)
#coin::wilcox_test(abs(rho)~robustDA | cohort, var_rob.cdhc_IBD)

var_rob.uchc_IBD<-df_cor_all_DAibd_all_sigs_IBD%>%
  filter(!is.na(effect))%>%
  filter(measure%in%c("rho","meanTE.UCHC"))%>%
  filter(UCHC%in%c("variable","robust"))%>%
  select(c(MGS,cohort,measure,effect,sample,UCHC))%>%
  spread(measure,effect)%>%
  rename("meanTE"="meanTE.UCHC")%>%
  rename("robustDA"="UCHC")

ggplot(var_rob.uchc_IBD, aes(x=robustDA, y=abs(rho)))+
  geom_boxplot()+
  geom_jitter()

var_rob.uchc_IBD$cohort<-as.factor(var_rob.uchc_IBD$cohort)
var_rob.uchc_IBD$robustDA<-as.factor(var_rob.uchc_IBD$robustDA)
coin::wilcox_test(abs(rho)~robustDA | cohort, var_rob.uchc_IBD)


```

## controls only

### crosssectional

```{r}
# count how many times MGS significant across cohorts
# select all that are significant in 1 or more cohorts

l_cor_cross_adjust_CTRL<-l_cor_cross_CTRL
df_cor_cross_CTRL<-do.call(rbind, l_cor_cross_adjust_CTRL)
df_cor_cross_DAibd_CTRL<-df_cor_cross_CTRL%>%
  full_join(ibd_ctrl.sig, by="MGS")

numSigCorr_CTRL<-df_cor_cross_DAibd_CTRL%>%
  mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
  group_by(MGS) %>%
  summarise(num_sigcorr = sum(sigcorr == "significant"))%>%
  left_join(df_cor_cross_DAibd_CTRL, by="MGS")

cross_corsig_CTRL<-numSigCorr_CTRL%>%filter(num_sigcorr>0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
cross_corsig_pos_CTRL<-numSigCorr_CTRL%>%filter(num_sigcorr>0)%>%filter(padjust<qval)%>%filter(rho>0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
cross_corsig_neg_CTRL<-numSigCorr_CTRL%>%filter(num_sigcorr>0)%>%filter(padjust<qval)%>%filter(rho<0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS

cross_corsig_robDA_CTRL<-numSigCorr_CTRL%>%filter(num_sigcorr>0 & robustDA=="robust")%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and robust DA 
cross_corsig_varDA_CTRL<-numSigCorr_CTRL%>%filter(num_sigcorr>0 & robustDA=="variable")%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and variable DA
cross_corsig_notDA_CTRL<-numSigCorr_CTRL%>%filter(num_sigcorr>0 & is.na(robustDA))%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and not DA

cross_notcorr_CTRL<-numSigCorr_CTRL%>%filter(num_sigcorr==0)%>%select(MGS)%>%unique()%>%nrow() # nonsignificant corr MGS
cross_allmgs_CTRL<-numSigCorr_CTRL%>%select(MGS)%>%unique()%>%nrow() # all MGS

cross_compare_CORvsDA_CTRL<-data.frame(
  Cor=cross_corsig_CTRL,
  Cor_robDA=cross_corsig_robDA_CTRL,
  Cor_varDA=cross_corsig_varDA_CTRL,
  Cor_notDA=cross_corsig_notDA_CTRL,
  NoCor=cross_notcorr_CTRL,
  numMGS=cross_allmgs_CTRL
)
cross_compare_CORvsDA_CTRL

df_cor_cross_DAibd_sigs_CTRL<-df_cor_cross_DAibd_CTRL%>%
  mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
  filter(sigcorr=="significant")%>%  
  mutate(rho2=rho)%>%
  gather(measure,effect,c(rho,meanTE))%>%
  arrange(MGS, rho2)
write.csv(df_cor_cross_DAibd_sigs_CTRL, paste0(PATHanalNov2023,"Downstream/Results/MGScorvsDAibd_CTRL.q0.1.csv"))


df_cor_cross_DAibd_sigs2_CTRL<-numSigCorr_CTRL%>%
  mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
  filter(sigcorr=="significant")%>%  
  filter(num_sigcorr>1)%>%  
  mutate(rho2=rho)%>%
  gather(measure,effect,c(rho,meanTE))%>%
  arrange(MGS, rho2)

# save table with all info
df_cor_cross_DAibd_all_CTRL<-df_cor_cross_DAibd_CTRL%>%
  full_join(da_all.sig, by="MGS")
df_cor_cross_DAibd_all_sigs_CTRL<-df_cor_cross_DAibd_all_CTRL%>%
  mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
  filter(sigcorr=="significant")%>%  
  mutate(rho2=rho)%>%
  gather(measure,effect,c(rho,meanTE.IBDHC,meanTE.CDHC,meanTE.UCHC,meanTE.CDUC))%>%
  arrange(MGS, rho2)
```


```{r fig.width=8, fig.height=7}
df_cor_cross_DAibd_all_sigs_CTRL<-df_cor_cross_DAibd_all_sigs_CTRL%>%
   mutate(meanTE=ifelse(measure=="rho",NA,effect))%>%
   mutate(effect=ifelse(IBDHC=="non-sig" & measure=="meanTE.IBDHC",NA,effect))%>%
   mutate(effect=ifelse(CDHC=="non-sig" & measure=="meanTE.CDHC",NA,effect))%>%
   mutate(effect=ifelse(UCHC=="non-sig" & measure=="meanTE.UCHC",NA,effect))%>%
   mutate(effect=ifelse(CDUC=="non-sig" & measure=="meanTE.CDUC",NA,effect))%>%
   mutate(robustDA=ifelse(measure=="meanTE.IBDHC",IBDHC,NA))%>%
   mutate(robustDA=ifelse(measure=="meanTE.CDHC",CDHC,robustDA))%>%
   mutate(robustDA=ifelse(measure=="meanTE.UCHC",UCHC,robustDA))%>%
   mutate(robustDA=ifelse(measure=="meanTE.CDUC",CDUC,robustDA))%>%
   mutate(robustDA=ifelse(is.na(robustDA),"corr",robustDA))

df_cor_cross_DAibd_all_sigs_CTRL$effect<-as.numeric(df_cor_cross_DAibd_all_sigs_CTRL$effect)
class(df_cor_cross_DAibd_all_sigs_CTRL$effect)

p_cor_cross_DAibd_all_sigs_CTRL<-ggplot(df_cor_cross_DAibd_all_sigs_CTRL%>%
   mutate(measure=ifelse(measure=="meanTE.IBDHC","meanTE IBD - control",measure))%>%
   mutate(measure=ifelse(measure=="meanTE.CDHC","meanTE CD - control",measure))%>%
   mutate(measure=ifelse(measure=="meanTE.UCHC","meanTE UC - control",measure))%>%
   mutate(measure=ifelse(measure=="meanTE.CDUC","meanTE CD - UC",measure)),
   aes(effect,reorder(gsub(".*;","",MGS),rho2),color=cohort,shape=robustDA))+
  geom_point(size=3, alpha=0.7) +
  facet_grid(.~measure, scales = "free_x") +
  ylab("MGS") +
  theme_bw()
p_cor_cross_DAibd_all_sigs_CTRL
```


### longitudinal

```{r}
#corDA_df_long$Sample2

l_cor_long_adjust_CTRL<-lapply(names(l_cor_long.resamp_CTRL), function(s){
    l_cor_long_adjtmp<-lapply(names(l_cor_long.resamp_CTRL[[s]]), function(x){
    corstmp<-l_cor_long.resamp_CTRL[[s]][[x]]
    #corstmp$padjust<-p.adjust(corstmp$pvalue, method = "BH") # already adjusted before
    return(corstmp)
  })
  names(l_cor_long_adjtmp)<-names(l_cor_long.resamp_CTRL[[s]])
  df_cor_long_samp<-do.call(rbind, l_cor_long_adjtmp)
  return(df_cor_long_samp)
})
names(l_cor_long_adjust_CTRL)<-names(l_cor_long.resamp_CTRL)

siglist<-list(ibd_ctrl.sig,cd_ctrl.sig,uc_ctrl.sig,cd_uc.sig)
names(siglist)<-c("ibd_ctrl","cd_ctrl","uc_ctrl","cd_uc")
  
corDA_df_long_CTRL<-lapply(names(l_cor_long_adjust_CTRL), function(x){
  df_cor_long<-l_cor_long_adjust_CTRL[[x]]
  corDA_plot_long_samp<-lapply(siglist, function(s){
  
    df.filt<-df_cor_long%>%
      full_join(s, by="MGS")%>%
      filter(padjust<qval)
    
  
    return(df.filt)
  })
  names(corDA_plot_long_samp)<-names(siglist)
  return(corDA_plot_long_samp)
})
names(corDA_df_long_CTRL)<-names(l_cor_long_adjust_CTRL)
corDA_df_long_CTRL$Sample1

# count how many times MGS significant across cohorts
# select all that are significant in 1 or more cohorts
# list longitudinal
numSigCorr.l_long_CTRL<-lapply(corDA_df_long_CTRL, function (x){
  numSigCorr<-x[["ibd_ctrl"]]%>%
    mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
    group_by(MGS) %>%
    summarise(num_sigcorr = sum(sigcorr == "significant"))%>%
    left_join(x[["ibd_ctrl"]], by="MGS")

  numSigCorr
})
names(numSigCorr.l_long_CTRL)<-names(corDA_df_long_CTRL)
#numSigCorr.l_long$Sample2

long_compare_CORvsDA.l_long_CTRL<-lapply(numSigCorr.l_long_CTRL, function (x){
  #x<-numSigCorr.l_long[[a]]
  long_corsig<-x%>%filter(num_sigcorr>0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
  long_corsig_pos<-x%>%filter(num_sigcorr>0)%>%filter(padjust<qval)%>%filter(rho>0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
  long_corsig_neg<-x%>%filter(num_sigcorr>0)%>%filter(padjust<qval)%>%filter(rho<0)%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS
  
  long_corsig_robDA<-x%>%filter(num_sigcorr>0 & robustDA=="robust")%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and robust DA 
  long_corsig_varDA<-x%>%filter(num_sigcorr>0 & robustDA=="variable")%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and variable DA
  long_corsig_notDA<-x%>%filter(num_sigcorr>0 & is.na(robustDA))%>%select(MGS)%>%unique()%>%nrow() # significant corr MGS and not DA
  
  long_notcorr<-x%>%filter(num_sigcorr==0)%>%select(MGS)%>%unique()%>%nrow() # nonsignificant corr MGS
  long_allmgs<-x%>%select(MGS)%>%unique()%>%nrow() # all MGS
  
  long_compare_CORvsDA<-data.frame(
    Cor=long_corsig,
    Cor_robDA=long_corsig_robDA,
    Cor_varDA=long_corsig_varDA,
    Cor_notDA=long_corsig_notDA,
    NoCor=long_notcorr,
    numMGS=long_allmgs
  )
  long_compare_CORvsDA
  })
names(long_compare_CORvsDA.l_long_CTRL)<-names(numSigCorr.l_long_CTRL)
long_compare_CORvsDA.df_CTRL<-do.call(rbind,long_compare_CORvsDA.l_long_CTRL)
long_compare_CORvsDA.df_CTRL

df_cor_long_DAibd_sigs_l_long_CTRL<-lapply(numSigCorr.l_long_CTRL, function (x){
  df_cor_long_DAibd_sigs<-x%>%
  mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
  filter(sigcorr=="significant")%>%  
  mutate(rho2=rho)%>%
  gather(measure,effect,c(rho,meanTE))%>%
  arrange(MGS, rho2)
  
  df_cor_long_DAibd_sigs
#write.csv(df_cor_long_DAibd_sigs, paste0(PATHanalNov2023,"Downstream/Results/MGScorvsDAibd.csv"))
})
names(df_cor_long_DAibd_sigs_l_long_CTRL)<-names(numSigCorr.l_long_CTRL)

df_cor_long_DAibd_sigs2_l_long_CTRL<-lapply(numSigCorr.l_long_CTRL, function (x){
  df_cor_long_DAibd_sigs2<-x%>%
    mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
    filter(sigcorr=="significant")%>%  
    filter(num_sigcorr>1)%>%  
    mutate(rho2=rho)%>%
    gather(measure,effect,c(rho,meanTE))%>%
    arrange(MGS, rho2)
  
  df_cor_long_DAibd_sigs2
})
names(df_cor_long_DAibd_sigs2_l_long_CTRL)<-names(numSigCorr.l_long_CTRL)

# save table with all info
df_cor_long_DAibd_all_l_long_CTRL<-lapply(numSigCorr.l_long_CTRL, function (x){
  df_cor_long_DAibd_all<-x%>%
    full_join(da_all.sig, by="MGS")
  df_cor_long_DAibd_all_sigs<-df_cor_long_DAibd_all%>%
    mutate(sigcorr=ifelse(padjust<qval,"significant","notSignificant"))%>%
    filter(sigcorr=="significant")%>%  
    mutate(rho2=rho)%>%
    gather(measure,effect,c(rho,meanTE.IBDHC,meanTE.CDHC,meanTE.UCHC,meanTE.CDUC))%>%
    arrange(MGS, rho2)
  
  df_cor_long_DAibd_all_sigs
})
names(df_cor_long_DAibd_all_l_long_CTRL)<-names(numSigCorr.l_long_CTRL)

```

```{r fig.width=8, fig.height=2}
#x<-df_cor_long_DAibd_all_l_long$Sample6

p_cor_long_DAibd_all_sigs.list_CTRL<-lapply(df_cor_long_DAibd_all_l_long_CTRL, function (x){
  x.df<-x%>%
    mutate(meanTE=ifelse(measure=="rho",NA,effect))%>%
    mutate(effect=ifelse(IBDHC=="non-sig" & measure=="meanTE.IBDHC",NA,effect))%>%
    mutate(effect=ifelse(CDHC=="non-sig" & measure=="meanTE.CDHC",NA,effect))%>%
    mutate(effect=ifelse(UCHC=="non-sig" & measure=="meanTE.UCHC",NA,effect))%>%
    mutate(effect=ifelse(CDUC=="non-sig" & measure=="meanTE.CDUC",NA,effect))%>%
    mutate(robustDA=ifelse(measure=="meanTE.IBDHC",IBDHC,NA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.CDHC",CDHC,robustDA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.UCHC",UCHC,robustDA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.CDUC",CDUC,robustDA))%>%
    mutate(robustDA=ifelse(is.na(robustDA),"corr",robustDA))
  
  x.df$effect<-as.numeric(x.df$effect)
  class(x.df$effect)
  
  p_cor_long_DAibd_all_sigs<-ggplot(x.df,aes(effect,reorder(gsub(".*;","",MGS),rho2),color=cohort,shape=robustDA))+
    geom_point(size=3, alpha=0.7) +
    facet_grid(.~measure, scales = "free_x") +
    ylab("MGS") +
    theme_bw()
  p_cor_long_DAibd_all_sigs
})

#p_cor_long_DAibd_all_sigs.list_CTRL$Sample1

```

```{r fig.width=8, fig.height=7}
#x<-df_cor_long_DAibd_all_l_long$Sample6

cor_long_DAibd_all_sigs.list_CTRL<-lapply(names(df_cor_long_DAibd_all_l_long_CTRL), function (x){
  x.df<-df_cor_long_DAibd_all_l_long_CTRL[[x]]%>%
    mutate(meanTE=ifelse(measure=="rho",NA,effect))%>%
    mutate(effect=ifelse(IBDHC=="non-sig" & measure=="meanTE.IBDHC",NA,effect))%>%
    mutate(effect=ifelse(CDHC=="non-sig" & measure=="meanTE.CDHC",NA,effect))%>%
    mutate(effect=ifelse(UCHC=="non-sig" & measure=="meanTE.UCHC",NA,effect))%>%
    mutate(effect=ifelse(CDUC=="non-sig" & measure=="meanTE.CDUC",NA,effect))%>%
    mutate(robustDA=ifelse(measure=="meanTE.IBDHC",IBDHC,NA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.CDHC",CDHC,robustDA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.UCHC",UCHC,robustDA))%>%
    mutate(robustDA=ifelse(measure=="meanTE.CDUC",CDUC,robustDA))%>%
    mutate(robustDA=ifelse(is.na(robustDA),"corr",robustDA))%>%
    mutate(sample=x)
  
  x.df$effect<-as.numeric(x.df$effect)
  x.df
})
names(cor_long_DAibd_all_sigs.list_CTRL)<-names(df_cor_long_DAibd_all_l_long_CTRL)

cor_long_DAibd_all_sigs.df_CTRL<-do.call(rbind,cor_long_DAibd_all_sigs.list_CTRL)


p_cor_long_DAibd_all_sigs_CTRL<-ggplot(cor_long_DAibd_all_sigs.df_CTRL,aes(effect,reorder(gsub(".*;","",MGS),rho2),color=cohort,shape=robustDA))+
    geom_point(size=3, alpha=0.7) +
    facet_grid(.~measure, scales = "free_x") +
    ylab("MGS") +
    theme_bw()
p_cor_long_DAibd_all_sigs_CTRL
```

```{r}
all(names(df_cor_cross_DAibd_all_sigs_CTRL%>%mutate(sample="cross"))==names(cor_long_DAibd_all_sigs.df_CTRL%>%select(-num_sigcorr)))
df_cor_all_DAibd_all_sigs_CTRL<-rbind(df_cor_cross_DAibd_all_sigs_CTRL%>%mutate(sample="cross"),cor_long_DAibd_all_sigs.df_CTRL%>%select(-num_sigcorr))
length(unique(df_cor_all_DAibd_all_sigs_CTRL$MGS)) 
write.csv(df_cor_all_DAibd_all_sigs_CTRL, paste0(PATHanalNov2023,"Downstream/Results/MGScorvsDAibd.crosslong.CTRLonly.q0.1.csv"))
saveRDS(df_cor_all_DAibd_all_sigs_CTRL, paste0(PATHanalNov2023,"humanReads/MGScorvsDAibd.crosslong.CTRLonly.q0.1.rds"))


testdf_CTRL<-df_cor_all_DAibd_all_sigs_CTRL%>%
  filter(robustDA=="corr")%>%
  group_by(MGS,cohort)%>%
  count()

# how many MGS more than 1 cohort
testdf_CTRL%>%group_by(MGS)%>%count()%>%filter(n>1)%>%nrow()

# direction
testdf2_CTRL<-df_cor_all_DAibd_all_sigs_CTRL%>%
  filter(robustDA=="corr")%>%
  mutate(direction=ifelse(rho2<0,"negative","positive"))%>%
  select(MGS,direction)%>%
  group_by(MGS,direction)%>%
  count()
# positive
testdf2_CTRL%>%filter(direction=="positive")%>%nrow()
# negative
testdf2_CTRL%>%filter(direction=="negative")%>%nrow()
# which MGS have discrepancy
testdf2_CTRL%>%group_by(MGS)%>%count()%>%filter(n>1)
```

```{r fig.width=10, fig.height=5}
p_cor_all_DAibd_all_sigs_CTRL<-ggplot(df_cor_all_DAibd_all_sigs_CTRL%>%
   mutate(measure=ifelse(measure=="meanTE.IBDHC","meanTE IBD - control",measure))%>%
   mutate(measure=ifelse(measure=="meanTE.CDHC","meanTE CD - control",measure))%>%
   mutate(measure=ifelse(measure=="meanTE.UCHC","meanTE UC - control",measure))%>%
   mutate(measure=ifelse(measure=="meanTE.CDUC","meanTE CD - UC",measure)),
   aes(effect,reorder(gsub(".*;","",gsub(";MGS."," MGS.",MGS)),rho2),color=cohort,shape=robustDA))+
    geom_point(size=3, alpha=0.7) +
    facet_grid(.~measure, scales = "free_x") +
    ylab("MGS") +
    theme_bw()
p_cor_all_DAibd_all_sigs_CTRL
```


```{r fig.width=7, fig.height=6}
df_cor_all_DAibd_rhoVSmeanTE_CTRL<-df_cor_all_DAibd_all_sigs_CTRL%>%
  filter(!is.na(effect))%>%
  select(c(MGS,cohort,measure,effect,sample))%>%
  spread(measure,effect)%>%
  gather(comparison,meanTE,c("meanTE.CDUC","meanTE.IBDHC","meanTE.UCHC","meanTE.CDHC"))%>%
  mutate(comparison=gsub("meanTE.","",comparison))%>%
  mutate(comparison=gsub("HC"," vs. control",comparison))%>%
  mutate(comparison=gsub("CDUC","CD vs. UC",comparison))

(df_cor_all_DAibd_rhoVSmeanTE_CTRL%>%filter(!is.na(meanTE)))$MGS%>%unique()%>%length()

p2_cor_all_DAibd_all_sigs_CTRL<-ggplot(df_cor_all_DAibd_rhoVSmeanTE_CTRL,aes(rho,meanTE))+
    geom_smooth(method="lm", se = FALSE, color="grey70", size=1)+
    geom_point(size=1, alpha=0.7, aes(color=cohort)) +
    facet_wrap(.~comparison, scales = "free_x") +
    ylab("Mean differential abundance effect size (mean TE)") +
    theme_bw()
p2_cor_all_DAibd_all_sigs_CTRL

```


```{r}
df_cor_all_DAibd_rhoVSmeanTE_CTRL$comparison%>%unique()


cduc_dfcorr_CTRL<-df_cor_all_DAibd_rhoVSmeanTE_CTRL%>%filter(comparison=="CD vs. UC")
cdhc_dfcorr_CTRL<-df_cor_all_DAibd_rhoVSmeanTE_CTRL%>%filter(comparison=="CD vs. control")
uchc_dfcorr_CTRL<-df_cor_all_DAibd_rhoVSmeanTE_CTRL%>%filter(comparison=="UC vs. control")
ibd_dfcorr_CTRL<-df_cor_all_DAibd_rhoVSmeanTE_CTRL%>%filter(comparison=="IBD vs. control")

lm_cduc_CTRL<-lmerTest::lmer(meanTE ~ rho + (1|cohort), data=cduc_dfcorr_CTRL)
lm_cdhc_CTRL<-lmerTest::lmer(meanTE ~ rho + (1|cohort), data=cdhc_dfcorr_CTRL)
lm_uchc_CTRL<-lmerTest::lmer(meanTE ~ rho + (1|cohort), data=uchc_dfcorr_CTRL)
lm_CTRL_CTRL<-lmerTest::lmer(meanTE ~ rho + (1|cohort), data=ibd_dfcorr_CTRL)

sum_lm_cduc_CTRL<-summary(lm_cduc_CTRL)
sum_lm_cdhc_CTRL<-summary(lm_cdhc_CTRL)
sum_lm_uchc_CTRL<-summary(lm_uchc_CTRL)
sum_lm_CTRL_CTRL<-summary(lm_CTRL_CTRL)

lm_meanTE_rho_CTRL<-data.frame(
  comparison=c("CD vs. UC","IBD vs. control","UC vs. control","CD vs. control"),
  p=c(sum_lm_cduc_CTRL$coefficients[2,5],sum_lm_CTRL_CTRL$coefficients[2,5],sum_lm_uchc_CTRL$coefficients[2,5],sum_lm_cdhc_CTRL$coefficients[2,5]),
  rho=c(sum_lm_cduc_CTRL$coefficients[2,1],sum_lm_CTRL_CTRL$coefficients[2,1],sum_lm_uchc_CTRL$coefficients[2,1],sum_lm_cdhc_CTRL$coefficients[2,1])
)
lm_meanTE_rho_CTRL$padjust<-p.adjust(lm_meanTE_rho_CTRL$p)
lm_meanTE_rho_CTRL

cor_cduc_CTRL<-cor.test((df_cor_all_DAibd_rhoVSmeanTE_CTRL%>%filter(comparison=="CD vs. UC"))$meanTE, (df_cor_all_DAibd_rhoVSmeanTE_CTRL%>%filter(comparison=="CD vs. UC"))$rho, use = "complete.obs", method="spearman")

cor_CTRL_CTRL<-cor.test((df_cor_all_DAibd_rhoVSmeanTE_CTRL%>%filter(comparison=="IBD vs. control"))$meanTE, (df_cor_all_DAibd_rhoVSmeanTE_CTRL%>%filter(comparison=="IBD vs. control"))$rho, use = "complete.obs", method="spearman")

cor_uchc_CTRL<-cor.test((df_cor_all_DAibd_rhoVSmeanTE_CTRL%>%filter(comparison=="UC vs. control"))$meanTE, (df_cor_all_DAibd_rhoVSmeanTE_CTRL%>%filter(comparison=="UC vs. control"))$rho, use = "complete.obs", method="spearman")

cor_cdhc_CTRL<-cor.test((df_cor_all_DAibd_rhoVSmeanTE_CTRL%>%filter(comparison=="CD vs. control"))$meanTE, (df_cor_all_DAibd_rhoVSmeanTE_CTRL%>%filter(comparison=="CD vs. control"))$rho, use = "complete.obs", method="spearman")

cor_meanTE_rho_CTRL<-data.frame(
  comparison=c("CD vs. UC","IBD vs. control","UC vs. control","CD vs. control"),
  p=c(cor_cduc_CTRL$p.value,cor_CTRL_CTRL$p.value,cor_uchc_CTRL$p.value,cor_cdhc_CTRL$p.value),
  rho=c(cor_cduc_CTRL$estimate,cor_CTRL_CTRL$estimate,cor_uchc_CTRL$estimate,cor_cdhc_CTRL$es)
)
cor_meanTE_rho_CTRL$padjust<-p.adjust(cor_meanTE_rho_CTRL$p)
cor_meanTE_rho_CTRL
```

## comapre IBD and CTRL rho

```{r}
df_cor_all_DAibd_all_sigs_IBD_mod<-df_cor_all_DAibd_all_sigs_IBD%>%
  mutate(group="IBD")
  
df_cor_all_DAibd_all_sigs_CTRL_mod<-df_cor_all_DAibd_all_sigs_CTRL%>%
  mutate(group="control")

df_cor_all_DAibd_all_sigs_IBD_CTRL<-rbind(df_cor_all_DAibd_all_sigs_IBD_mod,df_cor_all_DAibd_all_sigs_CTRL_mod)

df_cor_all_DAibd_all_sigs_IBD_mod2<-df_cor_all_DAibd_all_sigs_IBD_mod%>%
  filter(measure=="rho")%>%
  select(c(MGS, rho2, pvalue, padjust, enrichedIn))
df_cor_all_DAibd_all_sigs_CTRL_mod2<-df_cor_all_DAibd_all_sigs_CTRL_mod%>%
  filter(measure=="rho")%>%
  select(c(MGS, rho2, pvalue, padjust, enrichedIn))
df_cor_all_DAibd_all_sigs_IBD_CTRL_mod2<-df_cor_all_DAibd_all_sigs_IBD_mod2%>%
  full_join(df_cor_all_DAibd_all_sigs_IBD_mod2, by="MGS")

ggplot(df_cor_all_DAibd_all_sigs_IBD_CTRL_mod2, aes(x=rho2.x,y=rho2.y))+
  geom_point()
```


## correlation of Falks individual-blocked species

```{r}
MGScor_blInd<-read.table(paste0(PATHanalNov2023,"humanReads/ibd_ctrl_blockedInd_MGS.txt"), header=TRUE, check.names = FALSE, sep="\t")%>%
  mutate(MGS=`Feature name`)
MGScor_blInd$Direction<-as.numeric(MGScor_blInd$Direction)
MGScor_blInd$`q-value`<-as.numeric(MGScor_blInd$`q-value`)
MGScor_blInd_DA<-MGScor_blInd%>%
  left_join(da_all.sig, by="MGS")


MGScor_blInd_DA_gat<-MGScor_blInd_DA%>%
  filter(`q-value`<qval)%>%
  mutate(meanTE.CDHC=ifelse(CDHC=="non-sig",NA,meanTE.CDHC))%>%
  mutate(meanTE.UCHC=ifelse(UCHC=="non-sig",NA,meanTE.UCHC))%>%
  mutate(meanTE.CDUC=ifelse(CDUC=="non-sig",NA,meanTE.CDUC))%>%
  mutate(meanTE.IBDHC=ifelse(IBDHC=="non-sig",NA,meanTE.IBDHC))%>%
  gather(measure,effect,c(Direction,meanTE.IBDHC,meanTE.CDHC,meanTE.UCHC,meanTE.CDUC))%>%
  filter(!is.na(effect))%>%
  select(c(MGS,measure,effect))%>%
  spread(measure,effect)%>%
  gather(comparison,meanTE,c("meanTE.CDUC","meanTE.IBDHC","meanTE.UCHC","meanTE.CDHC"))%>%
  mutate(comparison=gsub("meanTE.","",comparison))%>%
  mutate(comparison=gsub("HC"," vs. control",comparison))%>%
  mutate(comparison=gsub("CDUC","CD vs. UC",comparison))
MGScor_blInd_DA_gat$Direction<-as.numeric(MGScor_blInd_DA_gat$Direction)
MGScor_blInd_DA_gat$meanTE<-as.numeric(MGScor_blInd_DA_gat$meanTE)

(MGScor_blInd_DA_gat%>%filter(!is.na(meanTE)))$MGS%>%unique()%>%length()

p2_cor_all_DAibd_blInd<-ggplot(MGScor_blInd_DA_gat,aes(Direction,meanTE))+
    geom_smooth(method="lm", se = FALSE, color="grey70", size=1)+
    geom_point(size=1, alpha=0.7) +
    facet_wrap(.~comparison, scales = "free_x") +
    ylab("Mean differential abundance effect size (mean TE)") +
    theme_bw()
p2_cor_all_DAibd_blInd
```

```{r}
cduc_dfcorr_blInd<-MGScor_blInd_DA_gat%>%filter(comparison=="CD vs. UC")
cdhc_dfcorr_blInd<-MGScor_blInd_DA_gat%>%filter(comparison=="CD vs. control")
uchc_dfcorr_blInd<-MGScor_blInd_DA_gat%>%filter(comparison=="UC vs. control")
ibd_dfcorr_blInd<-MGScor_blInd_DA_gat%>%filter(comparison=="IBD vs. control")

lm_cduc_blInd<-lm(meanTE ~ Direction, data=cduc_dfcorr_blInd)
lm_cdhc_blInd<-lm(meanTE ~ Direction, data=cdhc_dfcorr_blInd)
lm_uchc_blInd<-lm(meanTE ~ Direction, data=uchc_dfcorr_blInd)
lm_ibd_blInd<-lm(meanTE ~ Direction, data=ibd_dfcorr_blInd)

sum_lm_cduc_blInd<-summary(lm_cduc_blInd)
sum_lm_cdhc_blInd<-summary(lm_cdhc_blInd)
sum_lm_uchc_blInd<-summary(lm_uchc_blInd)
sum_lm_ibd_blInd<-summary(lm_ibd_blInd)

lm_meanTE_direction_blInd<-data.frame(
  comparison=c("CD vs. UC","IBD vs. control","UC vs. control","CD vs. control"),
  p=c(sum_lm_cduc_blInd$coefficients[2,4],sum_lm_ibd_blInd$coefficients[2,4],sum_lm_uchc_blInd$coefficients[2,4],sum_lm_cdhc_blInd$coefficients[2,4]),
  direction=c(sum_lm_cduc_blInd$coefficients[2,1],sum_lm_ibd_blInd$coefficients[2,1],sum_lm_uchc_blInd$coefficients[2,1],sum_lm_cdhc_blInd$coefficients[2,1])
)
lm_meanTE_direction_blInd$padjust<-p.adjust(lm_meanTE_direction_blInd$p)
lm_meanTE_direction_blInd
```

```{r}
saveRDS(df_cor_all_DAibd_rhoVSmeanTE,paste0(PATHanalNov2023,"humanReads/df_cor_all_DAibd_rhoVSmeanTE.q0.1.rds"))
saveRDS(df_cor_all_DAibd_rhoVSmeanTE_IBD,paste0(PATHanalNov2023,"humanReads/df_cor_all_DAibd_rhoVSmeanTE_IBD.q0.1.rds"))
saveRDS(df_cor_all_DAibd_rhoVSmeanTE_CTRL,paste0(PATHanalNov2023,"humanReads/df_cor_all_DAibd_rhoVSmeanTE_CTRL.q0.1.rds"))
saveRDS(MGScor_blInd_DA_gat,paste0(PATHanalNov2023,"humanReads/MGScor_blInd_DA_gat.q0.1.rds"))

p1_cor_all_DA<-ggplot(df_cor_all_DAibd_rhoVSmeanTE%>%filter(comparison=="IBD vs. control"),aes(rho,meanTE))+
    geom_smooth(method="lm", se = FALSE, color="grey70", size=1)+
    geom_point(size=1, alpha=0.7) +
    #facet_wrap(.~comparison, scales = "free_x") +
    ylab("Mean differential abundance effect size (mean TE)") +
    theme_bw()+
    scale_x_reverse()+
  ggtitle("All individuals")
p1_cor_all_DA

p2_cor_ibd_DA<-ggplot(df_cor_all_DAibd_rhoVSmeanTE_IBD%>%filter(comparison=="IBD vs. control"),aes(rho,meanTE))+
    geom_smooth(method="lm", se = FALSE, color="grey70", size=1)+
    geom_point(size=1, alpha=0.7) +
    #facet_wrap(.~comparison, scales = "free_x") +
    ylab("Mean differential abundance effect size (mean TE)") +
    theme_bw()+
    scale_x_reverse()+
  ggtitle("IBD individuals")
p2_cor_ibd_DA

p3_cor_ibd_DA<-ggplot(df_cor_all_DAibd_rhoVSmeanTE_CTRL%>%filter(comparison=="IBD vs. control"),aes(rho,meanTE))+
    geom_smooth(method="lm", se = FALSE, color="grey70", size=1)+
    geom_point(size=1, alpha=0.7) +
    #facet_wrap(.~comparison, scales = "free_x") +
    ylab("Mean differential abundance effect size (mean TE)") +
    theme_bw()+
    scale_x_reverse()+
  ggtitle("Control individuals")
p3_cor_ibd_DA

p4_cor_blInd_DA<-ggplot(MGScor_blInd_DA_gat%>%filter(comparison=="IBD vs. control"),aes(Direction,meanTE))+
    geom_smooth(method="lm", se = FALSE, color="grey70", size=1)+
    geom_point(size=1, alpha=0.7) +
    #facet_wrap(.~comparison, scales = "free_x") +
    ylab("Mean differential abundance effect size (mean TE)") +
    theme_bw()+
    scale_x_reverse()+
  ggtitle("Blocked for individuals")
p4_cor_blInd_DA

plot_grid(p1_cor_all_DA,p2_cor_ibd_DA,p3_cor_ibd_DA,p4_cor_blInd_DA,nrow=1)
```

```{r}
corr_log_blocked<-function(taxa_abun, metaVar, var){
  #taxa_abun <- spec.cross
  #metaVar <- meta_hr
  #var <- "HumanRds"
    taxab.df<-taxa_abun
    testmgs.vec<-names(taxab.df%>%select(-cohort))
    #testmgs.vec<-names(taxab.df)[1:5] # debug
    taxmeta.df<-taxab.df%>%
      rownames_to_column("SmplID")%>%
      left_join(metaVar, by="SmplID")%>%
      filter(!is.na(!!sym(var)))

    cor_mgs<-lapply(testmgs.vec, function(m) {
     #m<-"Bacteria;Firmicutes_C;Negativicutes;Veillonellales;Megasphaeraceae;Megasphaera;Megasphaera sp000417505;MGS.966" # debug
      #m<-"Bacteria;Firmicutes_A;Clostridia;Lachnospirales;Lachnospiraceae;Faecalimonas;Faecalimonas sp900546325;MGS.1523 cohort"
      #print(m)
      taxab.vec<-taxmeta.df[,m]
      hr.vec<-as.numeric(c(taxmeta.df$HumanRds))
      cohort<-taxmeta.df$cohort
      
      class(taxab.vec)
      class(hr.vec)
      cohort<-as.factor(cohort)
      class(cohort)

      sd_taxab.vec<-sd(taxab.vec, na.rm = TRUE)
      sd_hr.vec<-sd(hr.vec, na.rm = TRUE)


      if(sd_taxab.vec == 0 | sd_hr.vec == 0){
        return(NULL)
      } else {
        tryCatch(
        {
          #cor_taxab_hr<-cor.test(formula= ~ taxab.vec + hr.vec | cohort, use = "complete.obs", method="spearman")
          #cor_taxab_hr<-lmerTest::lmer(taxab.vec ~ hr.vec +(1|cohort))
          cor_taxab_hr<-lmerTest::lmer(log2(taxab.vec+1) ~ log2(hr.vec+1) +(1|cohort))
          cor_taxab_hr_summary<-summary(cor_taxab_hr)
          
          
          #return(cor_taxab_hr)
          # Extract correlation coefficient and p-value
          corr_est <- cor_taxab_hr_summary$coefficients[2,1]
          corr_p <- cor_taxab_hr_summary$coefficients[2,5]
        },
        error = function(e) {
          message("Error: ", e$message)
          #return(NA)
          corr_rho <- NA
          corr_p <- NA
        })
        cor_cohort_tmp<-data.frame(
          MGS=m,
          pvalue=corr_p,
          estimate=corr_est
        )
        return(cor_cohort_tmp)
      }
    })
    names(cor_mgs)<-testmgs.vec
    cor_mgs<-cor_mgs[sapply(cor_mgs, function(x) !is.null(x))]
    cor_mgs_df<-do.call(rbind, cor_mgs)
    return(cor_mgs_df)
}


# crosssectional
glm_spec_hr_blocked_log<-corr_log_blocked(spec.cross,meta_hr,"HumanRds")
glm_spec_hr_blocked_log$padjust<-p.adjust(glm_spec_hr_blocked_log$pvalue, method="fdr")

#longitudinal
# make list with dataframe subset according to resampling
l_spec.long.subs<-lapply(names(df.row[1:20]), function(s){
  subsamps<-df.row[[s]]
  df_spec.samp<-spec.long[subsamps,]
  return(df_spec.samp)
})
names(l_spec.long.subs)<-names(df.row[1:20])

l_glm_blocked_long.resamp<-lapply(names(l_spec.long.subs), function(s){
  corr_long<-corr_blocked(l_spec.long.subs[[s]],meta_hr,"HumanRds")
  corr_long$padjust<-p.adjust(corr_long$pvalue, method="BH")
  return(corr_long)
})
names(l_glm_blocked_long.resamp)<-names(l_spec.long.subs)
l_glm_blocked_long.resamp$Sample3%>%View()

l_glm_blocked_long.resamp$Sample1%>%filter(padjust<qval)%>%nrow()
```


