---
title: "Figure Strains"
author: "Rebecca Ansorge"
date: "31/03/2023"
output: html_document
---

# Make strain figure

## Load libraries
```{r}
library(tidyverse)
library(gridExtra)
library(formattable)
library(cowplot)
library(ggtree)
library(ggtreeExtra)
library(ggnewscale)
library(ape)
library(ggVennDiagram)
library(ggpubr)
```

## Import data
```{r}
#PATHanalNov2023<-"~/Documents/Rebecca/onedrive/analysisNov2023/"
#PATHanalNov2023<-"~/Library/CloudStorage/OneDrive-Personal/analysisNov2023/"
PATHanalNov2023<-"~/Documents/Rebecca/onedrive/analysisNov2023/"
PATH<-"~/Documents/Rebecca/onedrive/submission/natcom/revision/checks/"

```

```{r}
# define color palette

#col_IBD2<-c("#E8A400","#E31A1C","#FB9A99","#1F78B4","#1F78B4")
col_IBD2<-c("gold2","#E31A1C","darkorange","#1F78B4","#1F78B4")
names(col_IBD2)<-c("IBD","CD","UC","healthy","control")



#display.brewer.pal(n=12, name="Paired")
#brewer.pal(n=12, name="Paired")
```

```{r}
strainR2vsTE.incTips<-readRDS(paste0(PATHanalNov2023,"strains/RDS/strainR2vsTE.incTips_q0.1.June23.rds"))
strainR2vsTE.inclNonSig<-readRDS(paste0(PATHanalNov2023,"strains/RDS/strainR2vsTE.inclNonSig_q0.1.June23.rds"))
x<-readRDS(paste0(PATHanalNov2023,"strains/RDS/StrainClust.overlapIBD_IBD2.June23.rds"))
y<-readRDS(paste0(PATHanalNov2023,"strains/RDS/StrainClust.overlapIBD_IBD2.variable.June23.rds"))

# lm correlating strainR2 and DA after accounting for number of tips
strainR2vsDA.lm<-readRDS(paste0(PATHanalNov2023,"strains/RDS/strainR2vsDA.lm.June23_q0.1.rds"))

# import top MGS shown in DA fig
DAmeans.topMGS.tax<-readRDS(paste0(PATHanalNov2023,"Downstream/Results/MGS.toptax_DAFigq0.1.rds"))

```

```{r}
# count robust and variable MGS
strainR2vsTE.inclNonSig%>%
  group_by(robustness)%>%
  count()

strainR2vsTE.incTips%>%
  group_by(robustStrains)%>%
  count()

strainR2vsTE.incTips%>%
  filter(robustStrains=="robust")

#write.csv(strainR2vsTE.incTips,paste0(PATHanalNov2023,"Downstream/Results/StrainsDAcompare.csv"))

robustMGS<-(strainR2vsTE.incTips%>%filter(robustStrains=="robust"))$MGS
variableMGS<-(strainR2vsTE.incTips%>%filter(robustStrains=="variable"))$MGS

strainR2vsTE.incTips%>%
  ungroup()%>%
  count(robustStrains,robustDA)

strainR2vsTE.incTips%>%
  ungroup()%>%
  count(significance)
```


### compare robustly clustering strains IBD / IBD2

```{r}
names(x)<-c("IBD/control       ","       CD/UC/control")
venn.p1<-ggVennDiagram(x, label_alpha = 0, edge_size =0.5,set_size = 4,label_size=4)+
  scale_fill_gradient(low="white",high = "salmon")+
  ggtitle("Robust")+
  theme(legend.position = "none")+ 
  #scale_x_continuous(expand = expansion(mult = .4))+
  coord_flip()
```

### compare variable clustering strains IBD / IBD2

```{r}
names(y)<-c("IBD/control       ","       CD/UC/control")
venn.p2<-ggVennDiagram(y, label_alpha = 0, edge_size =0.5,set_size = 4,label_size=4)+
  scale_fill_gradient(low="white",high = "salmon")+
  ggtitle("Variable (longitudinal resampling)")+
  theme(legend.position = "none")+ 
  #scale_x_continuous(expand = expansion(mult = .4))+
  coord_flip()
```

# Make plots

## Plot correlation mean DA effect size vs strain mean R2 (MGS significant in each)
```{r}
strainR2vsDA.lm$anysigLog2$call
strainR2vsDA.lm$anysigLog2$coefficients
p.val.strainR2vsDA.lm<-strainR2vsDA.lm$anysigLog2$coefficients["abs(meanTE.random)","Pr(>|t|)"]

strainR2vsTE.incTips<-strainR2vsTE.incTips%>%
  mutate("Significant in"=gsub("sig","",significance))%>%
  mutate("Significant in"=gsub("DA","species DA",`Significant in`))

strainR2vsTE.p1<-ggplot(strainR2vsTE.incTips) +
    geom_errorbar(aes(x=log2(meanR2),y=abs(meanTE.random),xmin=log2(meanR2-stdevR2), xmax=log2(meanR2+stdevR2)), width=.1, color='grey')+
    geom_errorbar(aes(x=log2(meanR2),y=abs(meanTE.random),ymin=abs(meanTE.random)-stdevTE.random, ymax=abs(meanTE.random)+stdevTE.random), width=.1, color='grey')+
    geom_smooth(aes(x=log2(meanR2),y=abs(meanTE.random)),method='lm', color="black", linewidth=0.5, alpha=0.2) +
    geom_point(aes(x=log2(meanR2),y=abs(meanTE.random),color=`Significant in`), alpha=0.7, size=2) +
    #stat_cor(aes(x=log2(meanR2),y=abs(meanTE.random)),label.x = -10, label.y = 1,size=3) +
    geom_text(aes(x=-5,y=-1,label=paste("p=",round(p.val.strainR2vsDA.lm,3),"*","\nafter accounting for number of tips")), size=4) +
    theme_bw() +
    ggtitle("DA effect sizes VS R2 of permanova")+
    ylab("DA effect size") +
    xlab("log2(mean R2 of strain clustering IBD vs. healthy)") +
    theme(title = element_text(size=14),
          axis.text = element_text(size=14),
          axis.title = element_text(size=14),
        #legend.position = "top",
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        legend.key = element_rect(fill = "white", colour = NA),
        legend.key.size = unit(0.1, "cm"),
        legend.spacing.y = unit(0.1, "cm"),
        legend.position=c(.88,.9),
        legend.box.background = element_rect(colour = "black"),)+
  scale_color_brewer(palette="Set2")
strainR2vsTE.p1

```

## Plot correlation mean DA effect size vs strain mean R2 (MGS significant in both only)
```{r}
p.val.bothsig.strainR2vsDA.lm<-strainR2vsDA.lm$bothsigLog2$coefficients["abs(meanTE.random)","Pr(>|t|)"]

strainR2vsTE.p2<-ggplot(strainR2vsTE.incTips%>%filter(significance=="both sig")) +
    geom_errorbar(aes(x=log2(meanR2),y=abs(meanTE.random),xmin=log2(meanR2-stdevR2), xmax=log2(meanR2+stdevR2)), width=.1, color='grey')+
    geom_errorbar(aes(x=log2(meanR2),y=abs(meanTE.random),ymin=abs(meanTE.random)-stdevTE.random, ymax=abs(meanTE.random)+stdevTE.random), width=.1, color='grey')+
    geom_smooth(aes(x=log2(meanR2),y=abs(meanTE.random)),method='lm', color="black", size=0.5, alpha=0.2) +
    geom_point(aes(x=log2(meanR2),y=abs(meanTE.random)), alpha=0.7) +
    #stat_cor(aes(x=log2(meanR2),y=abs(meanTE.random)),label.x = -10, label.y = 1,size=3) +
    geom_text(aes(x=-8,y=1.5,label=paste("p-value:",round(p.val.bothsig.strainR2vsDA.lm,3),"*","\nafter accounting for number of tips")), size=3) +
    theme_bw() +
    ggtitle("DA effect sizes VS R2 of permanova (MGS sig in both) ")+
    ylab("DA effect size") +
    xlab("log2(mean R2 of strain clustering IBD vs. healthy)") +
    theme(title = element_text(size=10),
        #legend.position = "top",
        legend.title = element_text(size = 10), 
        legend.text = element_text(size = 10),
        legend.key = element_rect(fill = "white", colour = NA),
        legend.key.size = unit(0.1, "cm"),
        legend.spacing.y = unit(0.1, "cm"))
strainR2vsTE.p2


```

## Make table comparing strains and DA

```{r}
strainR2vsTE.inclNonSig.counts<-strainR2vsTE.inclNonSig%>%
  group_by(robustness)%>%
  count()
strainR2vsTE.inclNonSig.counts

both_nonsig<-(strainR2vsTE.inclNonSig.counts%>%filter(robustness=="strains_non-sig DA_non-sig"))$n
STRns_DArob<-(strainR2vsTE.inclNonSig.counts%>%filter(robustness=="strains_non-sig DA_robust"))$n
STRns_DAvar<-(strainR2vsTE.inclNonSig.counts%>%filter(robustness=="strains_non-sig DA_variable"))$n
STRrob_DAns<-(strainR2vsTE.inclNonSig.counts%>%filter(robustness=="strains_robust DA_non-sig"))$n
bothrob<-(strainR2vsTE.inclNonSig.counts%>%filter(robustness=="strains_robust DA_robust"))$n
STRrob_DAvar<-(strainR2vsTE.inclNonSig.counts%>%filter(robustness=="strains_robust DA_variable"))$n
STRvar_DAns<-(strainR2vsTE.inclNonSig.counts%>%filter(robustness=="strains_variable DA_non-sig"))$n
STRvar_DArob<-(strainR2vsTE.inclNonSig.counts%>%filter(robustness=="strains_variable DA_robust"))$n
STRvar_DAvar<-(strainR2vsTE.inclNonSig.counts%>%filter(robustness=="strains_variable DA_variable"))$n


confmat<-tibble("# MGS"=c("robust strains","variable strains","non-sig. strains"),"robust DA"=c(bothrob,STRvar_DArob,STRns_DArob),"variable DA"=c(STRrob_DAvar,STRvar_DAvar,STRns_DAvar),"non-sig. DA"=c(STRrob_DAns,STRvar_DAns,both_nonsig))

confmat.obj<-formattable(confmat,
            list(`#` = formatter("span", style = ~ style(font.weight = "bold"))))
confmat.obj
class(confmat.obj)
confmat.obj.tab<-tableGrob(confmat.obj,rows=c("","",""))

confmat.obj.tab2<-ggtexttable(confmat.obj, rows = NULL, theme = ttheme("light", base_size=13, ))
confmat.obj.tab2
```

## Make table comparing strains (CD/) and DA


```{r}
strainR2vsTE.inclNonSig.ibd_ibd2<-readRDS(paste0(PATHanalNov2023,"Downstream/Results/overview.MGS.DA_q0.1_strains_cor_genes_PosNegSeparate.rds"))
strainR2vsTE.inclNonSig.ibd_ibd2.counts<-strainR2vsTE.inclNonSig.ibd_ibd2%>%
  mutate(robustness=paste0("strains_",robustStrains_CDUCCTRL," DA_",robustDA))%>%
  group_by(robustness)%>%
  count()
strainR2vsTE.inclNonSig.ibd_ibd2.counts

both_nonsig_ibd2<-(strainR2vsTE.inclNonSig.ibd_ibd2.counts%>%filter(robustness=="strains_non-sig DA_non-sig"))$n
STRns_DArob_ibd2<-(strainR2vsTE.inclNonSig.ibd_ibd2.counts%>%filter(robustness=="strains_non-sig DA_robust"))$n
STRns_DAvar_ibd2<-(strainR2vsTE.inclNonSig.ibd_ibd2.counts%>%filter(robustness=="strains_non-sig DA_variable"))$n
STRrob_DAns_ibd2<-(strainR2vsTE.inclNonSig.ibd_ibd2.counts%>%filter(robustness=="strains_robust DA_non-sig"))$n
bothrob_ibd2<-(strainR2vsTE.inclNonSig.ibd_ibd2.counts%>%filter(robustness=="strains_robust DA_robust"))$n
STRrob_DAvar_ibd2<-(strainR2vsTE.inclNonSig.ibd_ibd2.counts%>%filter(robustness=="strains_robust DA_variable"))$n
STRvar_DAns_ibd2<-(strainR2vsTE.inclNonSig.ibd_ibd2.counts%>%filter(robustness=="strains_variable DA_non-sig"))$n
STRvar_DArob_ibd2<-(strainR2vsTE.inclNonSig.ibd_ibd2.counts%>%filter(robustness=="strains_variable DA_robust"))$n
STRvar_DAvar_ibd2<-(strainR2vsTE.inclNonSig.ibd_ibd2.counts%>%filter(robustness=="strains_variable DA_variable"))$n


confmat_ibd2<-tibble("# MGS"=c("robust strains","variable strains","non-sig. strains"),"robust DA"=c(bothrob_ibd2,STRvar_DArob_ibd2,STRns_DArob_ibd2),"variable DA"=c(STRrob_DAvar_ibd2,STRvar_DAvar_ibd2,STRns_DAvar_ibd2),"non-sig. DA"=c(STRrob_DAns_ibd2,STRvar_DAns_ibd2,both_nonsig_ibd2))

confmat.obj_ibd2<-formattable(confmat_ibd2,
            list(`#` = formatter("span", style = ~ style(font.weight = "bold"))))
confmat.obj_ibd2
class(confmat.obj_ibd2)
confmat.obj_ibd2.tab<-tableGrob(confmat.obj_ibd2,rows=c("","",""))

confmat.obj_ibd2.tab2<-ggtexttable(confmat.obj_ibd2, rows = NULL, theme = ttheme("light", base_size=13, ))
confmat.obj_ibd2.tab2
```


```{r}
strainvsall.subset<-strainR2vsTE.inclNonSig.ibd_ibd2%>%
  filter(robustStrains_IBD!="non-sig" | robustStrains_CDUCCTRL!="non-sig")

strainvsall.subset.mod<-strainvsall.subset%>%
  mutate(meanR2_IBD=ifelse(robustStrains_IBD=="non-sig",NA,meanR2_IBD))%>% 
  mutate(stdevR2_IBD=ifelse(robustStrains_IBD=="non-sig",NA,stdevR2_IBD))%>%
  mutate(meanR2_CDUCCTRL=ifelse(robustStrains_CDUCCTRL=="non-sig",NA,meanR2_CDUCCTRL))%>%
  mutate(stdevR2_CDUCCTRL=ifelse(robustStrains_CDUCCTRL=="non-sig",NA,stdevR2_CDUCCTRL))%>%
  mutate(n_sig_DA=ifelse(enrichedIn=="IBD",n_sig_DA*-1,n_sig_DA))%>%
  mutate(n_sig_DA.cdhc=ifelse(enrichedIn.cdhc=="CD",n_sig_DA.cdhc*-1,n_sig_DA.cdhc))%>%
  mutate(n_sig_DA.uchc=ifelse(enrichedIn.uchc=="UC",n_sig_DA.uchc*-1,n_sig_DA.uchc))%>%
  mutate(n_obs=ifelse(humanReadCorrelation=="negative",n_obs*-1,n_obs))%>%
  mutate(n_chrt=ifelse(humanReadCorrelation=="negative",n_chrt*-1,n_chrt))%>%
  mutate(MGS=gsub("$",")",gsub(".*;","",gsub(";MGS"," (MGS",bins))))


strainvsall.subset.mod$DAfig <-strainvsall.subset.mod$MGS %in%DAmeans.topMGS.tax$SpeciesMGS
strainvsall.subset.mod<-strainvsall.subset.mod%>%
  mutate(highlight=ifelse(DAfig==TRUE, "bold","plain"))
strainvsall.subset.mod2<-strainvsall.subset.mod%>%
  select(-c(humanReadCorrelation,n_chrt,n_obs))%>%
  distinct(.keep_all=TRUE)

strainvsall.subset.mod_gat<-strainvsall.subset.mod%>%
  gather(comparison, robustStrains, c(robustStrains_IBD,robustStrains_CDUCCTRL))%>%
  filter(robustStrains!="nonsig")%>%
  gather(group,meanR2,c(meanR2_IBD, meanR2_CDUCCTRL))

p1_R2strains<-ggplot(strainvsall.subset.mod)+
  geom_point(aes(x=meanR2_IBD,y=reorder(MGS,meanR2_IBD), color=robustStrains_IBD, shape="IBD-ctrl"),size=4)+ #shape=19,
  geom_errorbar(aes(x=meanR2_IBD,y=reorder(MGS,meanR2_IBD),xmin=meanR2_IBD-stdevR2_IBD, xmax=meanR2_IBD+stdevR2_IBD), width=.1, color='grey')+
  geom_point(aes(x=meanR2_CDUCCTRL,y=reorder(MGS,meanR2_IBD), color=robustStrains_CDUCCTRL, shape="CD-UC-ctrl"), size=4)+ #shape=1,
  geom_errorbar(aes(x=meanR2_CDUCCTRL,y=reorder(MGS,meanR2_IBD),xmin=meanR2_CDUCCTRL-stdevR2_CDUCCTRL, xmax=meanR2_CDUCCTRL+stdevR2_CDUCCTRL), width=.1, color='grey')+
  theme_bw()+
  theme(legend.position="bottom",
        legend.title=element_blank(),
        legend.text = element_text(size=14),
        legend.direction = "vertical",
        title = element_text(size=14),
        axis.ticks.y = element_blank())+
  theme(
    axis.text.y = element_text(size=14, face = (strainvsall.subset.mod2[order(strainvsall.subset.mod2$meanR2_IBD),])$highlight),
        #axis.text.y = element_text(size=14, face = highlight),
        axis.text.x = element_text(size=14),
        axis.title = element_text(size=14),
        legend.margin = margin(0,0,0,0, unit="cm"))+
  scale_color_manual(values=c("royalblue","lightblue","red"),breaks=c("robust", "variable"))+
  scale_shape_manual(values=c("IBD-ctrl" = 19, "CD-UC-ctrl" = 1))+
  theme(legend.box="vertical")+
  ylab("All species with significant strain signal\n\n")
      
p2_nsigstrains<-ggplot(strainvsall.subset.mod%>%
                         gather(set,n_sig,c(n_sig_IBD,n_sig_CDUCCTRL))%>%
                         mutate(set=ifelse(set=="n_sig_IBD","IBD-ctrl","CD-UC-ctrl")))+
  geom_bar(aes(x=n_sig,y=reorder(MGS,meanR2_IBD), fill=set), stat="identity",position=position_dodge())+
  theme_bw()+
  theme(legend.position="bottom",
        axis.text.y=element_blank(),
        axis.title.y=element_blank(),
        legend.title=element_blank(),
        legend.text = element_text(size=14),
        legend.direction = "vertical",
        title = element_text(size=12),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size=14))+
  scale_fill_manual(values = c("skyblue","gold2"), na.value = "white")+
  xlab("# sig. runs")


p3_meanTE_DA<-ggplot(strainvsall.subset.mod%>%
    mutate(meanTE.cdhc=ifelse(robustDA.cdhc=="non-sig",NA,meanTE.cdhc))%>%
    mutate(meanTE.uchc=ifelse(robustDA.uchc=="non-sig",NA,meanTE.uchc))%>%
    #mutate(meanTE.cduc=ifelse(robustDA.cduc=="non-sig",NA,meanTE.cduc))%>%
    gather(comparison,meanTE,c(meanTE.cdhc,meanTE.uchc))%>%
    mutate(comparison=gsub("meanTE.cdhc","CD",comparison))%>%
    mutate(comparison=gsub("meanTE.uchc","UC",comparison)))+
  geom_point(aes(x=meanTE,y=reorder(MGS,meanR2_IBD), color=comparison),size=3, alpha=0.7)+ #shape=19,
  geom_vline(xintercept = 0, linetype=2, color="grey60")+
  theme_bw()+
  theme(legend.position="bottom",
        axis.text.y=element_blank(),
        axis.title.y=element_blank(),
        legend.title=element_blank(),
        legend.text = element_text(size=14),
        legend.direction = "vertical",
        title = element_text(size=12),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size=14))+
  scale_color_manual(values=col_IBD2)+
  #scale_shape_manual(values=c("IBD-ctrl" = 19, "CD-UC-ctrl" = 1))+
  theme(legend.box="vertical")#+
  #ylab("mean TE")


corMGS_rho<-read.csv(paste0(PATHanalNov2023,"Downstream/Results/MGScorvsDAibd.q0.1.crosslong.csv"))%>%
  filter(measure=="rho")%>%
  mutate(MGS=gsub("$",")",gsub(".*;","",gsub(";MGS"," (MGS",MGS))))
corMGS_rho<-corMGS_rho%>%
  group_by(MGS)%>%
  summarize(meanRho=mean(rho2))%>%
  left_join(corMGS_rho, by="MGS")%>%
  distinct(MGS, comparison,.keep_all = TRUE)

p4_corrHR<-ggplot(strainvsall.subset.mod%>%
    left_join(corMGS_rho,by="MGS"))+
  geom_point(aes(x=meanRho,y=reorder(MGS,meanR2_IBD)),size=3, alpha=0.7)+ #shape=19,
  geom_vline(xintercept = 0, linetype=2, color="grey60")+
  theme_bw()+
  theme(legend.position="bottom",
        axis.text.y=element_blank(),
        axis.title.y=element_blank(),
        legend.title=element_blank(),
        legend.text = element_text(size=14),
        legend.direction = "vertical",
        title = element_text(size=12),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size=14))+
  #scale_color_manual(values=col_IBD2)+
  #scale_shape_manual(values=c("IBD-ctrl" = 19, "CD-UC-ctrl" = 1))+
  theme(legend.box="vertical")+
  xlab("mean rho")


strainvsall.subset.mod_gatGeneDif<-strainvsall.subset.mod%>%gather(set,geneDif,c(geneDif.ibd,geneDif.cdhc,geneDif.uchc))%>%
  mutate("log2(geneDif)"=log2(geneDif+1))%>%
  mutate(set=gsub("geneDif.cdhc","CD",set))%>%
  mutate(set=gsub("geneDif.uchc","UC",set))%>%
  mutate(set=gsub("geneDif.ibd","IBD",set))
p7_geneDif<-ggplot(strainvsall.subset.mod_gatGeneDif)+
  geom_bar(aes(x=`log2(geneDif)`,y=reorder(MGS,meanR2_IBD), fill=set), stat="identity",position=position_dodge())+
  theme_bw()+
  theme(legend.position="bottom",
        axis.text.y=element_blank(),
        axis.title.y=element_blank(),
        legend.title=element_blank(),
        legend.text = element_text(size=14),
        legend.direction = "vertical",
        title = element_text(size=12),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size=14))+
  scale_fill_manual(values = col_IBD2, na.value = "white", breaks=c("IBD","CD","UC"))+
  xlab("log2(\n# sig. genes)")

```

```{r}
p_MGSoverview3<-plot_grid(p1_R2strains+theme(
  axis.text.y = element_text(size=11),
  axis.title.x = element_text(size=12),
  axis.text.x = element_text(size=10),
  legend.text = element_text(size=12),
  legend.key.size = unit(0.1, "cm"),
  legend.spacing.y = unit(0.1, "cm")),
  p2_nsigstrains+ggtitle("Strain signal")+theme(
    title = element_text(size=10),
    axis.title.x = element_text(size=12),
    axis.text.x = element_text(size=10),
    legend.text = element_text(size=12),
    legend.key.size = unit(0.4, "cm"),
    legend.spacing.y = unit(0.1, "cm")),
  p3_meanTE_DA+ggtitle("DA species \nctrl<->disease")+theme(
    title = element_text(size=10),
    axis.title.x = element_text(size=12),
    axis.text.x = element_text(size=10),
    legend.text = element_text(size=12),
    legend.key.size = unit(0.4, "cm"),
    legend.spacing.y = unit(0.1, "cm"))+
    scale_x_reverse(),
  p4_corrHR+ggtitle("Species correl.\ninflammation idx")+theme(
    title = element_text(size=10),
    axis.title.x = element_text(size=12),
    axis.text.x = element_text(size=10),
    legend.text = element_text(size=12),
    legend.key.size = unit(0.4, "cm"),
    legend.spacing.y = unit(0.1, "cm")),
                         nrow=1,align ="h", rel_widths=c(3.5,1,1,1))
p_MGSoverview3
```

```{r fig.width=18, fig.height=10}
figA<-plot_grid(plot_grid(NA,venn.p1,venn.p2, NA, align="h",axis = "tb", nrow=1,rel_widths=c(0.5,2,2,0.5)),
                plot_grid(confmat.obj.tab2,confmat.obj_ibd2.tab2, nrow=1),
                nrow=2, rel_heights = c(1.5,1),labels = c('A', 'B'))
figA

figB<-plot_grid(figA, strainR2vsTE.p2, nrow=2, rel_heights = c(2,3),labels = c('', 'C'))
figB2<-plot_grid(figA, strainR2vsTE.p1, nrow=2, rel_heights = c(2,3),labels = c('', 'C'))
figB2


figAll4<-plot_grid(figB2,p_MGSoverview3, rel_widths = c(3,4), labels = c('', 'D'))
figAll4

pdf(paste0(PATH,"Figure_StrainsD_q0.1.pdf"), height=10, width=18)
  figAll4
dev.off()

ggsave(paste0(PATH, "Figure_StrainsD_q0.1.png"), figAll4,height=10, width=19)

```

```{r}
strainvsall.subset.mod%>%
  ungroup()%>%
  count(robustDA.cdhc,enrichedIn.cdhc)
```

```{r}
strainvsall.subset.mod%>%
  ungroup()%>%
  count(robustDA.uchc,enrichedIn.uchc)
```

```{r}
strainvsall.subset.mod%>%
  ungroup()%>%
  count(robustDA,enrichedIn)
```

```{r}
strainvsall.subset.mod%>%
  ungroup()%>%
  count(robustDA,robustDA.cdhc,robustDA.uchc)

#count(enrichedIn,enrichedIn.cdhc,enrichedIn.uchc)
```


