---
title: "Make Figure for Beta-diversity in IBD"
author: "Rebecca Ansorge"
date: "21/12/2023"
output: html_document
---

### Load libraries

```{r}
library(tidyverse)
library(ggExtra)
library(patchwork)
library(cowplot)
library(formattable)
library(gridExtra)
library(rstatix)
library(ggpubr)
library(biomeStats)
library(RColorBrewer)
```
### Import data

```{r}
#PATHanalNov2023<-"~/Documents/Rebecca/onedrive/analysisNov2023/"
#PATHanalNov2023<-"~/Library/CloudStorage/OneDrive-Personal/analysisNov2023/"
PATHanalNov2023<-"~/Documents/Rebecca/onedrive/analysisNov2023/"
```

```{r}
sc.m<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.pcoa.data.rds"))
sc.m.dbRDA<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.dbrda.data.rds"))
df_all_cross<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.df_all_cross.rds"))
l_spec.long.distances.df.20<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.l_spec.long.distances.df.20.rds"))
l_spec.long.distances.df.20ibd2<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.l_spec.long.distances.df.20ibd2.rds"))
df.beta.pair<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.pair.tempInd.rds"))
ad.bc.cohort<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.pmva.bc.cohort.rds"))
ad.bc<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.pmva.bc.ibd.rds"))
ad.bc.ibd2<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.pmvaPair.bc.ibd2.rds"))
perInd_dist_vs_deltaHR_meta_slope.df<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/perInd_dist_vs_deltaHR.rds"))
dist_hr_indNest_log2_lm_perCohort<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/dist_hr_indNest_log2_lm_perCohort.rds"))
chrts_overview.obj<-readRDS(paste0(PATHanalNov2023,"PrepData/cohort_summary.rds"))

```

```{r}
df_all_cross<-df_all_cross%>%
  mutate(IBD=as.character(IBD))%>%
  mutate(IBD2=as.character(IBD2))%>%
  mutate(IBD=ifelse(IBD=="healthy","control",IBD))%>%
  mutate(IBD2=ifelse(IBD2=="healthy","control",IBD2))%>% 
  mutate(IBD2=factor(IBD2, levels=c("control","UC","CD")))

sc.m<-sc.m%>%
  mutate(IBD=as.character(IBD))%>%
  mutate(IBD2=as.character(IBD2))%>%
  mutate(IBD=ifelse(IBD=="healthy","control",IBD))%>%
  mutate(IBD2=ifelse(IBD2=="healthy","control",IBD2))%>%
  mutate(IBD=as.factor(IBD))%>%
  mutate(IBD2=as.factor(IBD2))
  

sc.m.dbRDA<-sc.m.dbRDA%>%
  mutate(IBD=as.character(IBD))%>%
  mutate(IBD2=as.character(IBD2))%>%
  mutate(IBD=ifelse(IBD=="healthy","control",IBD))%>%
  mutate(IBD2=ifelse(IBD2=="healthy","control",IBD2))%>%
  mutate(IBD=as.factor(IBD))%>%
  mutate(IBD2=as.factor(IBD2))


l_spec.long.distances.df.20<-lapply(l_spec.long.distances.df.20, function(x){
  x2<-x%>%
    mutate(IBD=as.character(IBD))%>%
    mutate(IBD2=as.character(IBD2))%>%
    mutate(IBD=ifelse(IBD=="healthy","control",IBD))%>%
    mutate(IBD2=ifelse(IBD2=="healthy","control",IBD2))%>%
    mutate(IBD=as.factor(IBD))%>%
    mutate(IBD2=as.factor(IBD2))%>% 
  mutate(IBD2=factor(IBD2, levels=c("control","UC","CD")))
  x2
})

l_spec.long.distances.df.20ibd2<-lapply(l_spec.long.distances.df.20ibd2, function(x){
  x2<-x%>%
    mutate(IBD=as.character(IBD))%>%
    mutate(IBD2=as.character(IBD2))%>%
    mutate(IBD=ifelse(IBD=="healthy","control",IBD))%>%
    mutate(IBD2=ifelse(IBD2=="healthy","control",IBD2))%>%
    mutate(IBD=as.factor(IBD))%>%
    mutate(IBD2=as.factor(IBD2))%>% 
  mutate(IBD2=factor(IBD2, levels=c("control","UC","CD")))
  x2
})

df.beta.pair<-df.beta.pair%>%
  mutate(IBD=as.character(IBD))%>%
  mutate(IBD2=as.character(IBD2))%>%
  mutate(IBD=ifelse(IBD=="healthy","control",IBD))%>%
  mutate(IBD2=ifelse(IBD2=="healthy","control",IBD2))%>% 
  mutate(IBD2=factor(IBD2, levels=c("control","UC","CD")))

perInd_dist_vs_deltaHR_meta_slope.df<-perInd_dist_vs_deltaHR_meta_slope.df%>%
  mutate(IBD=as.character(IBD))%>%
  mutate(IBD2=as.character(IBD2))%>%
  mutate(IBD=ifelse(IBD=="healthy","control",IBD))%>%
  mutate(IBD2=ifelse(IBD2=="healthy","control",IBD2))%>% 
  mutate(IBD2=factor(IBD2, levels=c("control","UC","CD")))

```

```{r}
# define color palette

#col_IBD2<-c("#E8A400","#E31A1C","#FB9A99","#1F78B4","#1F78B4")
col_IBD2<-c("gold2","#E31A1C","darkorange","#1F78B4","#1F78B4")
names(col_IBD2)<-c("IBD","CD","UC","healthy","control")

#display.brewer.pal(n=12, name="Paired")
#brewer.pal(n=12, name="Paired")
```


#### permanovas

```{r}
# make df with permanova results

ibd.pmva<-as.data.frame(ad.bc)["meta$IBD",]%>%
  mutate(comparison="IBD vs control")
cdhc.pmva<-as.data.frame(ad.bc.ibd2$CD_vs_healthy)["IBD2",]%>%
  mutate(comparison="CD vs control")
uchc.pmva<-as.data.frame(ad.bc.ibd2$UC_vs_healthy)["IBD2",]%>%
  mutate(comparison="UC vs control")
cduc.pmva<-as.data.frame(ad.bc.ibd2$CD_vs_UC)["IBD2",]%>%
  mutate(comparison="CD vs UC")
cohort.pmva<-as.data.frame(ad.bc.cohort)["meta$Cohort_precise",]%>%
  mutate(comparison="cohort")

df.pmva<-rbind(ibd.pmva,cdhc.pmva,uchc.pmva,cduc.pmva,cohort.pmva)
df.pmva2<-data.frame(comparison=df.pmva$comparison,`p.value`=df.pmva$`Pr(>F)`,R2=df.pmva$R2)%>%
  mutate(R2=round(R2,3))


df.pmva2.obj<-formattable(df.pmva2,
            list(`#` = formatter("span", style = ~ style(font.weight = "bold"))))
df.pmva2.obj

library(grid) 
mytheme <- gridExtra::ttheme_minimal(
    core = list(fg_params=list(cex = 0.7)),
    colhead = list(fg_params=list(cex = 0.7)),
    rowhead = list(fg_params=list(cex = 0.7)),
    base_colour = "black")

df.pmva2.obj.tab<-tableGrob(df.pmva2.obj,rows=c("","","","",""), theme = mytheme)

plot(df.pmva2.obj.tab)

df.pmva2.obj.tab2<-ggtexttable(df.pmva2.obj, rows = NULL, theme = ttheme("light", base_size=8,padding = unit(c(2, 2), "mm")))
#df.pmva2.obj.tab2
```

### Make first time-point ordination

#### PCoA (unconditioned)

IBD 

```{r}
o.pcoa<-ggplot(sc.m, aes(Axis.1,   Axis.2, colour = IBD)) +
    geom_point(size = 3, alpha = 0.7, aes(shape=IBD2)) +
    theme_minimal() +
    #scale_colour_manual(values = c("#40B0A6", "cornflowerblue", "#9370DB")) +
    scale_colour_manual(values=col_IBD2, breaks = names(col_IBD2))+
    xlab("PCo1 [14.5%]") +
    ylab("PCo2 [6.2%]") +
  theme(
        #plot.caption  = element_text(size = 10),
        #plot.title = element_text(size = 12, face = "plain"),
        #plot.subtitle = element_text(size =14),
        panel.grid = element_blank(),
        axis.text = element_blank(),
        legend.position=c(.08,.84),
        legend.box.background = element_rect(colour = "black"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1)) +
        guides(color = guide_legend(keywidth=0.1,keyheight=0.1,default.unit="cm"),
               shape = guide_legend(keywidth=0.1,keyheight=0.1,default.unit="cm"))+
        annotation_custom(ggplotGrob(df.pmva2.obj.tab2), xmin=-0.5, xmax=1, ymin=0, ymax=0.89)+
    ggtitle("PCoA of Bray-Curtis dissimilarities (first time-point)")
    #labs(subtitle = "PCoA of Bray-Curtis dissimilarities (first time-point)")
    #theme(title = element_text(size = 6,face = "italic"))



og.cpoa<-ggMarginal(o.pcoa, groupFill = TRUE, size =8)
og.cpoa

#ggsave("~/OneDrive/work/HildebrandGroup/IBD/analysisR2023/Beta/PCoA.1tp.pdf",og.cpoa,height=5,width=6.5)
```

unconditioned PCoA colored by cohort

```{r}

o.pcoa<-ggplot(sc.m, aes(Axis.1,   Axis.2, colour = Cohort_precise)) +
    geom_point(size = 3, alpha = 0.7, aes(shape=IBD2)) +
    theme_minimal() +
    #scale_colour_manual(values = c("#40B0A6", "cornflowerblue", "#9370DB")) +
    #scale_colour_manual(values=col_IBD2, breaks = names(col_IBD2))+
    xlab("PCo1 [14.5%]") +
    ylab("PCo2 [6.2%]") +
  theme(
        #plot.caption  = element_text(size = 10),
        #plot.title = element_text(size = 12, face = "plain"),
        #plot.subtitle = element_text(size =14),
        panel.grid = element_blank(),
        axis.text = element_blank(),
        legend.position=c(.08,.84),
        legend.box.background = element_rect(colour = "black"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1)) +
        guides(color = guide_legend(keywidth=0.1,keyheight=0.1,default.unit="cm"),
               shape = guide_legend(keywidth=0.1,keyheight=0.1,default.unit="cm"))+
        annotation_custom(ggplotGrob(df.pmva2.obj.tab2), xmin=-0.5, xmax=1, ymin=0, ymax=0.89)+
    ggtitle("PCoA of Bray-Curtis dissimilarities (first time-point)")
    #labs(subtitle = "PCoA of Bray-Curtis dissimilarities (first time-point)")
    #theme(title = element_text(size = 6,face = "italic"))



og.cpoa<-ggMarginal(o.pcoa, groupFill = TRUE, size =8)
og.cpoa

#ggsave("~/OneDrive/work/HildebrandGroup/IBD/analysisR2023/Beta/PCoA.1tp.pdf",og.cpoa,height=5,width=6.5)
```

IBD2

```{r}
sc.m<-sc.m%>%mutate(IBD=as.character(IBD))%>%mutate(IBD2=as.character(IBD2))%>%
  mutate(IBD=ifelse(IBD=="healthy","control",IBD))%>%
  mutate(IBD2=ifelse(IBD2=="healthy","control",IBD2))

o.pcoaIBD2<-ggplot(sc.m, aes(Axis.1,   Axis.2, colour = IBD2)) +
    geom_point(size = 3, alpha = 0.7, aes(shape=IBD)) +
    theme_minimal() +
    scale_colour_manual(values = col_IBD2) +
    xlab("PCo1 [14.5%]") +
    ylab("PCo2 [6.2%]") +
  theme(
        #plot.caption  = element_text(size = 10),
        #plot.title = element_text(size = 16, face = "bold"),
        #plot.subtitle = element_text(size =14),
        panel.grid = element_blank(),
        axis.text = element_blank(),
        legend.position=c(.08,.84),
        legend.box.background = element_rect(colour = "black"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1)) +
        guides(color = guide_legend(keywidth=0.1,keyheight=0.1,default.unit="cm"),
               shape = guide_legend(keywidth=0.1,keyheight=0.1,default.unit="cm"))+ 
        #annotation_custom(ggplotGrob(df.pmva2.obj.tab2), xmin=-0.5, xmax=1, ymin=0, ymax=0.89)+
        ggtitle("PCoA of Bray-Curtis dissimilarities (first time-point)")
    #labs(title = "PCoA based on Bray-Curtis dissimilarities")

og.cpoaIBD2<-ggMarginal(o.pcoaIBD2, groupFill = TRUE, size =8)
og.cpoaIBD2


ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/PCoA.1tp.pdf"),og.cpoaIBD2,height=7,width=8.5)
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/PCoA.1tp.png"),og.cpoaIBD2,height=7,width=8.5)
```


#### dbRDA conditioned by cohort

IBD
```{r}
o.dbrda<-ggplot(sc.m.dbRDA, aes(MDS2, MDS1, colour = IBD)) +
    geom_point(size = 3, alpha = 0.7, aes(shape=IBD2)) +
    theme_minimal() +
    #scale_colour_manual(values = c("#40B0A6", "cornflowerblue", "#9370DB")) +
    scale_colour_manual(values=col_IBD2, breaks = names(col_IBD2))+
    #xlab("PCo1 [14.5%]") +
    #ylab("PCo2 [6.2%]") +
  theme(
        #plot.caption  = element_text(size = 10),
        #plot.title = element_text(size = 12, face = "plain"),
        #plot.subtitle = element_text(size =14),
        panel.grid = element_blank(),
        axis.text = element_blank(),
        legend.position=c(.92,.86),
        legend.box.background = element_rect(colour = "black"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1)) +
        guides(color = guide_legend(keywidth=0.1,keyheight=0.1,default.unit="cm"),
               shape = guide_legend(keywidth=0.1,keyheight=0.1,default.unit="cm"))+
        annotation_custom(ggplotGrob(df.pmva2.obj.tab2), xmin=-1.5, xmax=-1, ymin=-2.5, ymax=-2.2)+
    ggtitle("dbRDA of Bray-Curtis diss. (first time-point; conditioned by cohort)")
    #labs(subtitle = "PCoA of Bray-Curtis dissimilarities (first time-point)")
    #theme(title = element_text(size = 6,face = "italic"))


og.dbrda<-ggMarginal(o.dbrda, groupFill = TRUE, size =8)
og.dbrda

#ggsave("~/OneDrive/work/HildebrandGroup/IBD/analysisR2023/Beta/PCoA.1tp.pdf",og.cpoa,height=5,width=6.5)
```

IBD2

```{r}

o.dbrdaIBD2<-ggplot(sc.m.dbRDA, aes(MDS2, MDS1, colour = IBD2)) +
    geom_point(size = 1, alpha = 0.7) +
    theme_minimal() +
    scale_colour_manual(values = col_IBD2, breaks = names(col_IBD2)) +
  theme(title=element_text(size=9),
        panel.grid = element_blank(),
        axis.text = element_blank(),
        legend.position=c(.1,.86),
        legend.box.background = element_rect(colour = "black"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1)) +
        guides(color = guide_legend(keywidth=0.1,keyheight=0.1,default.unit="cm", "IBD subtype"))+ 
        annotation_custom(ggplotGrob(df.pmva2.obj.tab2), xmin=1.6, xmax=2.6, ymin=2.5, ymax=2.7)+
        ggtitle("dbRDA of BC dissimilarities (conditioned by cohort)")
    #labs(title = "PCoA based on Bray-Curtis dissimilarities")

og.dbrdaIBD2<-ggMarginal(o.dbrdaIBD2, groupFill = TRUE, size =8)
og.dbrdaIBD2

#ggsave("~/OneDrive/work/HildebrandGroup/IBD/analysisR2023/Beta/PCoA.1tp.pdf",og.cpoaIBD2,height=7,width=8.5)
```

### Distance from centroid 

#### Tests


#### Crosssectional

```{r}
# each cohort separately
stat.test <- df_all_cross %>%
  group_by(cohort) %>%
  wilcox_test(distances_ccentr~IBD) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()%>% 
  add_xy_position(x = "IBD")
stat.test
```

```{r}
# blocked by cohort
df_all_cross$cohort<-as.factor(df_all_cross$cohort)
df_all_cross$IBD<-as.factor(df_all_cross$IBD)
df_all_cross$IBD2<-as.factor(df_all_cross$IBD2)

stat.test.ibd.blocked_IBDHC<-coin::wilcox_test(distances_ccentr~IBD | cohort, df_all_cross)

stat.test.ibd.blocked<-data.frame(
  cohort="all",
  .y.="distances_ccentr",
  group1=c("IBD"),
  group2=c("control"),
  n1=nrow(df_all_cross%>%filter(IBD=="IBD")),
  n2=nrow(df_all_cross%>%filter(IBD=="control")),
  statistic=coin::statistic(stat.test.ibd.blocked_IBDHC),
  p=coin::pvalue(stat.test.ibd.blocked_IBDHC)
)

stat.test.ibd.blocked<-stat.test.ibd.blocked%>%add_significance()
stat.test.ibd.blocked<-stat.test.ibd.blocked%>%mutate(groups=paste0('c("',group1,'","',group2,'")'))
stat.test.ibd.blocked$y.position=c(0.84)
stat.test.ibd.blocked$xmin=c(1)
stat.test.ibd.blocked$xmax=c(2)
stat.test.ibd.blocked
```



```{r}
df_all_cross.plot<-ggplot(df_all_cross) +
  geom_violin(aes(x=IBD, y=distances_ccentr, fill = IBD))+
  scale_fill_manual(values=col_IBD2, breaks = names(col_IBD2))+
  geom_boxplot(aes(x=IBD, y=distances_ccentr),fill="white",width=0.2,alpha=0.5)+
  theme_bw()+
  facet_wrap(~ cohort, scales = "free", ncol = 5) +
  theme(axis.title.x = element_blank(),
        legend.position = c(1, 0),
        legend.justification = c(1, 0))+
  stat_pvalue_manual(stat.test,size = 3)+ 
  scale_y_continuous(expand = expansion(mult = 0.1))+
  ylab("Distance from centroid")

df_all_cross.plot
```

IBD average and wilcoxon test blocked by cohort

```{r}
df_all_cross.plot_mean<-ggplot(df_all_cross) +
  geom_violin(aes(x=IBD, y=distances_ccentr, fill = IBD))+
  scale_fill_manual(values=col_IBD2, breaks = names(col_IBD2))+
  geom_boxplot(aes(x=IBD, y=distances_ccentr),fill="white",width=0.2,alpha=0.5)+
  theme_bw()+
  #facet_wrap(~ cohort, scales = "free", ncol = 5) +
  theme(axis.title.x = element_blank(),
        legend.position = c(1, 0),
        legend.justification = c(1, 0))+
  stat_pvalue_manual(stat.test.ibd.blocked,size = 3)+ 
  scale_y_continuous(expand = expansion(mult = 0.1))+
  ylab("Distance from centroid") 

df_all_cross.plot_mean
```

IBD2 

```{r}

stat.test.ibd2 <- df_all_cross %>%
  group_by(cohort) %>%
  #t_test(distances_ccentr~IBD2) %>%
  wilcox_test(distances_ccentr~IBD2) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()%>% 
  add_xy_position(x = "IBD2")%>%
  mutate(xmax=ifelse(cohort%in%c("He","Tian"),2,xmax))%>%
  mutate(p.adj.signif=ifelse(p.adj>0.05 & p.adj < 0.1,".",p.adj.signif))

df_all_cross.plotIBD2<-ggplot(df_all_cross) +
  geom_violin(aes(IBD2, distances_ccentr, fill = IBD)) +
  #scale_fill_brewer(type = "div", palette = "Set3") +
  #scale_fill_manual(values=c("#40B0A6", "cornflowerblue", "#9370DB"))+
  scale_fill_manual(values=col_IBD2, breaks = names(col_IBD2))+
  geom_boxplot(aes(IBD2, distances_ccentr),width=0.2,fill="white",alpha=0.5) +
  theme_bw()+
  facet_wrap(~ cohort, scales = "free", ncol = 5) +
  ylab("Distance from centroid") +
  theme(axis.title.x = element_blank(),
        legend.position = c(1, 0),
        legend.justification = c(1, 0))+
  stat_pvalue_manual(stat.test.ibd2,size = 3)+ 
  scale_y_continuous(expand = expansion(mult = 0.1))
  #ggtitle("Distance of samples from centroids")
df_all_cross.plotIBD2
```

```{r}
# blocked by cohort
df_all_cross$cohort<-as.factor(df_all_cross$cohort)

df_all_cross_CDHC<-df_all_cross%>%
  filter(IBD2%in%c("CD","control"))
df_all_cross_UCHC<-df_all_cross%>%
  filter(IBD2%in%c("UC","control"))
df_all_cross_CDUC<-df_all_cross%>%
  filter(IBD2%in%c("CD","UC"))

stat.test.ibd2.blocked_CDHC<-coin::wilcox_test(distances_ccentr~IBD2 | cohort, df_all_cross_CDHC)
stat.test.ibd2.blocked_UCHC<-coin::wilcox_test(distances_ccentr~IBD2 | cohort, df_all_cross_UCHC)
stat.test.ibd2.blocked_CDUC<-coin::wilcox_test(distances_ccentr~IBD2 | cohort, df_all_cross_CDUC)


stat.test.ibd2.blocked<-data.frame(
  cohort="all",
  .y.="distances_ccentr",
  group1=c("CD","UC","CD"),
  group2=c("control","control","UC"),
  n1=c(nrow(df_all_cross_CDHC%>%filter(IBD2=="CD")),nrow(df_all_cross_UCHC%>%filter(IBD2=="UC")),nrow(df_all_cross_CDUC%>%filter(IBD2=="CD"))),
  n2=c(nrow(df_all_cross_CDHC%>%filter(IBD2=="control")),nrow(df_all_cross_UCHC%>%filter(IBD2=="control")),nrow(df_all_cross_CDUC%>%filter(IBD2=="UC"))),
  statistic=c(coin::statistic(stat.test.ibd2.blocked_CDHC),coin::statistic(stat.test.ibd2.blocked_UCHC),coin::statistic(stat.test.ibd2.blocked_CDUC)),
  effectSize=c(coin::statistic(stat.test.ibd2.blocked_CDHC)/sqrt(nrow(df_all_cross_CDHC)),coin::statistic(stat.test.ibd2.blocked_UCHC)/sqrt(nrow(df_all_cross_UCHC)),coin::statistic(stat.test.ibd2.blocked_CDUC)/sqrt(nrow(df_all_cross_CDUC))),
  p=c(coin::pvalue(stat.test.ibd2.blocked_CDHC),coin::pvalue(stat.test.ibd2.blocked_UCHC),coin::pvalue(stat.test.ibd2.blocked_CDUC))
)

stat.test.ibd2.blocked$p.adj<-p.adjust(stat.test.ibd2.blocked$p, method="BH")
stat.test.ibd2.blocked<-stat.test.ibd2.blocked%>%add_significance()
stat.test.ibd2.blocked<-stat.test.ibd2.blocked%>%mutate(groups=paste0('c("',group1,'","',group2,'")'))
stat.test.ibd2.blocked$y.position=c(0.84,0.80,0.88)
stat.test.ibd2.blocked$xmin=c(3,2,3)
stat.test.ibd2.blocked$xmax=c(1,1,2)
stat.test.ibd2.blocked
  
```

IBD2b

```{r}
df_all_cross.plotIBD2b<-ggplot(df_all_cross) +
  geom_violin(aes(IBD2, distances_ccentr, fill = IBD2)) +
  #scale_fill_brewer(type = "div", palette = "Set3") +
  scale_fill_manual(values=col_IBD2, name="")+
  #scale_fill_manual(values=c("#40B0A6","#E8A400"))+
  geom_boxplot(aes(IBD2, distances_ccentr),width=0.2,fill="white",alpha=0.5) +
  theme_bw()+
  facet_wrap(~ cohort, scales = "free", ncol = 5) +
  ylab("Distance from centroid") +
  theme(axis.title.x = element_blank(),
        legend.position = c(1, 0),
        legend.justification = c(1, 0))+
  stat_pvalue_manual(stat.test.ibd2,size = 3)+ 
  scale_y_continuous(expand = expansion(mult = 0.1))

  #ggtitle("Distance of samples from centroids")
df_all_cross.plotIBD2b
```

IBD2b

```{r}
df_all_cross.plotIBD2b_thin<-ggplot(df_all_cross) +
  geom_violin(aes(IBD2, distances_ccentr, fill = IBD2)) +
  #scale_fill_brewer(type = "div", palette = "Set3") +
  scale_fill_manual(values=col_IBD2)+
  #scale_fill_manual(values=c("#40B0A6","#E8A400"))+
  geom_boxplot(aes(IBD2, distances_ccentr),width=0.2,fill="white",alpha=0.5) +
  theme_bw()+
  facet_wrap(~ cohort, scales = "free_x", nrow = 1) +
  ylab("Distance from centroid") +
  theme(axis.title.x = element_blank(),
        legend.position = c(1, 0),
        legend.justification = c(1, 0))+
  stat_pvalue_manual(stat.test.ibd2,size = 3)+ 
  scale_y_continuous(expand = expansion(mult = 0.1))

  #ggtitle("Distance of samples from centroids")
df_all_cross.plotIBD2b_thin
```

IBD2b average and wilcoxon test blocked by cohort

```{r}
df_all_cross.plotIBD2b_mean<-ggplot(df_all_cross) +
  geom_violin(aes(IBD2, distances_ccentr, fill = IBD2)) +
  #scale_fill_brewer(type = "div", palette = "Set3") +
  scale_fill_manual(values=col_IBD2)+
  #scale_fill_manual(values=c("#40B0A6","#E8A400"))+
  geom_boxplot(aes(IBD2, distances_ccentr),width=0.2,fill="white",alpha=0.5) +
  theme_bw()+
  #facet_wrap(~ cohort, scales = "free", ncol = 5) +
  ylab("Distance from centroid\n") +
  theme(axis.title.x = element_blank(),
        legend.position = c(1, 0),
        legend.justification = c(1, 0))+
  stat_pvalue_manual(stat.test.ibd2.blocked,size = 3)+ 
  scale_y_continuous(expand = expansion(mult = 0.1))

  #ggtitle("Distance of samples from centroids")
df_all_cross.plotIBD2b_mean
```

#### Longitudinal

IBD

```{r}
l_spec.long.distances.df_all.20<-do.call(rbind,l_spec.long.distances.df.20)

l_spec.long.distances.df_all.mean20<-l_spec.long.distances.df_all.20%>%
  group_by(cohort, IBD, iteration)%>%
  summarize(mean_dist = mean(distances_ccentr, na.rm = TRUE))%>%
  group_by(cohort, IBD)%>%
  summarize(mean_mean_dist=mean(mean_dist),stdev_mean_dist=sd(mean_dist))
l_spec.long.distances.df_all.mean20

# geom_errorbar(data=l_spec.long.distances.df_all.mean20,aes(x=IBD,y=mean_dist,ymin=mean_dist-stdev_dist, ymax=mean_dist+stdev_dist), width=.1, color="grey50")+


mean_dist_resample.plot<-ggplot(l_spec.long.distances.df_all.mean20, aes(IBD, mean_mean_dist, color = IBD)) +
    geom_errorbar(aes(x=IBD,y=mean_mean_dist,ymin=mean_mean_dist-stdev_mean_dist, ymax=mean_mean_dist+stdev_mean_dist), width=.1, color="grey50")+
    geom_point(size=3, shape=4)+
    scale_color_manual(values=col_IBD2)+
    theme_bw()+
    facet_wrap(~ cohort, scales = "free",nrow = 1) +
    ylab(paste0("Mean distance","\n","from centroid")) +
    theme(axis.title.x = element_blank())
    #ggtitle("Mean distance of samples from centroids across resampling")
#mean_dist_resample.plot

```

### wilcoxon test for each resampling timepoint

```{r}
wt.ibd.20<-lapply(c(1:20), function(n){
  #n<-3
  wt.test.ibd <- l_spec.long.distances.df.20[[n]] %>%
  group_by(cohort) %>%
  wilcox_test(distances_ccentr~IBD) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()%>% 
  add_xy_position(x = "IBD")
})

wt.sum.ibd.20<-do.call(rbind, wt.ibd.20)%>%
  group_by(cohort,group1,group2)%>%
  count(p.adj.signif) %>%
  group_by(cohort,group1,group2)%>%
  summarise(sig = paste0(n[p.adj.signif!="ns"],"*"," / ",n[p.adj.signif=="ns"],"ns"))%>%
  mutate(p.adj.signif=ifelse(sig=="* / 20ns","ns",sig))


wt.ibd.20.stat<-wt.ibd.20[[1]]%>%
  rename(p.adj.signif.sample1=p.adj.signif)%>%
  left_join(wt.sum.ibd.20,by=c("cohort", "group1", "group2"))#%>%
  #mutate(xmax=ifelse(cohort%in%c("USA","Fang_UCSD_IBD_Biobank","Lewis"),2,xmax))%>%
  #mutate(xmin=ifelse(cohort%in%c("USA","Fang_UCSD_IBD_Biobank"),1,xmin))
```




### plot for each re-sampling time-point

```{r}
l_spec.long.distances.plots.20<-lapply(c(1:20), function(n){
#n<-1
  ggplot(l_spec.long.distances.df.20[[n]]) +
    geom_violin(aes(x=IBD, y=distances_ccentr, fill = IBD),alpha=0.5) +
    #scale_fill_brewer(type = "div", palette = "Set3") +
    scale_fill_manual(values=col_IBD2, breaks=names(col_IBD2))+
    #scale_color_manual(values=c("#40B0A6", "cornflowerblue", "#9370DB")) +
    geom_boxplot(aes(x=IBD, y=distances_ccentr),fill="white",width=0.2,alpha=0.5) +
    #geom_errorbar(data=l_spec.long.distances.df_all.mean20,aes(x=IBD,y=mean_mean_dist,ymin=mean_mean_dist-stdev_mean_dist, ymax=mean_mean_dist+stdev_mean_dist), width=.1, color="grey50")+
    #geom_jitter(aes(x=IBD2, y=distances_ccentr))+
    geom_point(data=l_spec.long.distances.df_all.mean20,aes(x=IBD, y=mean_mean_dist), size=3, shape=18)+
    theme_bw()+
    facet_wrap(~ cohort, scales = "free",ncol=6) +
    stat_pvalue_manual(wt.ibd.20.stat,size = 3)+ 
    scale_y_continuous(expand = expansion(mult = 0.1))+
    theme(axis.title.x=element_blank(),
          legend.position = "none")+
    ylab(paste0("Distance from","\n","centroid and mean"))
    #ggtitle("Mean distance of samples from centroids across resampling")
})
#l_spec.long.distances.plots.20[1]
```



IBD2

```{r}
l_spec.long.distances.df_all.20ibd2<-do.call(rbind,l_spec.long.distances.df.20ibd2)

l_spec.long.distances.df_all.mean20ibd2<-l_spec.long.distances.df_all.20ibd2%>%
  group_by(cohort, IBD2, iteration)%>%
  summarize(mean_dist = mean(distances_ccentr, na.rm = TRUE))%>%
  group_by(cohort, IBD2)%>%
  summarize(mean_mean_dist=mean(mean_dist),stdev_mean_dist=sd(mean_dist))%>%
  mutate(IBD=ifelse(IBD2=="control","control","IBD"))
l_spec.long.distances.df_all.mean20ibd2

# geom_errorbar(data=l_spec.long.distances.df_all.mean20,aes(x=IBD,y=mean_dist,ymin=mean_dist-stdev_dist, ymax=mean_dist+stdev_dist), width=.1, color="grey50")+


mean_dist_resample.ibd2.plot<-ggplot(l_spec.long.distances.df_all.mean20ibd2, aes(IBD2, mean_mean_dist, color = IBD)) +
    geom_errorbar(aes(x=IBD2,y=mean_mean_dist,ymin=mean_mean_dist-stdev_mean_dist, ymax=mean_mean_dist+stdev_mean_dist), width=.1, color="grey50")+
    geom_point(size=3, shape=4)+
    scale_color_manual(values=col_IBD2, breaks=names(col_IBD2))+
    theme_bw()+
    facet_wrap(~ cohort, scales = "free",nrow = 1)+
    ylab(paste0("Mean distance","\n","from centroid")) +
    theme(axis.title.x = element_blank())
    #ggtitle("Mean distance of samples from centroids across resampling")
#mean_dist_resample.ibd2.plot

```
IBD2 - mean distribution

```{r}
l_spec.long.distances.df_all.mean20ibd2.distr<-l_spec.long.distances.df_all.20ibd2%>%
  group_by(cohort, IBD2, iteration)%>%
  summarize(mean_dist = mean(distances_ccentr, na.rm = TRUE))%>%
  group_by(cohort, IBD2)%>%
  #summarize(mean_mean_dist=mean(mean_dist),stdev_mean_dist=sd(mean_dist))%>%
  mutate(IBD=ifelse(IBD2=="control","control","IBD"))
l_spec.long.distances.df_all.mean20ibd2.distr

mean_dist_resample.distr.ibd2.plot<-ggplot(l_spec.long.distances.df_all.mean20ibd2.distr) +
    geom_violin(aes(IBD2, mean_dist, fill = IBD))+
    geom_boxplot(aes(IBD2, mean_dist),width=0.2,fill="white",alpha=0.5)+
    #geom_jitter(aes(IBD2, mean_dist),size=0.5)+
    scale_fill_manual(values=col_IBD2, breaks=names(col_IBD2))+
    theme_bw()+
    facet_wrap(~ cohort, scales = "free",nrow = 1)+
    ylab(paste0("Mean distance","\n","from centroid")) +
    theme(axis.title.x = element_blank())
    #ggtitle("Mean distance of samples from centroids across resampling")
mean_dist_resample.distr.ibd2.plot

```


### wilcoxon test for each resampling timepoint
```{r}
wt.ibd2.20<-lapply(c(1:20), function(n){
  #n<-3
  wt.test.ibd <- l_spec.long.distances.df.20ibd2[[n]] %>%
  group_by(cohort) %>%
  wilcox_test(distances_ccentr~IBD2) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()%>% 
  add_xy_position(x = "IBD2")
})

wt.sum.ibd2.20<-do.call(rbind, wt.ibd2.20)%>%
  group_by(cohort,group1,group2)%>%
  count(p.adj.signif) %>%
  group_by(cohort,group1,group2)%>%
  summarise(sig = paste0(n[p.adj.signif!="ns"],"*"," / ",n[p.adj.signif=="ns"],"ns"))%>%
  mutate(p.adj.signif=ifelse(sig=="* / 20ns","ns",sig))


wt.ibd2.20.stat<-wt.ibd2.20[[1]]%>%
  rename(p.adj.signif.sample1=p.adj.signif)%>%
  left_join(wt.sum.ibd2.20,by=c("cohort", "group1", "group2"))%>%
  mutate(xmax=ifelse(cohort%in%c("USA","Fang_UCSD_IBD_Biobank","Lewis"),2,xmax))%>%
  mutate(xmin=ifelse(cohort%in%c("USA","Fang_UCSD_IBD_Biobank"),1,xmin))


```

### blocked wilcoxon test for each resampling timepoint

```{r}
wt.ibd2.blocked.20<-lapply(c(1:20), function(n){
  #n<-3
  wt.test.ibd <- l_spec.long.distances.df.20ibd2[[n]]
  
  # blocked by cohort
  wt.test.ibd$cohort<-as.factor(wt.test.ibd$cohort)

  df_all_long_CDHC<-wt.test.ibd%>%filter(IBD2%in%c("CD","control"))
  df_all_long_UCHC<-wt.test.ibd%>%filter(IBD2%in%c("UC","control"))
  df_all_long_CDUC<-wt.test.ibd%>%filter(IBD2%in%c("CD","UC"))

  wt.test.ibd2.blocked_CDHC<-coin::wilcox_test(distances_ccentr~IBD2 | cohort, df_all_long_CDHC)
  wt.test.ibd2.blocked_UCHC<-coin::wilcox_test(distances_ccentr~IBD2 | cohort, df_all_long_UCHC)
  wt.test.ibd2.blocked_CDUC<-coin::wilcox_test(distances_ccentr~IBD2 | cohort, df_all_long_CDUC)

  wt.test.ibd2.blocked<-data.frame(
    cohort="all",
    .y.="distances_ccentr",
    group1=c("CD","UC","CD"),
    group2=c("control","control","UC"),
    n1=c(nrow(df_all_long_CDHC%>%filter(IBD2=="CD")),nrow(df_all_long_UCHC%>%filter(IBD2=="UC")),nrow(df_all_long_CDUC%>%filter(IBD2=="CD"))),
    n2=c(nrow(df_all_long_CDHC%>%filter(IBD2=="control")),nrow(df_all_long_UCHC%>%filter(IBD2=="control")),nrow(df_all_long_CDUC%>%filter(IBD2=="UC"))),
    statistic=c(coin::statistic(wt.test.ibd2.blocked_CDHC),coin::statistic(wt.test.ibd2.blocked_UCHC),coin::statistic(wt.test.ibd2.blocked_CDUC)),
    p=c(coin::pvalue(wt.test.ibd2.blocked_CDHC),coin::pvalue(wt.test.ibd2.blocked_UCHC),coin::pvalue(wt.test.ibd2.blocked_CDUC))
  )
  
  wt.test.ibd2.blocked$p.adj<-p.adjust(wt.test.ibd2.blocked$p, method="BH")
  wt.test.ibd2.blocked<-wt.test.ibd2.blocked%>%add_significance()
  wt.test.ibd2.blocked<-wt.test.ibd2.blocked%>%mutate(groups=paste0('c("',group1,'","',group2,'")'))
  wt.test.ibd2.blocked$y.position=c(0.88,0.84,0.92)
  wt.test.ibd2.blocked$xmin=c(3,2,3)
  wt.test.ibd2.blocked$xmax=c(1,1,2)
  
  return(wt.test.ibd2.blocked)
})


wt.sum.ibd2.blocked.20<-do.call(rbind, wt.ibd2.blocked.20)%>%
  mutate(p.adj.signif2=ifelse(p.adj.signif!="ns","*",p.adj.signif))%>%
  group_by(group1,group2)%>%
  count(p.adj.signif2) %>%
  summarise(sig = paste0(n[p.adj.signif2!="ns"],"*"," / ",n[p.adj.signif2=="ns"],"ns"))%>%
  mutate(p.adj.signif=ifelse(sig=="* / 20ns","ns",sig))%>%
  mutate(p.adj.signif=ifelse(p.adj.signif=="20* / ns","*",p.adj.signif))


wt.ibd2.blocked.20.stat<-wt.ibd2.blocked.20[[1]]%>%
  rename(p.adj.signif.sample1=p.adj.signif)%>%
  left_join(wt.sum.ibd2.blocked.20,by=c("group1", "group2"))

```


```{r}
l_spec.long.distances.ibd2.plots.20<-lapply(c(1:20), function(n){
  #n<-1
  ggplot(l_spec.long.distances.df.20ibd2[[n]]) +
    geom_violin(aes(x=IBD2, y=distances_ccentr, fill = IBD),alpha=0.5) +
    #scale_fill_brewer(type = "div", palette = "Set3") +
    scale_fill_manual(values=col_IBD2, breaks=names(col_IBD2))+
    #scale_color_manual(values=c("#40B0A6", "cornflowerblue", "#9370DB")) +
    geom_boxplot(aes(x=IBD2, y=distances_ccentr),fill="white",width=0.2,alpha=0.5) +
    #geom_errorbar(data=l_spec.long.distances.df_all.mean20,aes(x=IBD,y=mean_mean_dist,ymin=mean_mean_dist-stdev_mean_dist, ymax=mean_mean_dist+stdev_mean_dist), width=.1, color="grey50")+
    #geom_jitter(aes(x=IBD2, y=distances_ccentr))+
    geom_point(data=l_spec.long.distances.df_all.mean20ibd2,aes(x=IBD2, y=mean_mean_dist), size=3, shape=18)+
    theme_bw()+
    facet_wrap(~ cohort, scales = "free",ncol=6) +
    stat_pvalue_manual(wt.ibd2.20.stat,size = 3)+ 
    scale_y_continuous(expand = expansion(mult = 0.1))+
    theme(axis.title.x=element_blank(),
          legend.position = "none")+
    ylab(paste0(paste0("Distance from","\n","centroid and mean")))
    #ggtitle("Mean distance of samples from centroids across resampling")
})
```

IBD2b

```{r}
mean_dist_resample.ibd2b.plot<-ggplot(l_spec.long.distances.df_all.mean20ibd2, aes(IBD2, mean_mean_dist, color = IBD2)) +
    geom_errorbar(aes(x=IBD2,y=mean_mean_dist,ymin=mean_mean_dist-stdev_mean_dist, ymax=mean_mean_dist+stdev_mean_dist), width=.1, color="grey50")+
    geom_point(size=3, shape=4)+
    #scale_color_manual(values=c("#40B0A6","#E8A400"))+
    scale_color_manual(values=col_IBD2, breaks=names(col_IBD2)) +
    theme_bw()+
    facet_wrap(~ cohort, scales = "free",nrow = 1)+
    ylab(paste0("Mean distance","\n","from centroid")) +
    theme(axis.title.x = element_blank())
    #ggtitle("Mean distance of samples from centroids across resampling")
#mean_dist_resample.ibd2b.plot

```


```{r}
l_spec.long.distances.ibd2b.plots.20<-lapply(c(1:20), function(n){
  #n<-1
  ggplot(l_spec.long.distances.df.20ibd2[[n]]) +
    geom_violin(aes(x=IBD2, y=distances_ccentr, fill = IBD2),alpha=0.5) +
    #scale_fill_brewer(type = "div", palette = "Set3") +
    #scale_fill_manual(values=c("#40B0A6","#E8A400"))+
    scale_fill_manual(values=col_IBD2, breaks=names(col_IBD2)) +
    geom_boxplot(aes(x=IBD2, y=distances_ccentr),fill="white",width=0.2,alpha=0.5) +
    #geom_errorbar(data=l_spec.long.distances.df_all.mean20,aes(x=IBD,y=mean_mean_dist,ymin=mean_mean_dist-stdev_mean_dist, ymax=mean_mean_dist+stdev_mean_dist), width=.1, color="grey50")+
    #geom_jitter(aes(x=IBD2, y=distances_ccentr))+
    geom_point(data=l_spec.long.distances.df_all.mean20ibd2,aes(x=IBD2, y=mean_mean_dist), size=3, shape=18)+
    theme_bw()+
    facet_wrap(~ cohort, scales = "free",ncol=6) +
    stat_pvalue_manual(wt.ibd2.20.stat,size = 3)+ 
    scale_y_continuous(expand = expansion(mult = 0.1))+
    theme(axis.title.x=element_blank(),
          legend.position = "none")+
    ylab(paste0(paste0("Distance from","\n","centroid and mean")))
    #ggtitle("Mean distance of samples from centroids across resampling")
})
```

blocked mean IBD2

```{r}
l_spec.long.distances.ibd2b.blocked.plots.20<-lapply(c(1:20), function(n){
  #n<-1
  ggplot(l_spec.long.distances.df.20ibd2[[n]]) +
    geom_violin(aes(x=IBD2, y=distances_ccentr, fill = IBD2),alpha=0.5) +
    #scale_fill_brewer(type = "div", palette = "Set3") +
    #scale_fill_manual(values=c("#40B0A6","#E8A400"))+
    scale_fill_manual(values=col_IBD2, breaks=names(col_IBD2)) +
    geom_boxplot(aes(x=IBD2, y=distances_ccentr),fill="white",width=0.2,alpha=0.5) +
    #geom_errorbar(data=l_spec.long.distances.df_all.mean20,aes(x=IBD,y=mean_mean_dist,ymin=mean_mean_dist-stdev_mean_dist, ymax=mean_mean_dist+stdev_mean_dist), width=.1, color="grey50")+
    #geom_jitter(aes(x=IBD2, y=distances_ccentr))+
    geom_point(data=l_spec.long.distances.df_all.mean20ibd2,aes(x=IBD2, y=mean_mean_dist, shape=cohort), size=1)+
    theme_bw()+
    #facet_wrap(~ cohort, scales = "free",ncol=6) +
    stat_pvalue_manual(wt.ibd2.blocked.20.stat,size = 3)+ 
    scale_y_continuous(expand = expansion(mult = 0.1))+
    theme(axis.title.x=element_blank(),
          legend.position = "none")+
    ylab(paste0(paste0("Distance from","\n","centroid and mean")))
    #ggtitle("Mean distance of samples from centroids across resampling")
})
#l_spec.long.distances.ibd2b.blocked.plots.20[[1]]
```

IBD2 distribution of means

```{r}
mean_dist_resample.distr.ibd2b.plot<-ggplot(l_spec.long.distances.df_all.mean20ibd2.distr) +
    geom_violin(aes(IBD2, mean_dist, fill = IBD2))+
    geom_boxplot(aes(IBD2, mean_dist),width=0.2,fill="white",alpha=0.5)+
    #geom_jitter(aes(IBD2, mean_dist),size=0.5)+
    #scale_fill_manual(values=c("#40B0A6","#E8A400"))+
    scale_fill_manual(values=col_IBD2, breaks=names(col_IBD2)) +
    theme_bw()+
    facet_wrap(~ cohort, scales = "free",nrow = 1)+
    ylab(paste0("Mean distance","\n","from centroid")) +
    theme(axis.title.x = element_blank())
    #ggtitle("Mean distance of samples from centroids across resampling")
mean_dist_resample.distr.ibd2b.plot
```


### Average beta-diversity among 4 consecutive samples per Individual

IBD

```{r}

stat.test.Ind <- df.beta.pair %>%
  wilcox_test(average~IBD) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance() %>% 
  add_xy_position(x = "IBD")

beta.pair.p1<-ggplot(df.beta.pair) +
  geom_violin(aes(IBD, average, fill = IBD)) +
  geom_boxplot(aes(IBD, average, fill = IBD),fill="white",width=0.2,alpha=0.5) +
  #geom_jitter()+
  #scale_fill_brewer(type = "div", palette = "Set2") +
  #scale_fill_manual(values=c("#40B0A6", "cornflowerblue", "#9370DB"))+
  scale_fill_manual(values=col_IBD2)+
  theme_bw() +
  ylab(paste0("Mean distance","\n","per individual")) +
  theme(axis.title.x = element_blank())+
  stat_pvalue_manual(stat.test.Ind,size = 3)+ 
  scale_y_continuous(expand = expansion(mult = 0.1))+
  ggtitle("Not accounted for cohort effect")
  #labs(title = "Average distances between 4 consecutive samples across conditions")
beta.pair.p1
```

IBD2

not accounted for cohort

```{r}

stat.test.Ind.ibd2 <- df.beta.pair %>%
  #filter(cohort%in%c("HMP","HMP2","LIBD"))%>%
  wilcox_test(average~IBD2) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()%>% 
  add_xy_position(x = "IBD2")
stat.test.Ind.ibd2
```

accounting for cohort


```{r}
#df.beta.pair %>%
# group_by(cohort)%>%
#  count(IBD2)%>%View()
```

Each cohort separately seems to have too few samples to reas significance

```{r}
df.beta.pair %>%
  filter(cohort!="Lewis")%>%
  group_by(cohort) %>%
  wilcox_test(average~IBD2) %>%
  adjust_pvalue(method = "BH")%>%
  add_significance()%>% 
  add_xy_position(x = "IBD2")
```

check whether cohort effect is significant

* for controls and UC there is no cohort effect
* for CD there is a cohort effect

```{r}
df.beta.pair%>%
  filter(IBD2%in%c("control"))%>%
  wilcox_test(average~cohort)

df.beta.pair%>%
  filter(IBD2%in%c("CD"))%>%
  wilcox_test(average~cohort)

df.beta.pair%>%
  filter(IBD2%in%c("UC"))%>%
  wilcox_test(average~cohort)
```

Testing the intra-individual effect between UC and controls across all cohorts

```{r}
df.beta.pair %>%
  filter(IBD2!="CD")%>%
  wilcox_test(average~IBD2) %>%
  add_significance()
```

Testing the intra-individual effect between CD and controls across all cohorts that were not significantly different from each other

```{r}
df.beta.pair %>%
  filter(cohort%in%c("HMP","HMP2","LIBD"))%>%
  #filter(IBD2!="UC")%>%
  wilcox_test(average~IBD2) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()%>% 
  add_xy_position(x = "IBD2")
```

Testinf intra-individual effect blocked by cohort

```{r}

# blocked by cohort
df.beta.pair$cohort<-as.factor(df.beta.pair$cohort)

df.beta.pair_CDHC<-df.beta.pair%>%
  filter(cohort!="Lewis")%>%
  filter(IBD2%in%c("CD","control"))
df.beta.pair_UCHC<-df.beta.pair%>%
  filter(cohort!="Lewis")%>%
  filter(IBD2%in%c("UC","control"))
df.beta.pair_CDUC<-df.beta.pair%>%
  filter(cohort!="Lewis")%>%
  filter(IBD2%in%c("CD","UC"))

stat.test.Ind.ibd2.blocked_CDHC<-coin::wilcox_test(average~IBD2 | cohort, df.beta.pair_CDHC)
stat.test.Ind.ibd2.blocked_UCHC<-coin::wilcox_test(average~IBD2 | cohort, df.beta.pair_UCHC)
stat.test.Ind.ibd2.blocked_CDUC<-coin::wilcox_test(average~IBD2 | cohort, df.beta.pair_CDUC)

stat.test.Ind.ibd2.blocked<-data.frame(
  cohort="all",
  .y.="distances_ccentr",
  group1=c("CD","UC","CD"),
  group2=c("control","control","UC"),
  n1=c(nrow(df.beta.pair_CDHC%>%filter(IBD2=="CD")),nrow(df.beta.pair_UCHC%>%filter(IBD2=="UC")),nrow(df.beta.pair_CDUC%>%filter(IBD2=="CD"))),
  n2=c(nrow(df.beta.pair_CDHC%>%filter(IBD2=="control")),nrow(df.beta.pair_UCHC%>%filter(IBD2=="control")),nrow(df.beta.pair_CDUC%>%filter(IBD2=="UC"))),
  statistic=c(coin::statistic(stat.test.Ind.ibd2.blocked_CDHC),coin::statistic(stat.test.Ind.ibd2.blocked_UCHC),coin::statistic(stat.test.Ind.ibd2.blocked_CDUC)),
  p=c(coin::pvalue(stat.test.Ind.ibd2.blocked_CDHC),coin::pvalue(stat.test.Ind.ibd2.blocked_UCHC),coin::pvalue(stat.test.Ind.ibd2.blocked_CDUC))
)

stat.test.Ind.ibd2.blocked$p.adj<-p.adjust(stat.test.Ind.ibd2.blocked$p, method="BH")
stat.test.Ind.ibd2.blocked<-stat.test.Ind.ibd2.blocked%>%add_significance()
stat.test.Ind.ibd2.blocked<-stat.test.Ind.ibd2.blocked%>%mutate(groups=paste0('c("',group1,'","',group2,'")'))
stat.test.Ind.ibd2.blocked$y.position=c(1.02,0.95,1.09)
stat.test.Ind.ibd2.blocked$xmin=c(3,2,3)
stat.test.Ind.ibd2.blocked$xmax=c(1,1,2)
stat.test.Ind.ibd2.blocked<-stat.test.Ind.ibd2.blocked%>%
  mutate(p.adj.signif=ifelse(p.adj>0.05 & p.adj<0.1, "o", p.adj.signif))
stat.test.Ind.ibd2.blocked<-stat.test.Ind.ibd2.blocked%>%
  mutate(effectSize=statistic/sqrt(n1+n2))
```

```{r}
beta.pair.p2<-ggplot(df.beta.pair) +
  geom_violin(aes(IBD2, average, fill = IBD)) +
  geom_boxplot(aes(IBD2, average, fill = IBD),fill="white",width=0.2,alpha=0.5) +
  scale_fill_manual(values=col_IBD2, breaks=names(col_IBD2))+
  theme_bw()+
  ylab(paste0("Mean distance","\n","per individual")) +
  theme(axis.title.x = element_blank())+
  stat_pvalue_manual(stat.test.Ind.ibd2.blocked,size = 3)+ 
  scale_y_continuous(expand = expansion(mult = 0.1))
  #labs(title = "Average distances between 4 consecutive samples across conditions")
beta.pair.p2
```




IBD2b

```{r}
beta.pair.p2b<-ggplot(df.beta.pair) +
  geom_violin(aes(IBD2, average, fill = IBD2)) +
  geom_boxplot(aes(IBD2, average, fill = IBD2),fill="white",width=0.2,alpha=0.5) +
  scale_fill_manual(values=col_IBD2, breaks=names(col_IBD2))+
  theme_bw()+
  ylab(paste0("Mean dissimilarity","\n","per individual")) +
  theme(axis.title.x = element_blank())+
  stat_pvalue_manual(stat.test.Ind.ibd2.blocked,size = 3)+ 
  scale_y_continuous(expand = expansion(mult = 0.1))
beta.pair.p2b
```


```{r fig.width=8, fig.height=3}
beta.pair.p2b.cohort<-ggplot(df.beta.pair%>%mutate(cohort=gsub("Fang_UCSD_","",cohort))) +
  geom_violin(aes(IBD2, average, fill = IBD2)) +
  geom_boxplot(aes(IBD2, average, fill = IBD2),fill="white",width=0.2,alpha=0.5) +
  geom_jitter(aes(IBD2, average, fill = IBD2), size=1, alpha=0.5)+
  scale_fill_manual(values=col_IBD2, breaks=names(col_IBD2), name="")+
  theme_bw()+
  ylab(paste0("Mean dissimilarity","\n","per individual")) +
  theme(axis.title.x = element_blank())+
  facet_grid(~cohort)+
  scale_y_continuous(expand = expansion(mult = 0.1))

beta.pair.p2b.cohort

pdf(paste0(PATHanalNov2023,"Downstream/Results/Figures/supplFig_meanIndDist.pdf"),width=8,height=3)
beta.pair.p2b.cohort
dev.off()
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/supplFig_meanIndDist.png"),beta.pair.p2b.cohort,width=8,height=3)

```

```{r fig.width=8, fig.height=3}
p_avDist_deltaHR_4TP<-ggplot(df.beta.pair, aes(log2(maxdist), log2(deltaHumanRds))) +
  geom_point(aes(color=cohort), alpha=0.3, size=1.5) +
  geom_smooth(aes(color=cohort), method = "lm",se = FALSE) +
  theme_bw() +
  facet_grid(.~IBD2, scales="free")+ 
  scale_colour_brewer(palette = "Dark2")+
  xlab("log2(Maxium pairwise distance among 4 consecutive samples)")+
  ylab(paste0("Maximum","\n","delta human reads"))
p_avDist_deltaHR_4TP

pdf(paste0(PATHanalNov2023,"Downstream/Results/Figures/supplFig_maxdeltaHR_maxDist.pdf"),width=8,height=3)
p_avDist_deltaHR_4TP
dev.off()
```

```{r fig.width=12, fig.height=3}
perInd_dist_vs_deltaHR_meta_slope.df<-perInd_dist_vs_deltaHR_meta_slope.df%>%
  mutate(cohort=as.character(cohort))%>%
  mutate(cohort=ifelse(cohort=="Fang_UCSD_IBD_Biobank","IBD_Biobank",cohort))%>%
  mutate(IBD2=as.character(IBD2))%>% 
  mutate(IBD2=factor(IBD2, levels=c("control","UC","CD")))

perInd_dist_vs_deltaHR_meta_singleslope.df<-perInd_dist_vs_deltaHR_meta_slope.df%>%
  distinct(IndividualID2, .keep_all = TRUE)

slope2_summary_Ind<-perInd_dist_vs_deltaHR_meta_singleslope.df%>%
  group_by(cohort,IBD2)%>%
  count(slope2)%>%
  spread(slope2,n)

cohort_summary_Ind<-perInd_dist_vs_deltaHR_meta_singleslope.df%>%
  group_by(cohort)%>%
  count()


slope_summary_Ind<-perInd_dist_vs_deltaHR_meta_singleslope.df%>%
  group_by(cohort,IBD2)%>%
  summarize(mean_slope=mean(slope))

slopeB_summary_Ind<-perInd_dist_vs_deltaHR_meta_singleslope.df%>%
  group_by(cohort,IBD2,slope2)%>%
  summarize(mean_slope=mean(slope))

# p_pairDist_deltaHR<-ggplot(perInd_dist_vs_deltaHR_meta_slope.df, 
#   aes(microbiome_bc_distance_log2,delta_human_reads_log2))+
#   geom_point(alpha=0.2, size=0.5, color="grey40")+
#   geom_smooth(aes(color=slope2, group = factor(IndividualID2)),method="lm", se = FALSE, size=0.3)+
#   #theme(legend.position="none")+
#   facet_grid(~IBD2,scales="free")+
#   scale_color_manual(#values=c("#1B9E77","#D95F02","#7570B3","#E7298A","#66A61E","grey20","red","#E6AB02"),
#     values=c("grey20","red"))+
#   theme_bw()+
#   xlab("log2(Pairwise distances between timepoints)")+
#   ylab(paste0("log2(Delta human reads","\n", "between timepoints)"))
# p_pairDist_deltaHR

slope2_summary_Ind_all<-slope2_summary_Ind%>%
  group_by(IBD2)%>%
  summarize(negative=sum(neg),positive=sum(pos))%>%
  mutate(negative_perc=paste0(negative," (",round((negative/(negative+positive))*100,0),"%)"))%>%
  mutate(positive_perc=paste0(positive," (",round((positive/(negative+positive))*100,0),"%)"))

p_pairDist_deltaHR<-ggplot(perInd_dist_vs_deltaHR_meta_slope.df, 
  aes(microbiome_bc_distance_log2,delta_human_reads_log2))+
  geom_point(alpha=0.2, size=0.5, color="grey40")+
  geom_smooth(method="lm" , se = FALSE, size=0.5,color="red")+ #method="lm", 
  geom_text(data=slope2_summary_Ind_all, aes(label=paste0("# positive: ",positive_perc),x=-4,y=-25), color="red")+
  geom_text(data=slope2_summary_Ind_all, aes(label=paste0("# negative: ",negative_perc),x=-4,y=-28), color="grey20")+
  #theme(legend.position="none")+
  facet_grid(~IBD2)+
  scale_color_manual(#values=c("#1B9E77","#D95F02","#7570B3","#E7298A","#66A61E","grey20","red","#E6AB02"),
    values=c("grey20","red"))+
  theme_bw()+
  xlab("log2(Pairwise distances between timepoints)")+
  ylab(paste0("log2(Delta human reads","\n", "between timepoints)"))
p_pairDist_deltaHR


slope2_summary_Ind_all_cohort<-slope2_summary_Ind%>%
  group_by(cohort,IBD2)%>%
  summarize(negative=sum(neg),positive=sum(pos))%>%
  mutate(negative_perc=paste0(negative," (",round((negative/(negative+positive))*100,0),"%)"))%>%
  mutate(positive_perc=paste0(positive," (",round((positive/(negative+positive))*100,0),"%)"))%>%
  mutate(neg_perc=paste0(round((negative/(negative+positive))*100,0),"%"))%>%
  mutate(pos_perc=paste0(round((positive/(negative+positive))*100,0),"%"))

disthr_sig_perCohort<-dist_hr_indNest_log2_lm_perCohort%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cohort")%>%
  mutate(significance=ifelse(p<0.001,"***",ifelse(p<0.01,"**",ifelse(p<0.05,"*",ifelse(p<0.1,".","")))))%>%
  mutate(cohortsig=paste(cohort,significance))

cohortsig<-disthr_sig_perCohort$cohortsig
names(cohortsig)<-disthr_sig_perCohort$cohort

perInd_dist_vs_deltaHR_meta_slope.df$cohort2<-factor(perInd_dist_vs_deltaHR_meta_slope.df$cohort, levels=c("Lewis","USA","IBD_Biobank","HMP2","HMP","LIBD"))


p_pairDist_deltaHR2<-ggplot(perInd_dist_vs_deltaHR_meta_slope.df, 
  aes(microbiome_bc_distance_log2,delta_human_reads_log2))+
  geom_point(alpha=0.2, size=0.5, aes(color=IBD2))+#color="grey40")+
  geom_smooth(aes(color=IBD2),method="lm" , se = FALSE, size=0.5)+ #method="lm", 
  scale_color_manual(values=col_IBD2)+
  facet_grid(~cohort2, labeller=labeller(cohort2=cohortsig))+
  theme_bw()+
  xlab("log2(Intra-individual dissimilarities between timepoints)")+
  ylab(paste0("log2(Difference inflammation","\n", "index between timepoints)"))+
  ggtitle("Intra-individual correlation of inflammation index and microbiome dissimilarity (p < 2e-16 ***)")
p_pairDist_deltaHR2

cohortN<-paste0(cohort_summary_Ind$cohort," n=",cohort_summary_Ind$n)
names(cohortN)<-cohort_summary_Ind$cohort

perInd_dist_vs_deltaHR_meta_singleslope.df$cohort2<-factor(perInd_dist_vs_deltaHR_meta_singleslope.df$cohort, levels=c("Lewis","USA","IBD_Biobank","HMP2","HMP","LIBD"))
slope2_summary_Ind_all_cohort$cohort2<-factor(slope2_summary_Ind_all_cohort$cohort, levels=c("Lewis","USA","IBD_Biobank","HMP2","HMP","LIBD"))
annot<-data.frame(x="control", y=0.38, lab=c("Positive iCC:"),cohort="Lewis")
annot$cohort2<-factor(annot$cohort, levels=c("Lewis","USA","IBD_Biobank","HMP2","HMP","LIBD"))

p_pindSlope_deltaHR<-ggplot(perInd_dist_vs_deltaHR_meta_singleslope.df, 
  aes(IBD2,slope))+
  geom_hline(yintercept=0, linetype=2, color="grey50")+
  geom_violin(aes(fill=IBD2))+
  geom_boxplot(outlier.size = 0.1, width=0.15)+
  #geom_text(data=slope2_summary_Ind_all_cohort, aes(x=IBD2,y=0.39, label=positive), size=3)+
  geom_text(data=annot, aes(x=x,y=y, label=lab), size=3.3, hjust=0.3)+
  geom_text(data=slope2_summary_Ind_all_cohort, aes(x=IBD2,y=0.38, label=pos_perc), size=3.5)+
  scale_fill_manual(values=col_IBD2)+
  ylim(-0.25,0.4)+
  theme_bw()+
  ylab(paste0("Correlation coefficient\nper individual"))+
  theme(axis.text.x=element_blank(),
        axis.title.x=element_blank(),
        axis.ticks.x = element_blank())+
  facet_grid(~cohort2, labeller=labeller(cohort2=cohortN))
p_pindSlope_deltaHR
```

```{r}
annot2<-data.frame(x=c("control","control"), y=c(0.25,0.75), lab=c("Positive","Negative"),cohort=c("Lewis","Lewis"))
annot2$cohort2<-factor(annot2$cohort, levels=c("Lewis","USA","IBD_Biobank","HMP2","HMP","LIBD"))

slope2_summary_Ind_all_cohort%>%
  ungroup()%>%
  filter(IBD2=="CD")%>%
  mutate(pos=as.numeric(gsub("%","",pos_perc)))%>%
  summarize(mean=mean(pos),median=median(pos))

slope2_summary_Ind_all_cohort%>%
  ungroup()%>%
  filter(IBD2=="UC")%>%
  mutate(pos=as.numeric(gsub("%","",pos_perc)))%>%
  summarize(mean=mean(pos),median=median(pos))

slope2_summary_Ind_all_cohort%>%
  ungroup()%>%
  filter(IBD2=="control")%>%
  mutate(pos=as.numeric(gsub("%","",pos_perc)))%>%
  summarize(mean=mean(pos),median=median(pos))

p_pind_barHR<-ggplot(slope2_summary_Ind_all_cohort%>%gather(slope,n,c(negative,positive)),
  aes(IBD2,n))+
  geom_bar(stat="identity",position="fill",aes(fill=slope))+
  geom_hline(yintercept=0.5, linetype=2, color="grey50")+
  geom_text(data=annot2, aes(x=x,y=y, label=lab), size=3.3, hjust=0.2, color=c("Salmon","#40B0A6"))+
  #geom_text(data=slope2_summary_Ind_all_cohort, aes(x=IBD2,y=45, label=pos_perc), size=3.5, hjust=0, color="Salmon")+
  scale_fill_manual(values = c("#40B0A6","salmon"), na.value = "white", breaks=c("negative","positive"))+
  #ylim(-0.25,0.4)+
  theme_bw()+
  ylab(paste0("Number of individuals"))+
  theme(#axis.text.x=element_blank(),
        axis.title.x=element_blank(),
        axis.ticks.x = element_blank(),
        legend.position="top")+
  facet_grid(~cohort2, labeller=labeller(cohort2=cohortN))+
  guides(fill = guide_legend(keywidth=0.3,keyheight=0.1,default.unit="cm", ""))
p_pind_barHR
```

### comparison of high vs low BD dissimilarities

```{r}
dist.HL_hr_cond_gat_intrChrt<-readRDS(paste0(PATHanalNov2023,"humanReads/dist.HL_hr_cond_gat_intrChrt.rds"))
dist.HL_hr_cond_gat_intrChrt<-dist.HL_hr_cond_gat_intrChrt%>%
  mutate(HumanRds_group1=gsub("HR","Iidx",HumanRds_group1))%>%
  mutate(HumanRds_group2=gsub("HR","Iidx",HumanRds_group2))%>%
  mutate(comparison=gsub("HR","Iidx",comparison))%>%
  mutate(comparison2=gsub("HR","Iidx",comparison2))

dist.HL_hr_ind_gat_intraInd<-readRDS(paste0(PATHanalNov2023,"humanReads/dist.HL_hr_ind_gat_intraInd.rds"))
dist.HL_hr_ind_gat_intraInd<-dist.HL_hr_ind_gat_intraInd%>%
  mutate(HumanRds_group1=gsub("HR","Iidx",HumanRds_group1))%>%
  mutate(HumanRds_group2=gsub("HR","Iidx",HumanRds_group2))%>%
  mutate(comparison=gsub("HR","Iidx",comparison))

col_comp<-c("darkorange1","khaki2","#1F78B4")
names(col_comp)<-c("control-HighIidx","control-LowIidx","control")
ctrllabs<-c("control-flare","control-remission","control")

testtest<-stat.test.Ind.ibd2.blocked%>%
  mutate(p.adj.signif="***")%>%
  select(c(p.adj.signif,y.position,xmin,xmax,group1,group2))
testtest$y.position<-c(1.2,1.1,1.3)
testtest$group1<-c("control-HighIidx","control-HighIidx","control-LowIidx")
testtest$group2<-c("control-LowIidx","control","control")


p_dist_HR_all<-
  ggplot(dist.HL_hr_cond_gat_intrChrt, aes(x=comparison2,y=dist))+
  geom_violin(aes(fill=comparison2))+
  geom_boxplot(outlier.size = 0.1, width=0.15)+  
  stat_pvalue_manual(testtest,size = 3)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45,hjust=1),
        legend.position="bottom",
        legend.direction="vertical",
        legend.title = element_blank(),
        legend.key.size = unit(0, 'lines'),
        legend.margin = margin(-2, -2, -2, -2))+
  scale_fill_manual(values=col_comp, breaks=names(col_comp), labels=ctrllabs)+
  ylab("Inter-individual\nBC dissimilarity")+
  theme(axis.text.x=element_blank(), axis.title.x=element_blank(), axis.ticks.x=element_blank())+
  guides(fill = guide_legend(keywidth=0.3,keyheight=0.1,default.unit="cm"))
p_dist_HR_all

col_comp2<-c("darkorange1","khaki2","violet")
names(col_comp2)<-c("HighIidx","LowIidx","LowIidx-HighIidx")
remlabs<-c("flare-like","remission-like","flare-remission")

testtest2<-stat.test.Ind.ibd2.blocked%>%
  select(c(p.adj.signif,y.position,xmin,xmax,group1,group2))
testtest2$y.position<-c(1.2,1.1,1.3)
testtest2$p.adj.signif<-c("ns","***","***")
testtest2$group1<-c("HighIidx","HighIidx","LowIidx")
testtest2$group2<-c("LowIidx","LowIidx-HighIidx","LowIidx-HighIidx")

p_iiv_all<-ggplot(dist.HL_hr_ind_gat_intraInd, aes(x=comparison,y=dist))+
  geom_violin(aes(fill=comparison))+
  geom_boxplot(outlier.size = 0.1, width=0.15)+
  stat_pvalue_manual(testtest2,size = 3)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45,hjust=1),
        legend.position="bottom",
        legend.direction="vertical",
        legend.title = element_blank(),
        legend.key.size = unit(0, 'lines'),
        legend.margin = margin(-2, -2, -2, -2))+
  scale_fill_manual(values=col_comp2, breaks=names(col_comp2),labels=remlabs)+
  ylab("Intra-individual\nBC dissimilarity")+
  theme(axis.text.x=element_blank(), axis.title.x=element_blank(), axis.ticks.x=element_blank())+
  guides(fill = guide_legend(keywidth=0.3,keyheight=0.1,default.unit="cm"))

p_iiv_all


(dist.HL_hr_ind_gat_intraInd%>%filter(HumanRds_group1=="LowIidx"))$SmplID1%>%unique()%>%length()

```

### combine crosssectional and longitudinal

```{r}
combi1<-plot_grid(l_spec.long.distances.plots.20[[1]]+ggtitle("Longitudinal cohorts - violin-plot displays sample 1"),beta.pair.p1+theme(legend.position = "none")+ggtitle("Intra-individual"),ncol=2,rel_widths = c(2.5,1), align = "h", axis = "b", labels=c("C","D"))
#combi1

combi2<-plot_grid(df_all_cross.plot+ggtitle("Crossectional cohorts"),combi1,ncol=1,rel_heights = c(2,1), labels=c("B",""))
#combi2

combi3<-plot_grid(og.dbrdaIBD2,combi2,ncol=2,rel_widths = c(1,1.2), align = "v", axis = "lr", labels=c("A",""))
combi3b<-plot_grid(og.dbrda,combi2,ncol=2,rel_widths = c(1,1.2), align = "v", axis = "lr", labels=c("A",""))
#combi3

combi3b
#ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/Fig_Beta.ibd.pdf"),combi3,height=7,width=15)
#ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/Fig_Beta.ibdb.pdf"),combi3b,height=7,width=15)

```


```{r}
combi1ibd2b<-plot_grid(l_spec.long.distances.ibd2b.plots.20[[1]]+ggtitle("Longitudinal cohorts - violin-plot displays sample 1"),beta.pair.p2b+theme(legend.position = "none")+ggtitle("Intra-individual"),ncol=2,rel_widths = c(3.8,1), align = "h", axis = "b", labels=c("C","D"))

combi2ibd2b<-plot_grid(df_all_cross.plotIBD2b+ggtitle("Crossectional cohorts"),combi1ibd2b,ncol=1,rel_heights = c(2,1), labels=c("B",""))

combi3ibd2b<-plot_grid(og.dbrdaIBD2,combi2ibd2b,ncol=2,rel_widths = c(1,1.5), align = "v", axis = "lr", labels=c("A",""))
combi3ibd2b
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/Fig_Beta.ibd2b.pdf"),combi3ibd2b,height=7,width=17)

```

```{r}

combi2ibd2b_supp<-plot_grid(df_all_cross.plotIBD2b+ggtitle("Crossectional cohorts"),l_spec.long.distances.ibd2b.plots.20[[1]]+ggtitle("Longitudinal cohorts - violin-plot displays sample 1"),ncol=1,rel_heights = c(2,1), labels=c("A","B"))

pdf(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_Fig_centroidCohorts.pdf"),height=7,width=10)
combi2ibd2b_supp
dev.off()
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_Fig_centroidCohorts.png"),combi2ibd2b_supp,height=7,width=10)
```



```{r, fig.height= 8, fig.width=14}

blockedViolins<-plot_grid(  
  df_all_cross.plotIBD2b_mean+ggtitle("Crossectional\ncohorts")+theme(legend.position="none", title = element_text(size=8.5), axis.text.x=element_text(angle=45, hjust=1)),
  l_spec.long.distances.ibd2b.blocked.plots.20[[1]]+ggtitle("Resampling\nlongit. cohorts")+theme(title = element_text(size=8.5), axis.text.x=element_text(angle=45, hjust=1)),
  beta.pair.p2b+theme(legend.position = "none")+ggtitle("Intra-individual\nmean BC")+theme(title = element_text(size=8.5), axis.text.x=element_text(angle=45, hjust=1)),
  p_dist_HR_all+ggtitle("Inflam. groups\ninter-indiv. BC")+theme(title = element_text(size=8.5)),
  p_iiv_all+ggtitle("Inflam. groups\nintra-indiv. BC")+theme(title = element_text(size=8.5)),
  nrow=1,rel_widths = c(1,1,1,1,1), align = "h", axis = "t", labels=c("C","D","E","F","G")
  )
  
# IntraIndCors<-plot_grid(
#   p_avDist_deltaHR_4TP+ggtitle("Maximum distance and delta human reads per individual")+theme(legend.position="none"),
#   p_pairDist_deltaHR+ggtitle("Intra-individual correlation of timepoint distances and delta human reads")+theme(legend.position="none"),
#   ncol=1,rel_heights = c(1,1), align = "v", axis = "b", labels=c("D","E")
#   )


# blockedViolins_IntraIndCors<-plot_grid(
#   blockedViolins,
#   IntraIndCors,
#   ncol=1, rel_heights=c(1,2,2)
# )

p_pind_deltaHR<-plot_grid(p_pairDist_deltaHR2 + theme(legend.position = "none",title = element_text(size=8.5)),
                          p_pind_barHR + ggtitle("Individuals’ correlation direction of Δ inflammation ~ Δ microbiome composition") + theme(legend.position = "none",title = element_text(size=8.5)),
                          ncol=1,align="v",axis="l", labels=c("H","I")
                          )


blockedViolins_IntraIndCors<-plot_grid(
  blockedViolins,
  p_pind_deltaHR,
  ncol=1, rel_heights=c(1,2), align="v", axis="lr"
)
blockedViolins_IntraIndCors


p2 <- cowplot::ggdraw() + cowplot::draw_image(paste0(PATHanalNov2023,"Downstream/Results/Figures/resampling.png"))

p2<-plot_grid(p2,NULL, nrow=1, labels=c("Resampling strategy",""), rel_widths = c(10,0.5),hjust = 0, vjust= 1, label_fontface="plain")

#p2
chrts_overview.obj.tab2<-ggtexttable(chrts_overview.obj, rows = NULL, theme = ttheme("light", base_size=10, padding = unit(c(3, 2), "mm")))
p3<-plot_grid(chrts_overview.obj.tab2,NULL, nrow=1, labels=c("Cohort overview",""), rel_widths = c(10,0.5),hjust = 0, vjust= 1, label_fontface="plain",label_size = 11)

combi1ibd2b.blocked<-plot_grid(
  plot_grid(p3,og.dbrdaIBD2,ncol=1,labels=c("A","B"), align = "v", axis = "l", hjust = 0, vjust= 1, rel_heights = c(1.5,2)),
  blockedViolins_IntraIndCors,
  ncol=2,rel_widths = c(1,1.4), align = "h", axis = "b")
combi1ibd2b.blocked



ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/Fig_Beta.ibd2b.blocked.pdf"),combi1ibd2b.blocked,height=8,width=14)
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/Fig_Beta.ibd2b.blocked.png"),combi1ibd2b.blocked,height=8,width=14)

```