---
title: "Prevalence filtration"
author: "SRomano (adapted by RAnsorge)"
date: '2023-11-11'
output: html_document
editor_options: 
  chunk_output_type: console
---
# Install load packages
```{r setup, include=FALSE}

list.of.packages <- c("dplyr", "reshape2", "meta", "ggplot2", "lattice")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq", "microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)

library(meta)
library(DHARMa)
library(nlme)
library(ggplot2)
library(phyloseq)
library(dplyr)
library(microbiome)
library(reshape2)
library(meta)
library(lattice)
#library(broomExtra)

```

```{r}
PATHanalNov2023<-"~/Documents/Rebecca/onedrive/analysisNov2023/"
```

# Load and clean data
Load the abundance file and the metadata

```{r, message=F}

spec<-read.delim(paste0(PATHanalNov2023,"PrepData/IBD.MGSabundance.txt"), header = T, row.names = 1)
dim(spec)

rownames(spec) %>% tail
```

The last row is "other". This will be used to calcualte the relative abundances but will then be removed
```{r, message =F, warning =F}

spec.tss<-as.data.frame(prop.table(as.matrix(spec), 2))
# now remove other

spec.tss<-spec.tss[-1*which(rownames(spec.tss) == "other"),]

dim(spec)[1]
```
There are `dim(spec)[1]` Species and 4839 samples


# Load and format the metadata

```{r, message =F, warning =F}
meta<-read.delim(paste0(PATHanalNov2023,"PrepData/IBDmeta.cur.filt.mc.matched.txt"), header = T, sep = "\t")
dim(meta)
sapply(meta, class)

meta$IBD<-factor(meta$IBD, levels = c("healthy", "IBD"))
meta$IBD2<-factor(meta$IBD2, levels = c("healthy", "UC", "CD"))
```

Load function and check prevalences
```{r}
source(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/Pevalence_functions.r"))
# Need to split datasets
spec<-as.data.frame(t(spec))
all(rownames(spec) == meta$SmplID)

spec$cohort<-meta$Cohort_precise

l_spec<-split(spec, f = as.factor(spec$cohort))
# remove cohort
l_spec<-lapply(l_spec, function(x) x[, which(names(x)!= "cohort")])
l_spec<-lapply(l_spec, function(x) x[, which(names(x)!= "other")])

l_spec<-lapply(l_spec, function(x) as.data.frame(t(x)))

# prevalence filtration (prev is in %)
spec_3x2.5<-prevalence.matching(l_spec, n.studies = 3, prev = 2.5) # for now choose 3 studies and 2.5% prevalence
length(spec_3x2.5)

write.csv(spec_3x2.5,paste0(PATHanalNov2023,"Downstream/Results/prev.filt.MGS.txt"), row.names = FALSE)

spec_5x10<-prevalence.matching(l_spec, n.studies = 3, prev = 10)
length(spec_5x10)
spec_5x20<-prevalence.matching(l_spec, n.studies = 3, prev = 20)
length(spec_5x20)
spec_5x30<-prevalence.matching(l_spec, n.studies = 3, prev = 30)
length(spec_5x30)

```

# Filter the MGS based on presence in at least 3 studies and prevalence of at least 2.5%
```{r}
l.meta<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.rds"))
filtMGS<-spec_3x2.5
length(filtMGS) # number of MGS after prevalence filtration

l.meta.prevfilt<-lapply(l.meta, function(x){
  x.prev<-x %>%
  filter(bins%in%filtMGS)%>%
  mutate(p.adjust=p.adjust(pval.random, method = "fdr"))
  x.prev
})

saveRDS(l.meta.prevfilt, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.prevfilt.rds"))

```

# Subset sign bin and plot concordances (prevalence-filtered data)
The strategy below was taken from this post: https://stackoverflow.com/questions/53575257/make-binary-presence-absence-data-matrix-from-multiple-lists-in-r

```{r, message =F, warning =F}
l.meta.prevfilt.sign<-lapply(l.meta.prevfilt, function(x) subset(x,  p.adjust < 0.05))
# create the datafram with the bins
saveRDS(l.meta.prevfilt.sign, aste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.prevfilt.sign.rds"))

df.all.prevfilt<-do.call(rbind, l.meta.prevfilt.sign)
df.all.prevfilt$in20<-vector(length = nrow(df.all.prevfilt))
for(i in 1: nrow(df.all.prevfilt)){
  df.all.prevfilt$in20[i]<-ifelse(sum(df.all.prevfilt$bins == df.all.prevfilt$bins[i]) == 20, "TRUE", "FALSE")
}
df.all.prevfilt<-df.all.prevfilt[!duplicated(df.all.prevfilt$bins),] # select only duplicated, meaning one per mGS
tax.prevfilt.sign<-df.all.prevfilt[df.all.prevfilt$in20 == TRUE,]$bins

length(tax.prevfilt.sign)
nrow(df.all.prevfilt)-length(tax.prevfilt.sign)
```

# Select only sign taxa (prevalence-filtered data)
```{r}
l.meta.prevfilt.sign.20<-lapply(l.meta.prevfilt.sign, function(x) subset(x, bins %in% tax.prevfilt.sign))

saveRDS(l.meta.prevfilt.sign.20, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.prevfilt.sign.20.rds"))
```


# CD-HC / UC-HC / CD-UC: Filter the MGS based on presence in at least 3 studies and prevalence of at least 2.5%
```{r}
filtMGS<-spec_3x2.5
length(filtMGS) # number of MGS after prevalence filtration

l.meta.cdhc<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_cdcontrols.rds"))
l.meta.uchc<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_uccontrols.rds"))
l.meta.cduc<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta_uccd.rds"))

l.meta.cdhc.prevfilt<-lapply(l.meta.cdhc, function(x){
  x.prev<-x %>%
  filter(bins%in%filtMGS)%>%
  mutate(p.adjust=p.adjust(pval.random, method = "fdr"))
  x.prev
})

l.meta.uchc.prevfilt<-lapply(l.meta.uchc, function(x){
  x.prev<-x %>%
  filter(bins%in%filtMGS)%>%
  mutate(p.adjust=p.adjust(pval.random, method = "fdr"))
  x.prev
})

l.meta.cduc.prevfilt<-lapply(l.meta.cduc, function(x){
  x.prev<-x %>%
  filter(bins%in%filtMGS)%>%
  mutate(p.adjust=p.adjust(pval.random, method = "fdr"))
  x.prev
})

saveRDS(l.meta.cdhc.prevfilt, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.cdhc.prevfilt.rds"))
saveRDS(l.meta.uchc.prevfilt, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.uchc.prevfilt.rds"))
saveRDS(l.meta.cduc.prevfilt, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.cduc.prevfilt.rds"))
```

# Subset sign bin and plot concordances (prevalence-filtered data)

```{r, message =F, warning =F}
l.meta.cdhc.prevfilt.sign<-lapply(l.meta.cdhc.prevfilt, function(x) subset(x,  p.adjust < 0.05))
l.meta.uchc.prevfilt.sign<-lapply(l.meta.uchc.prevfilt, function(x) subset(x,  p.adjust < 0.05))
l.meta.cduc.prevfilt.sign<-lapply(l.meta.cduc.prevfilt, function(x) subset(x,  p.adjust < 0.05))
# create the datafram with the bins
saveRDS(l.meta.cdhc.prevfilt.sign, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.cdhc.prevfilt.sign.rds"))
saveRDS(l.meta.uchc.prevfilt.sign, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.uchc.prevfilt.sign.rds"))
saveRDS(l.meta.cduc.prevfilt.sign, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.cduc.prevfilt.sign.rds"))

df.all.cdhc.prevfilt<-do.call(rbind, l.meta.cdhc.prevfilt.sign)
df.all.cdhc.prevfilt$in20<-vector(length = nrow(df.all.cdhc.prevfilt))
for(i in 1: nrow(df.all.cdhc.prevfilt)){
  df.all.cdhc.prevfilt$in20[i]<-ifelse(sum(df.all.cdhc.prevfilt$bins == df.all.cdhc.prevfilt$bins[i]) == 20, "TRUE", "FALSE")
}
df.all.cdhc.prevfilt<-df.all.cdhc.prevfilt[!duplicated(df.all.cdhc.prevfilt$bins),] # select only duplicated, meaning one per mGS
tax.cdhc.prevfilt.sign<-df.all.cdhc.prevfilt[df.all.cdhc.prevfilt$in20 == TRUE,]$bins

df.all.uchc.prevfilt<-do.call(rbind, l.meta.uchc.prevfilt.sign)
df.all.uchc.prevfilt$in20<-vector(length = nrow(df.all.uchc.prevfilt))
for(i in 1: nrow(df.all.uchc.prevfilt)){
  df.all.uchc.prevfilt$in20[i]<-ifelse(sum(df.all.uchc.prevfilt$bins == df.all.uchc.prevfilt$bins[i]) == 20, "TRUE", "FALSE")
}
df.all.uchc.prevfilt<-df.all.uchc.prevfilt[!duplicated(df.all.uchc.prevfilt$bins),] # select only duplicated, meaning one per mGS
tax.uchc.prevfilt.sign<-df.all.uchc.prevfilt[df.all.uchc.prevfilt$in20 == TRUE,]$bins

df.all.cduc.prevfilt<-do.call(rbind, l.meta.cduc.prevfilt.sign)
df.all.cduc.prevfilt$in20<-vector(length = nrow(df.all.cduc.prevfilt))
for(i in 1: nrow(df.all.cduc.prevfilt)){
  df.all.cduc.prevfilt$in20[i]<-ifelse(sum(df.all.cduc.prevfilt$bins == df.all.cduc.prevfilt$bins[i]) == 20, "TRUE", "FALSE")
}
df.all.cduc.prevfilt<-df.all.cduc.prevfilt[!duplicated(df.all.cduc.prevfilt$bins),] # select only duplicated, meaning one per mGS
tax.cduc.prevfilt.sign<-df.all.cduc.prevfilt[df.all.cduc.prevfilt$in20 == TRUE,]$bins

####
length(tax.cdhc.prevfilt.sign)
nrow(df.all.cdhc.prevfilt)-length(tax.cdhc.prevfilt.sign)
nrow(df.all.cdhc.prevfilt)

length(tax.uchc.prevfilt.sign)
nrow(df.all.uchc.prevfilt)-length(tax.uchc.prevfilt.sign)

length(tax.cduc.prevfilt.sign)
nrow(df.all.cduc.prevfilt)-length(tax.cduc.prevfilt.sign)
```

# Select only sign taxa (prevalence-filtered data)
```{r}
l.meta.cdhc.prevfilt.sign.20<-lapply(l.meta.cdhc.prevfilt.sign, function(x) subset(x, bins %in% tax.cdhc.prevfilt.sign))
l.meta.uchc.prevfilt.sign.20<-lapply(l.meta.uchc.prevfilt.sign, function(x) subset(x, bins %in% tax.uchc.prevfilt.sign))
l.meta.cduc.prevfilt.sign.20<-lapply(l.meta.cduc.prevfilt.sign, function(x) subset(x, bins %in% tax.cduc.prevfilt.sign))

saveRDS(l.meta.cdhc.prevfilt.sign.20, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.cdhc.prevfilt.sign.20.rds"))
saveRDS(l.meta.uchc.prevfilt.sign.20, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.uchc.prevfilt.sign.20.rds"))
saveRDS(l.meta.cduc.prevfilt.sign.20, paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.cduc.prevfilt.sign.20.rds"))
```
