---
title: "00_resampling"
author: "Rebecca Ansorge"
date: "09/12/2023"
output: html_document
---



# Install load packages
```{r setup, include=FALSE}

list.of.packages <- c("dplyr", "reshape2", "propr", "meta", "ggplot2", "lattice", "DHARMa", "broomExtra")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq", "microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)

library(meta)
library(DHARMa)
library(nlme)
library(ggplot2)
library(phyloseq)
library(dplyr)
library(microbiome)
library(reshape2)
library(meta)
library(lattice)
library(tidyverse)
#library(broomExtra)

```

```{r}
PATHanalNov2023<-"~/Documents/Rebecca/onedrive/analysisNov2023/"
```

# Load and format the metadata

```{r, message =F, warning =F}
meta<-read.delim(paste0(PATHanalNov2023,"PrepData/IBDmeta.cur.filt.mc.matched.txt"), header = T, sep = "\t")
dim(meta)
sapply(meta, class)

meta$IBD<-factor(meta$IBD, levels = c("healthy", "IBD"))
meta$IBD2<-factor(meta$IBD2, levels = c("healthy", "UC", "CD"))

meta.summary<-read.delim(paste0(PATHanalNov2023,"PrepData/IBDmeta.cur.filt.mc.overview.txt"), header = T, sep = "\t")


```

## Chec if all samples hava the right metadata

Check if the cohort is specified for all
```{r, message =F, warning =F}
meta$Cohort_precise %>% unique()
meta$Cohort %>% unique()

```

All have both cohort and cohort precise

## Check all samples have IBD assigned
```{r, message =F, warning =F}
meta$IBD %>% unique()
meta$IBD2 %>% unique()

```
In IBD2 there is IBDU as I converted them into factors I ignored this entry and is treated as an NA now

## Check all cohort have case_controls
```{r, message =F, warning =F}
meta %>%
  dplyr::group_by(Cohort_precise) %>%
  summarise(unique(IBD)) %>%
  print(n = 40)

```

# Split data in cross-sectiona and long
```{r, message =F, warning =F}
# remove cohort with only IBD
meta<-subset(meta, Cohort_precise != "Fang_UCSD_IBD_Biobank" & 
                Cohort_precise != "USA")


meta.cross<-subset(meta, Cohort_setup == "crosssectional")
meta.long<-subset(meta, Cohort_setup == "longitudinal")

```

## Longitudinal samples
# Explore the repeated samples
```{r, message=F, warning =F}
meta.long$visit_number<-as.factor(meta.long$days_after_first_visit)
meta.long$dummy<-rep(1, nrow(meta.long))

meta.long %>%
  dplyr::group_by(Cohort_precise, IndividualID) %>%
  dplyr::summarise(n = sum(dummy)) %>% 
  print(n=40) %>%
  ggplot(aes(Cohort_precise, n, fill = Cohort_precise)) +
  geom_boxplot(alpha = 0.7) + 
  scale_fill_brewer(type = "d") +
  geom_jitter(width = 0.4, alpha = 0.7, aes(fill = Cohort_precise), size = 4, pch = 21) +
  scale_color_brewer(type = "d") +
  ggtitle("Number of samples per participant across cohorts")
```

The number fo samples per single person is very variable, form 1 to 25.
the mx number of samples per individual is 25.
I will re sample up to 20 time the data

```{r, message=FALSE, warning=FALSE, error=FALSE}
#orde the data randomly and then select only a single sample
set.seed(342)
# I can split the metadata in cohort and then in Individuals and then get indices of each 

meta.long.cohort<-split(meta.long, f = meta.long$Cohort_precise)
meta.long.cohort.ind<-lapply(meta.long.cohort, function(x) split(x, f = x$IndividualID))

# createa a df with all ids that I will use
df.row<-data.frame("Sample1", "Sample2", "Sample3", "Sample4", "Sample5", "Sample6",
                     "Sample7", "Sample8", "Sample9", "Sample10", "Sample11",
                     "Sample12", "Sample13", "Sample14", "Sample15", "Sample16", "Sample17",
                     "Sample18", "Sample19", "Sample20",  "IndividualID", "Cohort")
for(i in 1:length(meta.long.cohort.ind)) {

  for(id in 1:length(meta.long.cohort.ind[[i]])){
    id.df<-meta.long.cohort.ind[[i]][[id]]
    rows<-id.df$SmplID
    if(length(rows) < 20){
      rows1<-sample(rows, size = 20,replace = T )
    } else {rows1<-sample(rows, size = 20,replace = F )}
    rows2<-c(rows1, unique(id.df$IndividualID), names(meta.long.cohort.ind)[i])
    df.row<-rbind.data.frame(df.row,rows2)
  }
  df.row<-as.data.frame(df.row)
}
names(df.row)<-df.row[1,]
df.row<-df.row[-1,] # because the first row is names

# check that the number of samples is the same as unique IDs
nrow(df.row) == length(unique(meta.long$IndividualID))
# I think that the reason why they are not unique is that there are similar ids in different studies
write_delim(df.row,paste0(PATHanalNov2023,"PrepData/resampling.20.txt"),delim="\t")

```


# Load and format the metadata

```{r, message =F, warning =F}
meta<-read.delim(paste0(PATHanalNov2023,"PrepData/IBDmeta.cur.filt.mc.matched.txt"), header = T, sep = "\t")
dim(meta)
sapply(meta, class)

meta$IBD<-factor(meta$IBD, levels = c("healthy", "IBD"))
meta$IBD2<-factor(meta$IBD2, levels = c("healthy", "UC", "CD"))

```

# Split data in cross-sectiona and long
```{r, message =F, warning =F}
meta.cross<-subset(meta, Cohort_setup == "crosssectional")
meta.long<-subset(meta, Cohort_setup == "longitudinal")

```


```{r, message=FALSE, warning=FALSE, error=FALSE}
#orde the data randomly and then select only a single sample
set.seed(342)
# I can split the metadata in cohort and then in Individuals and then get indices of each 

meta.long.cohort<-split(meta.long, f = meta.long$Cohort_precise)
meta.long.cohort.ind<-lapply(meta.long.cohort, function(x) split(x, f = x$IndividualID))

# createa a df with all ids that I will use
df.row<-data.frame("Sample1", "Sample2", "Sample3", "Sample4", "Sample5", "Sample6",
                     "Sample7", "Sample8", "Sample9", "Sample10", "Sample11",
                     "Sample12", "Sample13", "Sample14", "Sample15", "Sample16", "Sample17",
                     "Sample18", "Sample19", "Sample20",  "IndividualID", "Cohort")
for(i in 1:length(meta.long.cohort.ind)) {

  for(id in 1:length(meta.long.cohort.ind[[i]])){
    id.df<-meta.long.cohort.ind[[i]][[id]]
    rows<-id.df$SmplID
    if(length(rows) < 20){
      rows1<-sample(rows, size = 20,replace = T )
    } else {rows1<-sample(rows, size = 20,replace = F )}
    rows2<-c(rows1, unique(id.df$IndividualID), names(meta.long.cohort.ind)[i])
    df.row<-rbind.data.frame(df.row,rows2)
  }
  df.row<-as.data.frame(df.row)
}
names(df.row)<-df.row[1,]
df.row<-df.row[-1,] # because the first row is names

# check that the number of samples is the same as unique IDs
nrow(df.row) == length(unique(meta.long$IndividualID))
# I think that the reason why they are not unique is that there are similar ids in different studies
write_delim(df.row,paste0(PATHanalNov2023,"PrepData/resampling.CDUC.20.txt"),delim="\t")
```


