---
title: "strain_pmva"
author: "Rebecca Ansorge"
date: "13/06/2023"
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Correlate clustering of strains into clades according to disease state with species-level differential abundance effect sizes

*TODO*: For UC,CD,HC include numbers (even if not linking with DA)

```{r include = FALSE}
knitr::opts_chunk$set(echo=FALSE)
```

# import libraries
```{r include=FALSE}
library(tidyverse)
library(ggpubr)
library(ggtree)
library(ggtreeExtra)
library(ggnewscale)
library(ape)
library(ggVennDiagram)
```

# import and reshape data

```{r}
#PATHanalNov2023<-"~/Documents/Rebecca/onedrive/analysisNov2023/"
#PATHanalNov2023<-"~/Library/CloudStorage/OneDrive-Personal/analysisNov2023/"
PATHanalNov2023<-"~/Documents/Rebecca/onedrive/analysisNov2023/"
qval<-0.1
```

## strains
```{r include=FALSE}
# import prevalence filtered MGS IDs
prevMGS.tab<-read_csv(paste0(PATHanalNov2023,"Downstream/Results/prev.filt.MGS.txt"))%>%
  mutate(prevMGS=gsub('.*;MGS','MGS',x))
prevMGS<-prevMGS.tab$prevMGS

# import strain results into list
#strainpath="/Volumes/projects/c/cfa5d1aa-0df6-49de-ade6-c8334c54f9d1/data/results/IBD/strains/IBD8_Bin5SB/strainsIBD_pmva_out_Jan24/"
strainpath="/Users/ansorge/Downloads/strainsIBD_pmva_out_Jan24/"
## IBD
ibd.strains.files = list.files(path = strainpath,pattern="Strain_pmva_cndCohort_ctrIBD.sample*")
ibd.strains.pmva<-lapply(ibd.strains.files, function(x){
  read_delim(paste0(strainpath,x), delim="\t")%>%
  filter(var%in%prevMGS) %>% # only keep the prevalence-filtered MGS
  # adjust p-value for multiple testing using Benjamini-Hochberg
  mutate(padjustStrains=p.adjust(`Pr(>F)`,method = "BH"))%>%
  mutate(significance=ifelse(padjustStrains<0.1,"significant","non-significant"))
  })
ibd.strains.pmva.noPrevFilt<-lapply(ibd.strains.files, function(x){
  read_delim(paste0(strainpath,x), delim="\t")%>%
  #filter(var%in%prevMGS) %>% # only keep the prevalence-filtered MGS
  # adjust p-value for multiple testing using Benjamini-Hochberg
  mutate(padjustStrains=p.adjust(`Pr(>F)`,method = "BH"))%>%
  mutate(significance=ifelse(padjustStrains<0.1,"significant","non-significant"))
  })

# numbers before and after p-value correction
# make df
strainNum.list<-lapply(1:length(ibd.strains.pmva), function(n) {
  p0.05<-ibd.strains.pmva[[n]]%>%filter(`Pr(>F)`<0.05)%>%nrow()
  p0.05_percentPrevFilt<-(ibd.strains.pmva[[n]]%>%filter(`Pr(>F)`<0.05)%>%nrow())/nrow(ibd.strains.pmva[[n]])
  q0.1<-ibd.strains.pmva[[n]]%>%filter(padjustStrains<0.1)%>%nrow()
  q0.1_percentPrevFilt<-(ibd.strains.pmva[[n]]%>%filter(padjustStrains<0.1)%>%nrow())/nrow(ibd.strains.pmva[[n]])
  
  strainNum<-data.frame(
    sample=paste0("sample",n),
    p05=p0.05,
    p05_percentPrevFilt=(p0.05_percentPrevFilt*100),
    q01=q0.1,
    q01_percentPrevFilt=(q0.1_percentPrevFilt*100)
  )
  strainNum
}
)
strainNum.df<-do.call(rbind, strainNum.list)

# numbers before and after p-value correction without prev filtration
# make df
strainNum.list2<-lapply(1:length(ibd.strains.pmva.noPrevFilt), function(n) {
  p0.05<-ibd.strains.pmva.noPrevFilt[[n]]%>%filter(`Pr(>F)`<0.05)%>%nrow()
  p0.05_percentPrevFilt<-(ibd.strains.pmva.noPrevFilt[[n]]%>%filter(`Pr(>F)`<0.05)%>%nrow())/nrow(ibd.strains.pmva.noPrevFilt[[n]])
  q0.1<-ibd.strains.pmva.noPrevFilt[[n]]%>%filter(padjustStrains<0.1)%>%nrow()
  q0.1_percentPrevFilt<-(ibd.strains.pmva.noPrevFilt[[n]]%>%filter(padjustStrains<0.1)%>%nrow())/nrow(ibd.strains.pmva.noPrevFilt[[n]])
  
  strainNum<-data.frame(
    sample=paste0("sample",n),
    p05=p0.05,
    p05_percentPrevFilt=(p0.05_percentPrevFilt*100),
    q01=q0.1,
    q01_percentPrevFilt=(q0.1_percentPrevFilt*100)
  )
  strainNum
}
)
strainNum.df2<-do.call(rbind, strainNum.list2)

# check robustness 
df.ibd.strains.pmva<-do.call(rbind, ibd.strains.pmva)
count.ibd<-df.ibd.strains.pmva%>%
  filter(significance=="significant")%>%
  count(var)%>%
  rename("n_sig"=n)
count20.ibd<-count.ibd%>%
  filter(n_sig==20)
countLess20.ibd<-count.ibd%>%
  filter(n_sig<20)
robustMGSstrains<-count20.ibd$var
variableMGSstrains<-countLess20.ibd$var

# add info of robustness to all strain results
ibd.strains.pmva.all<-lapply(ibd.strains.pmva, function(m){
  m %>%
    mutate(robustStrains=ifelse(var%in%robustMGSstrains,"robust",ifelse(var%in%variableMGSstrains,"variable","non-sig")))%>%
    left_join(count.ibd, by="var")%>%
    mutate(n_sig=ifelse(robustStrains=="non-sig",0,n_sig))
})


## IBD2
ibd2.strains.files = list.files(path = strainpath,pattern="Strain_pmva_cndCohort_ctrIBD2.sample*")
ibd2.strains.pmva<-lapply(ibd2.strains.files, function(x){
  read_delim(paste0(strainpath,x), delim="\t")%>%
  filter(var%in%prevMGS) %>% # only keep the prevalence-filtered MGS
  # adjust p-value for multiple testing using Benjamini-Hochberg
  mutate(padjustStrains=p.adjust(`Pr(>F)`,method = "BH"))%>%
  mutate(significance=ifelse(padjustStrains<0.1,"significant","non-significant"))
  })

# check robustness 
df.ibd2.strains.pmva<-do.call(rbind, ibd2.strains.pmva)
count.ibd2<-df.ibd2.strains.pmva%>%
  filter(significance=="significant")%>%
  count(var)%>%
  rename("n_sig"=n)
count20.ibd2<-count.ibd2%>%
  filter(n_sig==20)
countLess20.ibd2<-count.ibd2%>%
  filter(n_sig<20)
robustMGSstrains2<-count20.ibd2$var
variableMGSstrains2<-countLess20.ibd2$var

# add info of robustness to all strain results
ibd2.strains.pmva.all<-lapply(ibd2.strains.pmva, function(m){
  m %>%
    mutate(robustStrains=ifelse(var%in%robustMGSstrains2,"robust",ifelse(var%in%variableMGSstrains2,"variable","non-sig")))%>%
    left_join(count.ibd2, by="var")%>%
    mutate(n_sig=ifelse(robustStrains=="non-sig",0,n_sig))
})

## IBD2 pairs
# ibd2.pair.strains.files = list.files(path = strainpath,pattern="Strain_pmva_cndCohort_ctrIBD2pairs.sample*")
# ibd2.pair.strains.pmva<-lapply(ibd2.pair.strains.files, function(x){
#   read.delim(paste0(strainpath,x), sep="\t")%>%
#   filter(var%in%prevMGS) %>% # only keep the prevalence-filtered MGS
#   # adjust p-value for multiple testing using Benjamini-Hochberg
#   mutate(padjustStrains=p.adjust(`Pr..F.`,method = "BH"))%>%
#   mutate(significance=ifelse(padjustStrains<0.1,"significant","non-significant"))
#   })
# 
# # check robustness 
# df.ibd2.pair.strains.pmva<-do.call(rbind, ibd2.pair.strains.pmva)
# count20.ibd2.pair<-df.ibd2.pair.strains.pmva%>%
#   filter(significance=="significant")%>%
#   count(var)%>%
#   filter(n==20)
# robustMGSstrains2.pair<-count20.ibd2.pair$var
# 
# # add info of robustness to all strain results
# ibd2.pair.strains.pmva.all<-lapply(ibd2.pair.strains.pmva, function(m){
#   m %>%
#   mutate(robustStrains=ifelse(var%in%robustMGSstrains2.pair,"robust","variable"))
# })

ibd.strains.pmva.all<-readRDS(paste0(PATHanalNov2023,"strains/RDS/ibd.strains.pmva.all.June23.rds"))
ibd2.strains.pmva.all<-readRDS(paste0(PATHanalNov2023,"strains/RDS/ibd2.strains.pmva.all.June23.rds"))
```

## DA
```{r}
# all species DA
metaDA<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.prevfilt.rds"))
# robust species DA
#metaDArobust<-readRDS(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/l.meta.prevfilt.sign0.1.20.rds"))
#robustMGSda<-metaDArobust[[1]][,"bins"] # get IDs of robust MGS

df.ibd.DA<-do.call(rbind, metaDA)
count.DA<-df.ibd.DA%>%
  mutate(significance=ifelse(p.adjust<qval,"significant","non-significant"))%>%
  filter(significance=="significant")%>%
  count(bins)%>%
  rename("n_sig"=n)
count20.DA<-count.DA%>%
  filter(n_sig==20)
countLess20.DA<-count.DA%>%
  filter(n_sig<20)
robustMGSDA<-count20.DA$bins
variableMGSDA<-countLess20.DA$bins

# add info of robustness to all DA results
metaDAall<-lapply(metaDA, function(m){
  m %>%
    mutate(robustDA=ifelse(bins%in%robustMGSDA,"robust",ifelse(bins%in%variableMGSDA,"variable","non-sig")))%>%
    left_join(count.DA, by="bins")%>%
    mutate(n_sig=ifelse(robustDA=="non-sig",0,n_sig))
})
```

## metadata and trees
```{r}

ibdMeta<-read.delim2(paste0(PATHanalNov2023,"PrepData/IBDmeta.cur.filt.mc.matched.txt"), sep="\t") %>%
  mutate(row=SmplID) %>%
  column_to_rownames("row")

spTree<-read.tree(paste0(PATHanalNov2023,"PrepData/IQtree_allsites.treefile"))

# strain trees location
treepath=paste0(PATHanalNov2023,"PrepData/strainTrees/")

```

# summarize strain grouping numbers
```{r}
table(ibd.strains.pmva.all[[1]]$robustStrains)
table(ibd2.strains.pmva.all[[1]]$robustStrains)
```

```{r}

ibdRes1<-do.call(rbind, ibd.strains.pmva.all)
ibdRes1<-ibdRes1%>%
  group_by(var)%>%
  mutate(meanR2=mean(R2))%>%
  mutate(stdevR2=sd(R2))%>%
  select(c(var,meanR2,stdevR2,robustStrains,n_sig))%>%
  distinct()

ibd.strains.p1<-ggplot(ibdRes1%>%filter(robustStrains=="robust"), aes(x=meanR2,y=var))+
  geom_point(alpha=0.7)+
  geom_errorbar(aes(xmin=meanR2-stdevR2, xmax=meanR2+stdevR2), width=.2)+
  ggtitle("robust sign. strain grouping mean R2 acc. to IBD vs HC")
ibd.strains.p1

ibdRes2<-do.call(rbind, ibd2.strains.pmva.all)
ibdRes2<-ibdRes2%>%
  group_by(var)%>%
  mutate(meanR2=mean(R2))%>%
  mutate(stdevR2=sd(R2))%>%
  select(c(var,meanR2,stdevR2,robustStrains,n_sig))%>%
  distinct()

ibd2.strains.p1<-ggplot(ibdRes2%>%filter(robustStrains=="robust"), aes(x=meanR2,y=var))+
  geom_point(alpha=0.7)+
  geom_errorbar(aes(xmin=meanR2-stdevR2, xmax=meanR2+stdevR2), width=.2)+
  ggtitle("robust sign. strain grouping mean R2 to CD vs UC vs HC")
ibd2.strains.p1

```

# overlap of MGS between IBD and IBD2
```{r}
x <- list(
  ibd = (ibdRes1%>%filter(robustStrains=="robust"))$var, 
  ibd2 = (ibdRes2%>%filter(robustStrains=="robust"))$var)

ggVennDiagram(x, label_alpha = 0, edge_size =0)+
  scale_fill_gradient(low="white",high = "salmon")+
  ggtitle("robust significant strain grouping")

y <- list(
  ibd = (ibdRes1%>%filter(robustStrains=="variable"))$var, 
  ibd2 = (ibdRes2%>%filter(robustStrains=="variable"))$var)

ggVennDiagram(y, label_alpha = 0, edge_size =0)+
  scale_fill_gradient(low="white",high = "salmon")+
  ggtitle("variable significant strain grouping")

#saveRDS(x,paste0(PATHanalNov2023,"strains/RDS/StrainClust.overlapIBD_IBD2.June23.rds"))
#saveRDS(y,paste0(PATHanalNov2023,"strains/RDS/StrainClust.overlapIBD_IBD2.variable.June23.rds"))

```

# plot strain clade R2 (sig. and non-sig.) vs DA effect size

no filtration of clustering data

```{r}
# consider removing low-prevalence ones to match with DA

# create df with mean effect sizes and R2
STRAINmeans<-ibdRes1%>%
  rename(MGS=var)

DAmeans<-do.call(rbind, metaDAall)%>%
  mutate(MGS=gsub(".*;MGS","MGS",bins))%>%
  mutate(TE.random=as.numeric(TE.random))
DAmeans<-DAmeans%>%
  group_by(MGS)%>%
  mutate(meanTE.random=mean(TE.random))%>%
  mutate(stdevTE.random=sd(TE.random))%>%
  select(c(bins,meanTE.random,stdevTE.random,robustDA))%>%
  distinct()

strainR2vsTE<-STRAINmeans %>%
  full_join(DAmeans, by="MGS") %>%
  mutate(abs.meanTE.random=abs(meanTE.random))%>%
  #mutate(significance=ifelse(robustStrains=="robust" & robustDA=="robust","both sig",ifelse(robustStrains=="robust" & robustDA!="robust","strains sig",ifelse(robustDA=="robust" & robustStrains!="robust","DA sig","not sig"))))%>%
  mutate(significance=ifelse(robustStrains!="non-sig" & robustDA!="non-sig","both sig",ifelse(robustStrains!="non-sig" & robustDA=="non-sig","strains sig",ifelse(robustDA!="non-sig" & robustStrains=="non-sig","DA sig","not sig"))))%>%
  mutate(robustness=paste0("strains_",robustStrains," DA_",robustDA))%>%
  filter(significance!="not sig") # remove MGS not significant in any

strainR2vsTE.inclNonSig<-STRAINmeans %>%
  full_join(DAmeans, by="MGS") %>%
  mutate(abs.meanTE.random=abs(meanTE.random))%>%
  #mutate(significance=ifelse(robustStrains=="robust" & robustDA=="robust","both sig",ifelse(robustStrains=="robust" & robustDA!="robust","strains sig",ifelse(robustDA=="robust" & robustStrains!="robust","DA sig","not sig"))))
  mutate(significance=ifelse(robustStrains!="non-sig" & robustDA!="non-sig","both sig",ifelse(robustStrains!="non-sig" & robustDA=="non-sig","strains sig",ifelse(robustDA!="non-sig" & robustStrains=="non-sig","DA sig","not sig"))))%>%
  mutate(robustness=paste0("strains_",robustStrains," DA_",robustDA))

#saveRDS(strainR2vsTE.inclNonSig,paste0(PATHanalNov2023,"strains/RDS/strainR2vsTE.inclNonSig.June23.rds"))
strainR2vsTE.inclNonSig<-readRDS(paste0(PATHanalNov2023,"strains/RDS/strainR2vsTE.inclNonSig.June23.rds"))

# swap over too DA q<0.1
strainR2vsTE.inclNonSig_q0.1<-strainR2vsTE.inclNonSig%>%
  select(colnames(strainR2vsTE.inclNonSig)[1:5])%>%
  full_join(DAmeans, by="MGS") %>%
  mutate(abs.meanTE.random=abs(meanTE.random))%>%
  mutate(significance=ifelse(robustStrains!="non-sig" & robustDA!="non-sig","both sig",ifelse(robustStrains!="non-sig" & robustDA=="non-sig","strains sig",ifelse(robustDA!="non-sig" & robustStrains=="non-sig","DA sig","not sig"))))%>%
  mutate(robustness=paste0("strains_",robustStrains," DA_",robustDA))

strainR2vsTE.inclNonSig_q0.1%>%
  group_by(robustDA)%>%
  count()

# only significant ones
strainR2vsTE_q0.1<-strainR2vsTE.inclNonSig_q0.1%>%
  mutate(significance=ifelse(robustStrains!="non-sig" & robustDA!="non-sig","both sig",ifelse(robustStrains!="non-sig" & robustDA=="non-sig","strains sig",ifelse(robustDA!="non-sig" & robustStrains=="non-sig","DA sig","not sig"))))%>%
  mutate(robustness=paste0("strains_",robustStrains," DA_",robustDA))%>%
  filter(significance!="not sig") # remove MGS not significant in any

strainR2vsTE<-strainR2vsTE_q0.1
strainR2vsTE.inclNonSig<-strainR2vsTE.inclNonSig_q0.1

saveRDS(strainR2vsTE.inclNonSig,paste0(PATHanalNov2023,"strains/RDS/strainR2vsTE.inclNonSig_q0.1.June23.rds"))
```

```{r}
# create df with mean effect sizes and R2
STRAINmeans_ibd<-ibdRes1%>%
  rename(MGS=var)%>%
  rename(meanR2_IBD=meanR2)%>%
  rename(stdevR2_IBD=stdevR2)%>%
  rename(robustStrains_IBD=robustStrains)%>%
  rename(n_sig_IBD=n_sig)
STRAINmeans_ibd2<-ibdRes2%>%
  rename(MGS=var)%>%
  rename(meanR2_CDUCCTRL=meanR2)%>%
  rename(stdevR2_CDUCCTRL=stdevR2)%>%
  rename(robustStrains_CDUCCTRL=robustStrains)%>%
  rename(n_sig_CDUCCTRL=n_sig)

DAmeans_ibd<-do.call(rbind, metaDAall)%>%
  mutate(MGS=gsub(".*;MGS","MGS",bins))%>%
  mutate(TE.random=as.numeric(TE.random))
DAmeans_ibd<-DAmeans_ibd%>%
  group_by(MGS)%>%
  mutate(meanTE.random=mean(TE.random))%>%
  mutate(stdevTE.random=sd(TE.random))%>%
  select(c(bins,meanTE.random,stdevTE.random,robustDA,n_sig))%>%
  distinct()%>%
  rename(n_sig_DA=n_sig)

strainR2vsTE.inclNonSig.ibd_ibd2<-STRAINmeans_ibd %>%
  full_join(STRAINmeans_ibd2, by="MGS") %>%
  full_join(DAmeans_ibd, by="MGS") %>%
  mutate(abs.meanTE.random=abs(meanTE.random))

#saveRDS(strainR2vsTE.inclNonSig.ibd_ibd2,paste0(PATHanalNov2023,"strains/RDS/strainR2vsTE.inclNonSig.ibd_ibd2.June23.rds"))
strainR2vsTE.inclNonSig.ibd_ibd2<-readRDS(paste0(PATHanalNov2023,"strains/RDS/strainR2vsTE.inclNonSig.ibd_ibd2.June23.rds"))

strainR2vsTE.inclNonSig.ibd_ibd2_q0.1<-strainR2vsTE.inclNonSig.ibd_ibd2%>%
  select(1:9)%>%
  full_join(DAmeans_ibd, by="MGS") %>%
  mutate(abs.meanTE.random=abs(meanTE.random))

strainR2vsTE.inclNonSig.ibd_ibd2<-strainR2vsTE.inclNonSig.ibd_ibd2_q0.1

saveRDS(strainR2vsTE.inclNonSig.ibd_ibd2,paste0(PATHanalNov2023,"strains/RDS/strainR2vsTE.inclNonSig_q0.1.ibd_ibd2.June23.rds"))

```

# tree stats
```{r}
# import all trees into list

treepath=paste0("~/Downloads/strainsIBD_pmva_out_Jan24/")

allfiles = list.files(path = treepath,pattern="*.pmvastrains.20iterations.RData")
tree.list<-lapply(allfiles, function(x){
  try(load(paste0(treepath,x)))
  #inTreeSub2_MP<-midpoint(inTreeSub2)
  inTreeSub2})
names(tree.list)<-allfiles # name with MGS

# number of tips for each tree
tip.counts<-lapply(tree.list, function(x) {
  tipnum<-length(x$tip.label)-1
  tipnum
})


# list with sample IDs per IBD status group
mgs.list<-lapply(c("IBD","healthy","UC","CD"), function(x) {
  if(x%in%c("IBD","healthy")){
    mgs<-ibdMeta[ibdMeta$IBD==x,]$SmplID
    mgs
  } else if(x%in%c("UC","CD")){
    mgs<-ibdMeta[ibdMeta$IBD2==x,]$SmplID
    mgs
  }
})
names(mgs.list)<-c("IBD","healthy","UC","CD")


# function to extract number of tips for each bin x per disease condition y
get_strat_tips<-function(x,y){
  # x is bin
  # y is either IBD, healthy, UC, CD
  stratTips<-sum(tree.list[[x]]$tip.label%in% mgs.list[[y]])
  return(stratTips)
  }

# lapply to count tips for each condition in every tree
numTips<-lapply(c("IBD","healthy","CD","UC"), function(stat) {
  xTips<-lapply(names(tree.list),function(bin) {get_strat_tips(bin,stat)})
  names(xTips)<-names(tree.list)
  xTips
})
names(numTips)<-c("IBD","healthy","CD","UC")


# combine tip counts in df
tree.df<-data.frame(Trees = names(tree.list)) 

tree.df$Tips<-do.call(cbind.data.frame, tip.counts) %>% t() 
tree.df$IBD<-do.call(cbind.data.frame, numTips$IBD) %>% t()
tree.df$healthy<-do.call(cbind.data.frame, numTips$healthy) %>% t()
tree.df$CD<-do.call(cbind.data.frame, numTips$CD) %>% t()
tree.df$UC<-do.call(cbind.data.frame, numTips$UC) %>% t()
tree.df$MGS<-gsub(".treefile","",tree.df$Trees)
# preview tree.df
head(tree.df)

tree.df.prev<-tree.df%>%
  mutate(MGS=gsub(".pmvastrains.20iterations.RData","",MGS))%>%
  filter(MGS%in%prevMGS)

tree.df.prev<-readRDS(paste0(PATHanalNov2023,"strains/RDS/tree.df.prev.June23.rds"))
```

# merge with tips
```{r}
strainR2vsTE.incTips<-strainR2vsTE %>%
  left_join(tree.df.prev, by="MGS")
strainR2vsTE.incTips

saveRDS(strainR2vsTE.incTips,paste0(PATHanalNov2023,"strains/RDS/strainR2vsTE.incTips_q0.1.June23.rds"))

strainR2vsTE.incTips
```

# LM 
```{r}

# run linear model with number of tips and DA effect size
lm.out<-lm(meanR2 ~ Tips + abs(meanTE.random), data=strainR2vsTE.incTips)
summary(lm.out)
lm.out.log2<-lm(log2(meanR2) ~ Tips + abs(meanTE.random), data=strainR2vsTE.incTips)
summary(lm.out.log2)
  
# run linear model with number of tips and DA effect size only significant ones in both
lm.out.sig<-lm(meanR2 ~ Tips + abs(meanTE.random), data=strainR2vsTE.incTips%>%filter(significance=="both sig"))
summary(lm.out.sig)
lm.out.sig.log<-lm(log2(meanR2) ~ Tips + abs(meanTE.random), data=strainR2vsTE.incTips%>%filter(significance=="both sig"))
summary(lm.out.sig.log)

lm.list<-list(anysig=summary(lm.out),anysigLog2=summary(lm.out.log2),bothsig=summary(lm.out.sig),bothsigLog2=summary(lm.out.sig.log))
saveRDS(lm.list,paste0(PATHanalNov2023,"strains/RDS/strainR2vsDA.lm.June23_q0.1.rds"))
```
The linear model shows a negative correlation between mean R2 (and log2 of mean) in strain clades compared to mean effect size of species differential abundance after accounting for number of tips. This was done on prevalence-filtered, robust MGS only.

# plot DA effect sizes against strain R2 (only robust) 
```{r}
strainR2vsTE.p1<-ggplot(strainR2vsTE) +
    geom_errorbar(aes(x=log2(meanR2),y=abs(meanTE.random),xmin=log2(meanR2-stdevR2), xmax=log2(meanR2+stdevR2)), width=.1, color='grey')+
    geom_errorbar(aes(x=log2(meanR2),y=abs(meanTE.random),ymin=abs(meanTE.random)-stdevTE.random, ymax=abs(meanTE.random)+stdevTE.random), width=.1, color='grey')+
    geom_smooth(aes(x=log2(meanR2),y=abs(meanTE.random)),method='lm', color="black", size=0.5, alpha=0.2) +
    geom_point(aes(x=log2(meanR2),y=abs(meanTE.random),color=significance), alpha=0.7) +
    #stat_cor(aes(x=log2(meanR2),y=abs(meanTE.random)),label.x = -10, label.y = 1,size=3) +
    theme_bw() +
    ggtitle("DA effect sizes VS R2 of permanova")+
    ylab("DA effect size") +
    xlab("log2(mean R2 of strain clustering IBD vs. healthy)") +
    theme(title = element_text(size=10),
        #legend.position = "top",
        legend.title = element_text(size = 10), 
        legend.text = element_text(size = 10),
        legend.key = element_rect(fill = "white", colour = NA),
        legend.key.size = unit(0.1, "cm"),
        legend.spacing.y = unit(0.1, "cm"))
strainR2vsTE.p1
ggsave(paste0(PATHanalNov2023,"strains/strainR2vsTE.eachsig_q0.1.June23.pdf"),strainR2vsTE.p1,height=5,width=6)
```

# plot significant mean DA effect size against significant mean R2 only
```{r}
strainR2vsTE.p2<-ggplot(strainR2vsTE%>%filter(significance=="both sig")) +
    geom_errorbar(aes(x=log2(meanR2),y=abs(meanTE.random),xmin=log2(meanR2-stdevR2), xmax=log2(meanR2+stdevR2)), width=.1, color='grey')+
    geom_errorbar(aes(x=log2(meanR2),y=abs(meanTE.random),ymin=abs(meanTE.random)-stdevTE.random, ymax=abs(meanTE.random)+stdevTE.random), width=.1, color='grey')+
    geom_smooth(aes(x=log2(meanR2),y=abs(meanTE.random)),method='lm', color="black", size=0.5, alpha=0.2) +
    geom_point(aes(x=log2(meanR2),y=abs(meanTE.random)), alpha=0.7) +
    #stat_cor(aes(x=log2(meanR2),y=abs(meanTE.random)),label.x = -10, label.y = 1,size=3) +
    theme_bw() +
    ggtitle("DA effect sizes VS R2 of permanova (MGS sig in both) ")+
    ylab("DA effect size") +
    xlab("log2(mean R2 of strain clustering IBD vs. healthy)") +
    theme(title = element_text(size=10),
        #legend.position = "top",
        legend.title = element_text(size = 10), 
        legend.text = element_text(size = 10),
        legend.key = element_rect(fill = "white", colour = NA),
        legend.key.size = unit(0.1, "cm"),
        legend.spacing.y = unit(0.1, "cm"))
strainR2vsTE.p2
ggsave(paste0(PATHanalNov2023,"strains/strainR2vsTE.bothsig_q0.1.June23.pdf"),strainR2vsTE.p2,height=5,width=5)
```

# confusion matrix
```{r}
strainR2vsTE.inclNonSig.counts<-strainR2vsTE.inclNonSig%>%
  group_by(significance)%>%
  count()
strainR2vsTE.inclNonSig.counts

```


