---
title: "DA_IBD_LM"
author: "Stefano Romano"
date: "May 15, 2022"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: inline
---
# Install load packages
```{r setup, include=FALSE}

list.of.packages <- c("dplyr", "reshape2", "ggplot2", "lattice")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq","microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)


library(ggplot2)
library(phyloseq)
library(dplyr)
library(microbiome)
library(reshape2)
library(tidyverse)
library(RColorBrewer)
library(cowplot)
library(vegan)
library(MatchIt)

```

# Load and clean data

```{r}
PATHanalNov2023<-"~/Documents/Rebecca/onedrive/analysisNov2023/"
```

```{r}
#colors
col_IBD2<-c("gold2","#E31A1C","darkorange","#1F78B4","#1F78B4","#1F78B4", "grey40","#E8A400")
names(col_IBD2)<-c("IBD","CD","UC","healthy","control", "Healthy","variable","disease")

```

Load the abundance file and the metadata

```{r, message=F}

spec<-read.delim(paste0(PATHanalNov2023,"PrepData/IBD.MGSabundance.txt"), header = T, row.names = 1)
dim(spec)

rownames(spec) %>% tail
```

The last row is "other". This will be used to calculate the relative abundances but will then be removed
```{r, message =F, warning =F}

spec.tss<-as.data.frame(prop.table(as.matrix(spec), 2))


# now remove other
spec.tss<-spec.tss[-1*which(rownames(spec.tss) == "other"),]
dim(spec.tss)


# check NAs
spec.na<-apply(spec,2, function(x) is.na(x))
all(spec.na == FALSE) 
xl<-apply(is.na(spec.tss), 2, which)
na.samples<-gsub("[.].*", "", names(unlist(xl))) %>% unique()
# this samples need to be removed

if(length(na.samples)>0){
  spec.tss<-spec.tss[, -1*which(names(spec.tss) %in% na.samples )]
} else {
  cat("no NAs found")
}
  
# are there empty rows
spec.tss[rowSums(spec.tss) == 0,] %>% rownames()


```
There are 2036 Species and 4839 samples


# Load and format the metadata

```{r, message =F, warning =F}
meta<-read.delim(paste0(PATHanalNov2023,"PrepData/IBDmeta.cur.filt.mc.matched.txt"), header = T, sep = "\t") # load a single time point metadata
dim(meta)
sapply(meta, class)

# check how many UIBD and remove 
meta %>%
  group_by(IBD2)%>%
  count()
meta<-meta[which(meta$IBD2!="IBDU"),]

meta$IBD<-factor(meta$IBD, levels = c("healthy", "IBD"))
meta$IBD2<-factor(meta$IBD2, levels = c("healthy", "UC", "CD"))

# filter the species according to samples in the meta
spec.tss<-spec.tss[, which(names(spec.tss) %in% meta$SmplID)]
dim(spec.tss)

spec.tss[rowSums(spec.tss) == 0,] %>% rownames() # now there are species with 0. Remove them

spec.tss<-spec.tss[rowSums(spec.tss) != 0,] 
dim(spec.tss)

spec.tss[,colSums(spec.tss) == 0] %>% colnames()
spec.tss<-spec.tss[,colSums(spec.tss) != 0] # now there are samples with 0. Remove them
dim(spec.tss)

# I need to remove the same from the metadata
meta<-meta[meta$SmplID %in% names(spec.tss),]
meta<-meta[order(meta$SmplID),]
spec.tss<-spec.tss[, order(names(spec.tss))]
all(names(spec.tss) == meta$SmplID)
dim(spec.tss)
dim(meta)

```

## import human reads

```{r}
humanReads<-read.delim2(paste0(PATHanalNov2023,"humanReads/IBD2022.4.map"), comment.char = "#", header=FALSE)%>%
  select(1,36)
colnames(humanReads)<-c("SmplID","HumanRds")
head(humanReads)

meta<-inner_join(meta,humanReads,by="SmplID")
meta$HumanRds<-as.numeric(meta$HumanRds)
```

## Check if all samples hava the right metadata

Check if the cohort is specified for all
```{r, message =F, warning =F}
meta$Cohort_precise %>% unique()
meta$Cohort %>% unique()

```

All have both cohort and cohort precise

## Check all samples have IBD assigned
```{r, message =F, warning =F}
meta$IBD %>% unique()
meta$IBD2 %>% unique()

```


## Check all cohort have case_controls
```{r, message =F, warning =F}
meta %>%
  dplyr::group_by(Cohort_precise) %>%
  summarise(unique(IBD)) %>%
  print(n = 40)

```

There are two cohorts with no controls:
Fang_UCSD_IBD_Biobank,USA



# Select first 4 samples for each indiviual
```{r}
# add number of visits
meta<-meta[order(meta$days_after_first_visit),]
meta<- meta %>%
  group_by(IndividualID) %>%
  mutate(visit = row_number())

# select only visits from 1:4

meta4<-subset(meta, visit %in% 1:4) # change here for more or less reps
#meta4<-meta 

# select only a single patient ID and calcualte beta
df.beta.pair<-data.frame(
  IndividualID = vector(length = length(unique(meta4$IndividualID))),
  average = vector(length = length(unique(meta4$IndividualID))),
  maxdist = vector(length = length(unique(meta4$IndividualID))),
  stdev = vector(length = length(unique(meta4$IndividualID))),
  cohort = vector(length = length(unique(meta4$IndividualID))),
  IBD = vector(length = length(unique(meta4$IndividualID))),
  IBD2 = vector(length = length(unique(meta4$IndividualID))),
  deltaHumanRds = vector(length = length(unique(meta4$IndividualID)))
)

for(i in 1:length(unique(meta4$IndividualID))) {

  meta.tmp<-subset(meta4, IndividualID == unique(meta4$IndividualID)[i])
  if(nrow(meta.tmp) == 4){  # change here for more or less reps, needs to match above
  spec.tss.tmp<-spec.tss[, which(names(spec.tss) %in% meta.tmp$SmplID)]
  spec.tss.tmp<-spec.tss.tmp[rowSums(spec.tss.tmp) != 0,] # to remove empty rows otherwise vegan complains
  spec.tss.tmp<-t(spec.tss.tmp) # transpose here
  dist.v<-as.vector(vegan::vegdist(spec.tss.tmp, method = "bray"))

  delta_hr<-(max(meta.tmp$HumanRds)-min(meta.tmp$HumanRds))

  df.beta.pair$IndividualID[i]<-meta.tmp$IndividualID[1]
  df.beta.pair$average[i]<-mean(dist.v)
  df.beta.pair$maxdist[i]<-max(dist.v)
  df.beta.pair$stdev[i]<-sd(dist.v)
  df.beta.pair$cohort[i]<-meta.tmp$Cohort_precise[1]
  df.beta.pair$IBD[i]<-meta.tmp$IBD[1] %>% as.character()
  df.beta.pair$IBD2[i]<-meta.tmp$IBD2[1] %>% as.character()
  df.beta.pair$deltaHumanRds[i]<-delta_hr

  }
}

# remove the skipped samples
df.beta.pair<-subset(df.beta.pair, cohort != FALSE)
df.beta.pair$IBD2<-factor(df.beta.pair$IBD2, levels = c("healthy", "UC", "CD"))
df.beta.pair<-df.beta.pair%>%
  mutate(average_log2=log2(average))%>%
  mutate(deltaHumanRds_log2=log2(deltaHumanRds))%>%
  mutate(maxdist_log2=log2(maxdist))

# plot this
ggplot(df.beta.pair, aes(IBD2, average, fill = IBD2)) +
  geom_boxplot() +
  geom_jitter()+
  scale_fill_brewer(type = "div", palette = "Set2") +
  theme_bw() +
  labs(title = "Average distances between 4 consecutive samples across conditions",
       subtitle = "Bray-Crutis distances were used")

```


```{r}
cor_iiv_hr<-cor.test(df.beta.pair$average, df.beta.pair$deltaHumanRds, use = "complete.obs", method="spearman")
cor_iiv_hr_CD<-cor.test((df.beta.pair%>%filter(IBD2=="CD"))$average, (df.beta.pair%>%filter(IBD2=="CD"))$deltaHumanRds, use = "complete.obs", method="spearman")
cor_iiv_hr_UC<-cor.test((df.beta.pair%>%filter(IBD2=="UC"))$average, (df.beta.pair%>%filter(IBD2=="UC"))$deltaHumanRds, use = "complete.obs", method="spearman")
cor_iiv_hr_CTRL<-cor.test((df.beta.pair%>%filter(IBD2=="healthy"))$average, (df.beta.pair%>%filter(IBD2=="healthy"))$deltaHumanRds, use = "complete.obs", method="spearman")


cor.test((df.beta.pair%>%filter(cohort=="LIBD" & IBD2=="CD"))$average, (df.beta.pair%>%filter(cohort=="LIBD" & IBD2=="CD"))$deltaHumanRds, use = "complete.obs", method="spearman")
cor.test((df.beta.pair%>%filter(cohort=="LIBD"))$average, (df.beta.pair%>%filter(cohort=="LIBD"))$deltaHumanRds, use = "complete.obs", method="spearman")

# fit log2 transformed
dist_hr_log2_lm<-lmerTest::lmer(maxdist_log2 ~ deltaHumanRds_log2*IBD2 + (1|cohort), data=df.beta.pair)
summary(dist_hr_log2_lm)
anova(dist_hr_log2_lm)
lmerTest::ranova(dist_hr_log2_lm) # cohort effect
qqnorm(resid(dist_hr_log2_lm))

emmeans::emm(dist_hr_log2_lm,pairwise~deltaHumanRds_log2)

saveRDS(df.beta.pair,paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/beta.pair.tempInd.rds"))
```

# Testing correlation between BC distances and human read differences
# Select 4 or more samples for each indiviual
```{r fig.height=20, fig.width=10}

# There are some Individual IDs that overlap between HMP and HMP2
# In total 88 overlaps and 46 unique in HMP and 18 unique in HMP2
# To proceed with per-individual calculations that we will account for cohort later we treat the overlapping individuals as separate individuals
HMP_Inds<-(meta%>%filter(Cohort_precise=="HMP"))$IndividualID%>%unique()
HMP2_Inds<-(meta%>%filter(Cohort_precise=="HMP2"))$IndividualID%>%unique()
HMP_uniqueInds<-setdiff(HMP_Inds,HMP2_Inds)
HMP2_uniqueInds<-setdiff(HMP2_Inds,HMP_Inds)
HMP_HMP2_sharedInds<-HMP_Inds[HMP_Inds%in%HMP2_Inds]
length(HMP_uniqueInds)
length(HMP2_uniqueInds)
length(HMP_HMP2_sharedInds)
length(HMP_uniqueInds)+length(HMP2_uniqueInds)+length(HMP_HMP2_sharedInds)
length((meta%>%filter(Cohort_precise%in%c("HMP","HMP2")))$IndividualID%>%unique())

# ammend the cohort to overlapping Individuals IDs and save to new column IndividualID2
meta_splitHMP<-meta%>%
  mutate(IndividualID2=ifelse(IndividualID%in%HMP_HMP2_sharedInds,paste0(IndividualID,"_",Cohort_precise),IndividualID))

perInd_dist_vs_deltaHR<-lapply(unique(meta_splitHMP$IndividualID2), function(x) {
  meta.tmp<-meta_splitHMP%>%filter(IndividualID2 == x)

    if(nrow(meta.tmp) >= 4){  # change here for more or less reps, needs to match above
    spec.tss.tmp<-spec.tss[, which(names(spec.tss) %in% meta.tmp$SmplID)]
    spec.tss.tmp<-spec.tss.tmp[rowSums(spec.tss.tmp) != 0,] # to remove empty rows otherwise vegan complains
    spec.tss.tmp<-t(spec.tss.tmp) ## EDIT needed to transpose here
    dist_ab<-vegan::vegdist(spec.tss.tmp, method = "bray")
    
    meta.tmp2<-meta.tmp%>%
      column_to_rownames("SmplID")%>%
      select(c("HumanRds"))
  
    mat_dist_ab<-as.matrix(dist_ab)
    mat_hr<-mat_dist_ab
    mat_hr[!is.na(mat_hr)]<-NA
    all(rownames(mat_hr)==colnames(mat_hr))
    
    for (n in (1:ncol(mat_hr))){
      for (i in (1:nrow(mat_hr))){
        a=rownames(mat_hr)[i]
        b=rownames(mat_hr)[n]
        #print(paste(a,b))
        deltaHR<-(meta.tmp2[a,]-meta.tmp2[b,])
        mat_hr[a,b]<-abs(deltaHR)
      }
    }
    
    dist_hr<-as.dist(mat_hr)
    all(labels(dist_hr)==labels(dist_ab))
    
    
    df.tmp<-data.frame(
      delta_human_reads=as.vector(dist_hr),
      microbiome_bc_distance=as.vector(dist_ab),
      IndividualID2=paste(x)
    )
    
    return(df.tmp)
    }
  
})
names(perInd_dist_vs_deltaHR)<-unique(meta_splitHMP$IndividualID2)


perInd_dist_vs_deltaHR.df<-do.call(rbind, perInd_dist_vs_deltaHR)
perInd_dist_vs_deltaHR_meta.df<-perInd_dist_vs_deltaHR.df%>%
  left_join(meta_splitHMP%>%
              ungroup(IndividualID) %>%
              select(c(IndividualID2,IBD,IBD2,Cohort_precise))%>%
              distinct(),
            by="IndividualID2")%>%
  rename(cohort=Cohort_precise)%>%
  mutate(delta_human_reads_log2=log2(delta_human_reads))%>% 
  mutate(microbiome_bc_distance_log2=log2(microbiome_bc_distance)) 

# number of Individuals
perInd_dist_vs_deltaHR_meta.df$IndividualID2%>%unique()%>%length()

```


```{r}
# fitting linear models
## set variables to factors
perInd_dist_vs_deltaHR_meta.df$cohort<-as.factor(perInd_dist_vs_deltaHR_meta.df$cohort)
perInd_dist_vs_deltaHR_meta.df$IBD2<-as.factor(perInd_dist_vs_deltaHR_meta.df$IBD2)
## make sure testing values are numeric
class(perInd_dist_vs_deltaHR_meta.df$microbiome_bc_distance_log2)
class(perInd_dist_vs_deltaHR_meta.df$delta_human_reads_log2)

# remove rows with NA
nrow(perInd_dist_vs_deltaHR_meta.df[-which(is.na(perInd_dist_vs_deltaHR_meta.df)),])
nrow(perInd_dist_vs_deltaHR_meta.df)
perInd_dist_vs_deltaHR_meta.df<-perInd_dist_vs_deltaHR_meta.df[-which(is.na(perInd_dist_vs_deltaHR_meta.df)),]
anyNA(perInd_dist_vs_deltaHR_meta.df)

# remove Sample with 0 in both values
perInd_dist_vs_deltaHR_meta.df<-perInd_dist_vs_deltaHR_meta.df%>%
  filter(microbiome_bc_distance>0 & delta_human_reads>0)

# fitting mixed effects model with random effect Individual nested within cohort
# log2 transformed
dist_hr_indNest_log2_lm<-lmerTest::lmer(microbiome_bc_distance_log2 ~ delta_human_reads_log2*IBD2 + (1|cohort/IndividualID2), data=perInd_dist_vs_deltaHR_meta.df)
summary(dist_hr_indNest_log2_lm)
anova(dist_hr_indNest_log2_lm)

lmerTest::ranova(dist_hr_indNest_log2_lm) # cohort effect
qqnorm(resid(dist_hr_indNest_log2_lm))

# fitting mixed effects model with random effect Individual separately per cohort
## log2 transformed per cohort
cohorts<-perInd_dist_vs_deltaHR_meta.df$cohort%>%as.character()%>%unique()

dist_hr_indNest_log2_lm_listCohorts<-lapply(cohorts, function(c){
  if(c=="Lewis"){
    df_cohort<-perInd_dist_vs_deltaHR_meta.df%>%filter(cohort==c)
    dist_hr_indNest_log2_lm<-lmerTest::lmer(microbiome_bc_distance_log2 ~ delta_human_reads_log2 + (1|IndividualID2), data=df_cohort)
  } else {
    df_cohort<-perInd_dist_vs_deltaHR_meta.df%>%filter(cohort==c)
    dist_hr_indNest_log2_lm<-lmerTest::lmer(microbiome_bc_distance_log2 ~ delta_human_reads_log2*IBD2 + (1|IndividualID2), data=df_cohort)
  }
  return(dist_hr_indNest_log2_lm)
})
names(dist_hr_indNest_log2_lm_listCohorts)<-cohorts

summary(dist_hr_indNest_log2_lm_listCohorts$USA)
summary(dist_hr_indNest_log2_lm_listCohorts$Fang_UCSD_IBD_Biobank)
summary(dist_hr_indNest_log2_lm_listCohorts$HMP2)
summary(dist_hr_indNest_log2_lm_listCohorts$HMP)
summary(dist_hr_indNest_log2_lm_listCohorts$Lewis)
summary(dist_hr_indNest_log2_lm_listCohorts$LIBD)

dist_hr_indNest_log2_lm_perCohort<-data.frame(
  USA=c(summary(dist_hr_indNest_log2_lm_listCohorts$USA)$coefficients[2,5],summary(dist_hr_indNest_log2_lm_listCohorts$USA)$coefficients[2,1]),
  IBD_Biobank=c(summary(dist_hr_indNest_log2_lm_listCohorts$Fang_UCSD_IBD_Biobank)$coefficients[2,5],summary(dist_hr_indNest_log2_lm_listCohorts$Fang_UCSD_IBD_Biobank)$coefficients[2,1]),
  HMP2=c(summary(dist_hr_indNest_log2_lm_listCohorts$HMP2)$coefficients[2,5],summary(dist_hr_indNest_log2_lm_listCohorts$HMP2)$coefficients[2,1]),
  HMP=c(summary(dist_hr_indNest_log2_lm_listCohorts$HMP)$coefficients[2,5],summary(dist_hr_indNest_log2_lm_listCohorts$HMP)$coefficients[2,1]),
  Lewis=c(summary(dist_hr_indNest_log2_lm_listCohorts$Lewis)$coefficients[2,5],summary(dist_hr_indNest_log2_lm_listCohorts$Lewis)$coefficients[2,1]),
  LIBD=c(summary(dist_hr_indNest_log2_lm_listCohorts$LIBD)$coefficients[2,5],summary(dist_hr_indNest_log2_lm_listCohorts$LIBD)$coefficients[2,1])
)
rownames(dist_hr_indNest_log2_lm_perCohort)<-c("p","estimate")


perInd_dist_vs_deltaHR_meta.df%>%
  group_by(cohort)%>%
  count(IBD2)

saveRDS(dist_hr_indNest_log2_lm,paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/dist_hr_indNest_log2_lm.rds"))
saveRDS(dist_hr_indNest_log2_lm_perCohort,paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/dist_hr_indNest_log2_lm_perCohort.rds"))

```

```{r fig.width=8, fig.height=3}
# assess per individual correlation between pairwise BC dissimilarity and human read difference

## association coefficient across Individuals and cohorts
acrossIndChrts_dist_vs_deltaHR_meta_slope.df<-perInd_dist_vs_deltaHR_meta.df%>%
    ungroup()%>%
    filter(IBD2 !="CD" | cohort!="HMP")%>%
    group_by(IBD2) %>%
    summarize(slope = lm(microbiome_bc_distance_log2 ~ delta_human_reads_log2)$coeff[2])

## association coefficient across Individuals per cohort and condition
acrossInd_dist_vs_deltaHR_meta_slope.df<-perInd_dist_vs_deltaHR_meta.df%>%
    ungroup()%>%
    group_by(cohort, IBD2) %>%
    summarize(slope = lm(microbiome_bc_distance_log2 ~ delta_human_reads_log2)$coeff[2])

## association coefficient per Individual
perInd_dist_vs_deltaHR_meta_slope.df<-perInd_dist_vs_deltaHR_meta.df%>%
    group_by(IndividualID2) %>%
    mutate(slope = lm(microbiome_bc_distance_log2 ~ delta_human_reads_log2)$coeff[2])%>%
    mutate(slope2=ifelse(slope>0,"pos","neg"))

perInd_dist_vs_deltaHR_meta_slope.df%>%
  group_by(cohort)%>%
  count(slope2)

perInd_dist_vs_deltaHR_meta_slope.df%>%
  group_by(IBD2)%>%
  count(slope2)

slope_summary_Ind<-perInd_dist_vs_deltaHR_meta_slope.df%>%
  group_by(cohort,IBD2)%>%
  count(slope2)%>%
  spread(slope2,n)

## mean slope of association 
slopeDegree_summary_acrossInd<-acrossInd_dist_vs_deltaHR_meta_slope.df%>%
  group_by(cohort, IBD2)%>%
  summarize(mean(slope))

slopeDF<-perInd_dist_vs_deltaHR_meta_slope.df%>%distinct(IndividualID2, .keep_all = TRUE)
ggplot(slopeDF, aes(x=IBD2,y=slope))+
  geom_boxplot(aes(fill=IBD2))

slopeDF$IBD2<-as.factor(slopeDF$IBD2)
slopeDF$cohort<-as.factor(slopeDF$cohort)
coin::wilcox_test(slope ~ IBD2 | cohort, data=slopeDF%>%filter(IBD2!="CD"))
coin::wilcox_test(slope ~ IBD2 | cohort, data=slopeDF%>%filter(IBD2!="UC"))


saveRDS(perInd_dist_vs_deltaHR_meta_slope.df,paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RDS/perInd_dist_vs_deltaHR.rds"))

```

### Compare high HR vs low HR to controls

Use 20th and 80 quantile of human read percent to define high and low HR

```{r fig.height=4, fig.width=14}
# group per cohort since percentages of HR vary between cohorts
meta_hr_percentiles<-meta%>%
  group_by(Cohort_precise)%>%
  summarize(HighHR_perc80=quantile(as.numeric(HumanRds), 0.8, na.rm=TRUE),LowHR_perc20=quantile(as.numeric(HumanRds), 0.2, na.rm=TRUE))

# join classification of samples into low and high HR groups with metadata
meta_hr_num<-meta%>%
  left_join(meta_hr_percentiles,by="Cohort_precise")%>%
  mutate(HumanRds_group=ifelse(HumanRds>HighHR_perc80,"HighHR",ifelse(HumanRds<LowHR_perc20,"LowHR",NA)))%>% 
  filter(!is.na(HumanRds_group))%>%
  distinct(IndividualID, .keep_all = TRUE)


# group per cohort and conditionto obtain disease subtype-specific classification of high and low HR groups
meta_hr_cond_percentiles<-meta%>%
  group_by(Cohort_precise, IBD2)%>%
  summarize(HighHR_perc80=quantile(as.numeric(HumanRds), 0.8, na.rm=TRUE),LowHR_perc20=quantile(as.numeric(HumanRds), 0.2, na.rm=TRUE))

# join classification of samples into low and high HR groups with metadata
meta_hr_cond_num<-meta%>%
  left_join(meta_hr_cond_percentiles,by=c("Cohort_precise","IBD2"))%>%
  mutate(HumanRds_group=ifelse(IBD2=="healthy","control",ifelse(HumanRds>HighHR_perc80,"HighHR",ifelse(HumanRds<LowHR_perc20,"LowHR",NA))))%>% 
  filter(!is.na(HumanRds_group))%>%
  distinct(IndividualID, .keep_all = TRUE)

#check number of Individuals with samples classified as high or low HR for both
meta_hr_num$IndividualID%>%length()
meta_hr_num$IndividualID%>%unique()%>%length()
meta_hr_cond_num$IndividualID%>%length()
meta_hr_cond_num$IndividualID%>%unique()%>%length()
```

Use classification based on per-cohort classification:

```{r}

## subset species abundances in samples to samples only that are classified as high or low HR

meta_hr_num%>%
  group_by(Cohort_precise)%>%
  count(HumanRds_group)

spec.tss.HL_hr<-spec.tss[, which(names(spec.tss) %in% meta_hr_num$SmplID)]
spec.tss.HL_hr<-spec.tss.HL_hr[rowSums(spec.tss.HL_hr) != 0,] # to remove empty rows otherwise vegan complains

all(colnames(spec.tss.HL_hr)==meta_hr_num$SmplID)
spec.tss.HL_hr <- spec.tss.HL_hr[,meta_hr_num$SmplID]
all(colnames(spec.tss.HL_hr)==meta_hr_num$SmplID)

spec.tss.HL_hr<-t(spec.tss.HL_hr) # transpose
dist.HL_hr<-vegan::vegdist(spec.tss.HL_hr, method = "bray") # bray-curtis dissimilarities between samples

all(labels(dist.HL_hr)==meta_hr_num$SmplID)

## Permanova to assess whether significant differences in bray-curtis dissimilarity between high and low HR groups - restrict permutations within cohort
set.seed(222)
perm<-how(blocks = meta_hr_num$Cohort_precise, nperm = 999)
pmva.HL_hr<-adonis2(formula = dist.HL_hr ~ HumanRds_group, data = meta_hr_num, parallel = T, permutations = perm)
pmva.HL_hr

# Get pairwise dissimilarities into dataframe
dist.HL_hr_gat<-dist.HL_hr%>%
  as.matrix()%>%
  as.data.frame()%>%
  rownames_to_column("SmplID1")%>%
  gather(SmplID2,dist,-SmplID1)%>%
  left_join(meta_hr_num,by=c("SmplID1"="SmplID"))%>%
  select(SmplID1,SmplID2,dist,HumanRds_group,Cohort_precise,IBD2)%>%
  rename("HumanRds_group1"=HumanRds_group)%>%
  rename("Cohort_precise1"=Cohort_precise)%>%
  rename("IBD21"=IBD2)%>%
  left_join(meta_hr_num,by=c("SmplID2"="SmplID"))%>%
  select(SmplID1,SmplID2,dist,HumanRds_group1,Cohort_precise1,IBD21,HumanRds_group,Cohort_precise,IBD2)%>%
  rename("HumanRds_group2"=HumanRds_group)%>%
  rename("Cohort_precise2"=Cohort_precise)%>%
  rename("IBD22"=IBD2)%>%
  mutate(comparison=ifelse(HumanRds_group1==HumanRds_group2,HumanRds_group1,"LowHR-HighHR"))

# compare pairwise BC dissimilarities among lowHR samples with BC dissimilarities among HighHR samples
dist.HL_hr_gat.sub1<-dist.HL_hr_gat%>%
  filter(comparison!="LowHR-HighHR")%>%
  filter(Cohort_precise1==Cohort_precise2)%>%
  mutate(comparison=as.factor(comparison))%>%
  mutate(cohort=as.factor(Cohort_precise1))

# wilcoxon test blocked by cohort
wt_HL_hr<-coin::wilcox_test(dist ~comparison | cohort, data=dist.HL_hr_gat.sub1)
wt_HL_hr
# calculate effect size r
n<-nrow(dist.HL_hr_gat.sub1)
coin::statistic(wt_HL_hr)/sqrt(n)

# create plot
col_compHL<-c("darkorange1","khaki2")
names(col_compHL)<-c("HighHR","LowHR")

p_HL_IBD<-ggplot(dist.HL_hr_gat%>%filter(comparison!="LowHR-HighHR"), aes(x=comparison,y=dist))+
  #geom_violin()+
  geom_boxplot(aes(fill=comparison))+
  scale_fill_manual(values=col_compHL, breaks=names(col_compHL), name="inflammation index", labels=c("flare-like","remission-like"))+
  theme_bw()+
  ylab("BC dissimilarity")+
  xlab("")
```

Use classification based on per-cohort and per-condotion classification:

```{r}
# use classification based on per-cohort and per-condotion classification

## subset species abundances in samples to samples only that are classified as high or low HR

spec.tss.HL_hr_cond<-spec.tss[, which(names(spec.tss) %in% meta_hr_cond_num$SmplID)]
spec.tss.HL_hr_cond<-spec.tss.HL_hr_cond[rowSums(spec.tss.HL_hr_cond) != 0,] # to remove empty rows otherwise vegan complains

all(colnames(spec.tss.HL_hr_cond)==meta_hr_cond_num$SmplID)
spec.tss.HL_hr_cond <- spec.tss.HL_hr_cond[,meta_hr_cond_num$SmplID]
all(colnames(spec.tss.HL_hr_cond)==meta_hr_cond_num$SmplID)

spec.tss.HL_hr_cond<-t(spec.tss.HL_hr_cond)
dist.HL_hr_cond<-vegan::vegdist(spec.tss.HL_hr_cond, method = "bray")

all(labels(dist.HL_hr_cond)==meta_hr_cond_num$SmplID)

col_comp<-c("darkorange1","khaki2","#1F78B4")
names(col_comp)<-c("control-HighHR","control-LowHR","control")

# Get pairwise dissimilarities into dataframe
dist.HL_hr_cond_gat<-dist.HL_hr_cond%>%
  as.matrix()%>%
  as.data.frame()%>%
  rownames_to_column("SmplID1")%>%
  gather(SmplID2,dist,-SmplID1)%>%
  left_join(meta_hr_cond_num,by=c("SmplID1"="SmplID"))%>%
  select(SmplID1,SmplID2,dist,HumanRds_group,Cohort_precise,IBD2)%>%
  rename("HumanRds_group1"=HumanRds_group)%>%
  rename("Cohort_precise1"=Cohort_precise)%>%
  rename("IBD21"=IBD2)%>%
  left_join(meta_hr_cond_num,by=c("SmplID2"="SmplID"))%>%
  select(SmplID1,SmplID2,dist,HumanRds_group1,Cohort_precise1,IBD21,HumanRds_group,Cohort_precise,IBD2)%>%
  rename("HumanRds_group2"=HumanRds_group)%>%
  rename("Cohort_precise2"=Cohort_precise)%>%
  rename("IBD22"=IBD2)%>%
  mutate(comparison=ifelse(HumanRds_group1==HumanRds_group2,HumanRds_group1,paste0(HumanRds_group1,"_",HumanRds_group2)))

ctrl_low<-c("control_LowHR","LowHR_control")
ctrl_high<-c("control_HighHR","HighHR_control")

dist.HL_hr_cond_gat<-dist.HL_hr_cond_gat%>%
  mutate(comparison2=ifelse(comparison%in%ctrl_low,"control-LowHR",ifelse(comparison%in%ctrl_high,"control-HighHR",comparison)))

# wilcoxon test to test differences in pairwise BC dissimilarity between groups control-HoghHR and control-lowHR and highHR-lowHR
wilcox.test((dist.HL_hr_cond_gat%>%filter(comparison=="control"))$dist,(dist.HL_hr_cond_gat%>%filter(comparison=="HighHR"))$dist)
wilcox.test((dist.HL_hr_cond_gat%>%filter(comparison=="control"))$dist,(dist.HL_hr_cond_gat%>%filter(comparison=="LowHR"))$dist)
wilcox.test((dist.HL_hr_cond_gat%>%filter(comparison%in%ctrl_high))$dist,(dist.HL_hr_cond_gat%>%filter(comparison%in%ctrl_low))$dist)

# only within cohort
dist.HL_hr_cond_gat_intrChrt<-dist.HL_hr_cond_gat%>%
  filter(Cohort_precise1==Cohort_precise2)%>%
  filter(comparison%in%c(ctrl_low,ctrl_high,"control"))%>%
  mutate(cohort=gsub("Fang_UCSD_","",Cohort_precise1))%>%
  mutate(cohort=gsub("cohort1000","1000",Cohort_precise1))

# wilcoxon tests
dist.HL_hr_cond_gat_intrChrt$comparison<-as.factor(dist.HL_hr_cond_gat_intrChrt$comparison)
dist.HL_hr_cond_gat_intrChrt$comparison2<-factor(dist.HL_hr_cond_gat_intrChrt$comparison2, levels=names(col_comp))
dist.HL_hr_cond_gat_intrChrt$cohort<-as.factor(dist.HL_hr_cond_gat_intrChrt$cohort)
wt.ctrlLow_vs_ctrlHigh<-coin::wilcox_test(dist ~ comparison2 | cohort, data=dist.HL_hr_cond_gat_intrChrt%>%filter(comparison%in%c(ctrl_low,ctrl_high)))
n.ctrlLow_vs_ctrlHigh<-dist.HL_hr_cond_gat_intrChrt%>%filter(comparison%in%c(ctrl_low,ctrl_high))%>%nrow()

wt.ctrl_vs_ctrlLow<-coin::wilcox_test(dist ~ comparison2 | cohort, data=dist.HL_hr_cond_gat_intrChrt%>%filter(comparison2%in%c("control","control-LowHR")))
n.ctrl_vs_ctrlLow<-dist.HL_hr_cond_gat_intrChrt%>%filter(comparison2%in%c("control","control-LowHR"))%>%nrow()

wt.ctrl_vs_ctrlHigh<-coin::wilcox_test(dist ~ comparison2 | cohort, data=dist.HL_hr_cond_gat_intrChrt%>%filter(comparison2%in%c("control","control-HighHR")))
n.ctrl_vs_ctrlHigh<-dist.HL_hr_cond_gat_intrChrt%>%filter(comparison2%in%c("control","control-HighHR"))%>%nrow()

# merge wilcoxon test results into df
lowHR_vs_highHR.df<-data.frame(
  comparison=c("ctrlLow_vs_ctrlHigh","ctrl_vs_ctrlLow","ctrl_vs_ctrlHigh"),
  p=c(coin::pvalue(wt.ctrlLow_vs_ctrlHigh),coin::pvalue(wt.ctrl_vs_ctrlLow),coin::pvalue(wt.ctrl_vs_ctrlHigh)),
  effectSize=c(coin::statistic(wt.ctrlLow_vs_ctrlHigh)/sqrt(n.ctrlLow_vs_ctrlHigh),coin::statistic(wt.ctrl_vs_ctrlLow)/sqrt(n.ctrl_vs_ctrlLow),coin::statistic(wt.ctrl_vs_ctrlHigh)/sqrt(n.ctrl_vs_ctrlHigh))
)
# adjust p-value
lowHR_vs_highHR.df$padjust<-p.adjust(lowHR_vs_highHR.df$p, method = "BH" )
```

```{r}
# create plots

# per cohort
dist.HL_hr_cond_gat_intrChrt<-dist.HL_hr_cond_gat_intrChrt%>%
  mutate(comparison3=gsub("HR","Iidx",comparison2))
col_comp2<-col_comp
names(col_comp2)<-gsub("HR","Iidx",names(col_comp))
p_dist_HR_chrts<-ggplot(dist.HL_hr_cond_gat_intrChrt, aes(x=comparison3,y=dist))+
  geom_boxplot(aes(fill=comparison3),outlier.size = 0.1)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45,hjust=1))+
  scale_fill_manual(values=col_comp2, breaks=names(col_comp2), name="comparison", labels=c("control-flare","control-remission","control"))+
  facet_grid(.~cohort)+
  ylab("Pairwise BC dissimilarity")
p_dist_HR_chrts

# across cohorts
p_dist_HR_all<-
  ggplot(dist.HL_hr_cond_gat_intrChrt, aes(x=comparison2,y=dist))+
  geom_boxplot(aes(fill=comparison2),outlier.size = 0.1)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45,hjust=1))+
  scale_fill_manual(values=col_comp, breaks=names(col_comp))+
  ylab("Pairwise BC dissimilarity")
p_dist_HR_all

# combine across and per-cohort
p_dist_HR_combined<-plot_grid(p_dist_HR_all+theme(legend.position="none", axis.text.x=element_blank(), axis.title.x=element_blank(), axis.ticks.x=element_blank()),
          p_dist_HR_chrts+theme(legend.position="top",axis.text.x=element_blank(), axis.text.y=element_blank(),axis.title=element_blank(), axis.ticks=element_blank()),
          nrow=1, align="h", axis="tb", rel_widths = c(1,8.5),labels=c("A","B"))
p_dist_HR_combined
```

```{r}
# export figs
pdf(paste0(PATHanalNov2023,"Downstream/Results/Figures/dist_ctrl_hlHR.pdf"), width=14,height=4)
p_dist_HR_combined
dev.off()


pdf(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_dist_ctrl_hlHR.pdf"), width=13,height=3)
p_dist_HR_chrts+theme(legend.position="top",axis.text.x=element_blank(), axis.title.x=element_blank(), axis.ticks=element_blank())
dev.off()

ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_dist_ctrl_hlHR.png"),p_dist_HR_chrts+theme(legend.position="top",axis.text.x=element_blank(), axis.title.x=element_blank(), axis.ticks=element_blank()), width=14,height=4)


saveRDS(dist.HL_hr_cond_gat_intrChrt,paste0(PATHanalNov2023,"humanReads/dist.HL_hr_cond_gat_intrChrt.rds"))

```

Assess difference in high HR and low HR groups in controls only

```{r}
## controls only high infl vs low Infl
meta_hr_cond_num_CTRL<-meta%>%
  left_join(meta_hr_cond_percentiles,by=c("Cohort_precise","IBD2"))%>%
  filter(IBD2=="healthy")%>%
  mutate(HumanRds_group=ifelse(HumanRds>HighHR_perc80,"HighHR",ifelse(HumanRds<LowHR_perc20,"LowHR",NA)))%>% 
  filter(!is.na(HumanRds_group))%>%
  distinct(IndividualID, .keep_all = TRUE)

meta_hr_cond_num_CTRL%>%
  group_by(Cohort_precise)%>%
  count(HumanRds_group)

spec.tss.HL_hr_cond_CTRL<-spec.tss[, which(names(spec.tss) %in% meta_hr_cond_num_CTRL$SmplID)]
spec.tss.HL_hr_cond_CTRL<-spec.tss.HL_hr_cond_CTRL[rowSums(spec.tss.HL_hr_cond_CTRL) != 0,] # to remove empty rows otherwise vegan complains

all(colnames(spec.tss.HL_hr_cond_CTRL)==meta_hr_cond_num_CTRL$SmplID)
spec.tss.HL_hr_cond_CTRL <- spec.tss.HL_hr_cond_CTRL[,meta_hr_cond_num_CTRL$SmplID]
all(colnames(spec.tss.HL_hr_cond_CTRL)==meta_hr_cond_num_CTRL$SmplID)

spec.tss.HL_hr_cond_CTRL<-t(spec.tss.HL_hr_cond_CTRL)
dist.HL_hr_cond_CTRL<-vegan::vegdist(spec.tss.HL_hr_cond_CTRL, method = "bray")

all(labels(dist.HL_hr_cond_CTRL)==meta_hr_cond_num_CTRL$SmplID)

dist.HL_hr_cond_gat_CTRL<-dist.HL_hr_cond_CTRL%>%
  as.matrix()%>%
  as.data.frame()%>%
  rownames_to_column("SmplID1")%>%
  gather(SmplID2,dist,-SmplID1)%>%
  left_join(meta_hr_cond_num_CTRL,by=c("SmplID1"="SmplID"))%>%
  select(SmplID1,SmplID2,dist,HumanRds_group,Cohort_precise,IBD2)%>%
  rename("HumanRds_group1"=HumanRds_group)%>%
  rename("Cohort_precise1"=Cohort_precise)%>%
  rename("IBD21"=IBD2)%>%
  left_join(meta_hr_cond_num_CTRL,by=c("SmplID2"="SmplID"))%>%
  select(SmplID1,SmplID2,dist,HumanRds_group1,Cohort_precise1,IBD21,HumanRds_group,Cohort_precise,IBD2)%>%
  rename("HumanRds_group2"=HumanRds_group)%>%
  rename("Cohort_precise2"=Cohort_precise)%>%
  rename("IBD22"=IBD2)%>%
  mutate(comparison=ifelse(HumanRds_group1==HumanRds_group2,HumanRds_group1,paste0(HumanRds_group1,"_",HumanRds_group2)))

# wicoxon test to test difference in pairwise BC dissimilarities between high HR and low HR groups in controls; blocked by cohort

dist.HL_hr_cond_gat_CTRL_filt<-dist.HL_hr_cond_gat_CTRL%>%
  filter(Cohort_precise1==Cohort_precise2)%>%
  filter(comparison%in%c("LowHR","HighHR"))

dist.HL_hr_cond_gat_CTRL_filt$comparison<-as.factor(dist.HL_hr_cond_gat_CTRL_filt$comparison)
dist.HL_hr_cond_gat_CTRL_filt$cohort<-as.factor(dist.HL_hr_cond_gat_CTRL_filt$Cohort_precise2)
wt.low_vs_high_CTRL<-coin::wilcox_test(dist ~ comparison | cohort, data=dist.HL_hr_cond_gat_CTRL_filt)
wt.low_vs_high_CTRL

# calculate effect size r
n<-nrow(dist.HL_hr_cond_gat_CTRL_filt)
coin::statistic(wt.low_vs_high_CTRL)/sqrt(n)

# plot
p_HL_CTRL<-ggplot(dist.HL_hr_cond_gat_CTRL_filt, aes(x=comparison,y=dist))+
  geom_boxplot(aes(fill=comparison))+
  scale_fill_manual(values=col_compHL, breaks=names(col_compHL), name="inflammation\nindex", labels=c("flare-like","remission-like"))+
  theme_bw()+
  ylab("BC dissimilarity")+
  xlab("")
p_HL_CTRL

```

```{r}
# combine HR-LR figure supplement
p_suppl_HL<-plot_grid(p_HL_IBD+theme(legend.position = "none",axis.text.x = element_blank(),axis.ticks.x = element_blank())+ggtitle("IBD individuals"), p_HL_CTRL+theme(axis.title.y = element_blank(),axis.text.x = element_blank(),axis.ticks.x = element_blank())+ggtitle("Control individuals"), labels = c("A","B"), rel_widths = c(1,1.45))

ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/supplFig.HL_ibd_ctrls.pdf"),p_suppl_HL,height=4,width=7)
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/supplFig.HL_ibd_ctrls.png"),p_suppl_HL,height=4,width=7)
```

Matching of HR percentages between control and IBD 

```{r}
## matched high infl vs low Infl control - IBD
filtered_HRmaxmin<-meta%>%
  filter(!is.na(HumanRds))%>%
  ungroup()%>%
  group_by(IBD2,Cohort_precise)%>%
  summarise(maxHR=max(HumanRds), minHR=min(HumanRds))%>%
  ungroup()%>%
  group_by(Cohort_precise)%>%
  summarize(maxHRl=min(maxHR),minHRh=max(minHR))%>%
  left_join(meta, by="Cohort_precise")%>%
  filter(HumanRds<maxHRl & HumanRds>minHRh)

# Convert to binary treatment variable
cohortlist<-unique(filtered_HRmaxmin$Cohort_precise)

# do matching per cohort
matched_data.list<-lapply(cohortlist, function(c){
  df <- filtered_HRmaxmin %>%
    distinct(IndividualID, .keep_all = TRUE)%>% # single sample for timeseries data only
    filter(IBD%in%c("healthy","IBD"))%>%
    filter(Cohort_precise==c)%>%
    mutate(treatment = ifelse(IBD == "healthy", 1, 0))
    

  if("healthy" %in% unique(df$IBD)){
    # Perform nearest neighbor matching
    m.out <- matchit(treatment ~ HumanRds, data = df, method = "nearest", distance = "mahalanobis")
    # Get matched dataset
    matched_data <- match.data(m.out)
    } else {
    matched_data <- NULL
    }
  return(matched_data)
})
names(matched_data.list)<-cohortlist

p1_md_hr.list<-lapply(cohortlist, function(c){
  md<-matched_data.list[[c]]
    if(!is.null(md)){
    plot<-ggplot(md, aes(x=IBD,y=HumanRds))+
    geom_boxplot(aes(fill=IBD))
  } else {
    plot<-NULL
  }
  return(plot)
})
names(p1_md_hr.list)<-cohortlist
```

```{r}
# create list with BC dissimilarity per cohort
dist_md_hr.list<-lapply(cohortlist, function(c){
  md<-matched_data.list[[c]]
  if(!is.null(md)){
    md.spec<-spec.tss[, which(names(spec.tss) %in% md$SmplID)]
    md.spec<-md.spec[rowSums(md.spec) != 0,] # to remove empty rows otherwise vegan complains
    
    all(colnames(md.spec)==md$SmplID)
    md.spec <- md.spec[,md$SmplID]
    all(colnames(md.spec)==md$SmplID)
    
    md.spec<-t(md.spec)
    md.dist<-vegan::vegdist(md.spec, method = "bray")
  } else {
    md.dist<- NULL
  }
  return(md.dist)
})
names(dist_md_hr.list)<-cohortlist
```

```{r}
# compute permanova to test wether BC dissimilarites differ between high HR and low HR groups (matched)
pmva_md_hr.list<-lapply(cohortlist, function(c){
  md<-matched_data.list[[c]]
  md.dist<-dist_md_hr.list[[c]]
  if(!is.null(md)){
    print(all(labels(md.dist)==md$SmplID))
    set.seed(222)
    pmva<-adonis2(md.dist ~ IBD2, data = md)
  } else {
    pmva<-NULL
  }
  return(pmva)
})
names(pmva_md_hr.list)<-cohortlist

pmva_md_hr.listDF<-lapply(cohortlist, function(c){
  pmva<-pmva_md_hr.list[[c]]
  if(!is.null(pmva)){
    pmva.df<-data.frame(
      cohort=c,
      p=pmva$`Pr(>F)`[1],
      R2=pmva$R2[1],
      n=nrow(matched_data.list[[c]])
    )

  } else {
    pmva.df<-NULL
  }
  return(pmva.df)
})
names(pmva_md_hr.listDF)<-cohortlist
pmva_md_hr.df<-do.call(rbind, pmva_md_hr.listDF)
pmva_md_hr.df$q<-p.adjust(pmva_md_hr.df$p, method = "BH")
pmva_md_hr.df

# export permanova results to csv
write.csv(pmva_md_hr.df, paste0(PATHanalNov2023,"Downstream/Results/pmva.matchedInfl.csv"), row.names = FALSE)
```

```{r}
# create boxplots per cohort
df.p_box_md_hr.list<-lapply(cohortlist, function(c){
  md<-matched_data.list[[c]]
  md.dist<-dist_md_hr.list[[c]]
  if(!is.null(md)){
    md.dist_gat<-md.dist%>%
      as.matrix()%>%
      as.data.frame()%>%
      rownames_to_column("SmplID1")%>%
      gather(SmplID2,dist,-SmplID1)%>%
      left_join(md,by=c("SmplID1"="SmplID"))%>%
      select(SmplID1,SmplID2,dist,subclass,Cohort_precise,IBD)%>%
      rename("subclass1"=subclass)%>%
      rename("Cohort_precise1"=Cohort_precise)%>%
      rename("IBD1"=IBD)%>%
      left_join(md,by=c("SmplID2"="SmplID"))%>%
      select(SmplID1,SmplID2,dist,subclass1,Cohort_precise1,IBD1,subclass,Cohort_precise,IBD)%>%
      rename("subclass2"=subclass)%>%
      rename("Cohort_precise2"=Cohort_precise)%>%
      rename("IBD2"=IBD)%>%
        filter(SmplID1!=SmplID2)%>%
      mutate(IBD=ifelse(IBD1==IBD2,paste(IBD1),"control-IBD"))%>%
      mutate(IBD=ifelse(IBD=="healthy","control",paste(IBD)))%>%
      filter(IBD!="control-IBD")%>%
      mutate(cohort=Cohort_precise1)
  
  } else {
    md.dist_gat<-NULL
    #plot<-NULL
  }
  return(md.dist_gat)
})
names(df.p_box_md_hr.list)<-cohortlist
df.p_box_md_hr.gat<-do.call(rbind, df.p_box_md_hr.list)
p_box_md_hr<-ggplot(df.p_box_md_hr.gat, aes(x=cohort,y=dist))+
  geom_boxplot(aes(fill=IBD))+
  scale_fill_manual(values=col_IBD2, name="")+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ylab("BC dissimilarity")
p_box_md_hr

# export plots
pdf(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_p_box_matched_hr.pdf"), width=7,height=3)
p_box_md_hr
dev.off()
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_p_box_matched_hr.png"), p_box_md_hr,width=8,height=4)
```

```{r}
# wilcoxon to test difference in the distribution of BC dissimilarities between high HR and low HR groups of matched samples
wt_md_hr.list<-lapply(cohortlist, function(c){
  md<-matched_data.list[[c]]
  md.dist<-dist_md_hr.list[[c]]
  if(!is.null(md)){
    md.dist_gat<-md.dist%>%
      as.matrix()%>%
      as.data.frame()%>%
      rownames_to_column("SmplID1")%>%
      gather(SmplID2,dist,-SmplID1)%>%
      left_join(md,by=c("SmplID1"="SmplID"))%>%
      select(SmplID1,SmplID2,dist,subclass,Cohort_precise,IBD)%>%
      rename("subclass1"=subclass)%>%
      rename("Cohort_precise1"=Cohort_precise)%>%
      rename("IBD1"=IBD)%>%
      left_join(md,by=c("SmplID2"="SmplID"))%>%
      select(SmplID1,SmplID2,dist,subclass1,Cohort_precise1,IBD1,subclass,Cohort_precise,IBD)%>%
      rename("subclass2"=subclass)%>%
      rename("Cohort_precise2"=Cohort_precise)%>%
      rename("IBD2"=IBD)%>%
        filter(SmplID1!=SmplID2)%>%
      mutate(IBD=ifelse(IBD1==IBD2,paste(IBD1),"control-IBD"))%>%
      mutate(IBD=ifelse(IBD=="healthy","control",paste(IBD)))
   
    md.dist_gat$IBD<-as.factor(md.dist_gat$IBD)
    wt<-coin::wilcox_test(dist ~ IBD, data=md.dist_gat%>%filter(IBD1==IBD2))
    
    
  } else {
    wt<-NULL
  }
  return(wt)
})
names(wt_md_hr.list)<-cohortlist


wt_md_hr.listDF<-lapply(cohortlist, function(c){
  wt<-wt_md_hr.list[[c]]
  if(!is.null(wt)){
    wt.df<-data.frame(
      cohort=c,
      p=coin::pvalue(wt),
      Z=coin::statistic(wt),
      n=nrow(matched_data.list[[c]])
    )
  } else {
    wt.df<-NULL
  }
  return(wt.df)
})
names(wt_md_hr.listDF)<-cohortlist

# collect wilcoxon results in df
wt_md_hr.df<-do.call(rbind, wt_md_hr.listDF)
wt_md_hr.df$padjust<-p.adjust(wt_md_hr.df$p, method = "BH")
wt_md_hr.df<-wt_md_hr.df%>%
  mutate(effectSize=Z/sqrt(n))
wt_md_hr.df

```


### Compare high HR vs low HR within individuals

```{r fig.height=4, fig.width=8}

col_comp2<-c("darkorange1","khaki2","violet","darkorange1","khaki2","violet")
names(col_comp2)<-c("HighHR","LowHR","LowHR-HighHR","flare-like","remission-like","flare-remission")

# select individuals with 4 or more samples
Ind_4plusTP<-(meta%>%
  filter(Cohort_setup=="longitudinal")%>%
  group_by(IndividualID)%>%
  count()%>%
  filter(n>3))$IndividualID

# per individual
meta_hr_ind_percentiles<-meta%>%
  filter(Cohort_setup=="longitudinal")%>%
  filter(IndividualID%in%Ind_4plusTP)%>%
  group_by(IndividualID)%>%
  summarize(HighHR_perc80=quantile(as.numeric(HumanRds), 0.8, na.rm=TRUE),LowHR_perc20=quantile(as.numeric(HumanRds), 0.2, na.rm=TRUE))

meta_hr_ind_num<-meta%>%
  left_join(meta_hr_ind_percentiles,by="IndividualID")%>%
  mutate(HumanRds_group=ifelse(HumanRds>HighHR_perc80,"HighHR",ifelse(HumanRds<LowHR_perc20,"LowHR",NA)))%>% 
  filter(!is.na(HumanRds_group))#%>%
  #distinct(IndividualID, .keep_all = TRUE)

meta_hr_ind_num$IndividualID%>%length()
meta_hr_ind_num$IndividualID%>%unique()%>%length()

###

meta_hr_ind_num%>%
  group_by(Cohort_precise)%>%
  count(HumanRds_group)

spec.tss.HL_hr_ind<-spec.tss[, which(names(spec.tss) %in% meta_hr_ind_num$SmplID)]
spec.tss.HL_hr_ind<-spec.tss.HL_hr_ind[rowSums(spec.tss.HL_hr_ind) != 0,] # to remove empty rows otherwise vegan complains

all(colnames(spec.tss.HL_hr_ind)==meta_hr_ind_num$SmplID)
spec.tss.HL_hr_ind <- spec.tss.HL_hr_ind[,meta_hr_ind_num$SmplID]
all(colnames(spec.tss.HL_hr_ind)==meta_hr_ind_num$SmplID)

spec.tss.HL_hr_ind<-t(spec.tss.HL_hr_ind)
dist.HL_hr_ind<-vegan::vegdist(spec.tss.HL_hr_ind, method = "bray")

all(labels(dist.HL_hr_ind)==meta_hr_ind_num$SmplID)

dist.HL_hr_ind_gat<-dist.HL_hr_ind%>%
  as.matrix()%>%
  as.data.frame()%>%
  rownames_to_column("SmplID1")%>%
  gather(SmplID2,dist,-SmplID1)%>%
  left_join(meta_hr_ind_num,by=c("SmplID1"="SmplID"))%>%
  select(SmplID1,SmplID2,dist,HumanRds_group,Cohort_precise,IBD2,IndividualID)%>%
  rename("HumanRds_group1"=HumanRds_group)%>%
  rename("Cohort_precise1"=Cohort_precise)%>%
  rename("IBD21"=IBD2)%>%
  rename("IndividualID1"=IndividualID)%>%
  left_join(meta_hr_ind_num,by=c("SmplID2"="SmplID"))%>%
  select(SmplID1,SmplID2,dist,HumanRds_group1,Cohort_precise1,IBD21,IndividualID1,HumanRds_group,Cohort_precise,IBD2,IndividualID)%>%
  rename("HumanRds_group2"=HumanRds_group)%>%
  rename("Cohort_precise2"=Cohort_precise)%>%
  rename("IBD22"=IBD2)%>%
  rename("IndividualID2"=IndividualID)%>%
  mutate(comparison=ifelse(HumanRds_group1==HumanRds_group2,HumanRds_group1,"LowHR-HighHR"))

dist.HL_hr_ind_gat_intraInd<-
  dist.HL_hr_ind_gat%>%
  filter(IndividualID1==IndividualID2)%>%
  filter(SmplID1!=SmplID2)%>%
  mutate(cohort=gsub("Fang_UCSD_","",Cohort_precise1))%>%
  mutate(cohort=gsub("cohort1000","1000",cohort))
```

```{r}
# wilcoxon tests blocked by cohort
# make sure variables are factors
dist.HL_hr_ind_gat_intraInd$comparison<-as.factor(dist.HL_hr_ind_gat_intraInd$comparison)
dist.HL_hr_ind_gat_intraInd$cohort<-as.factor(dist.HL_hr_ind_gat_intraInd$cohort)

# low HR vs High HR
wt_HL_hr_ind.Low_vs_High<-coin::wilcox_test(dist ~ comparison | cohort, data=dist.HL_hr_ind_gat_intraInd%>%filter(comparison%in%c("LowHR","HighHR")))
wt_HL_hr_ind.Low_vs_High
## effect size
coin::statistic(wt_HL_hr_ind.Low_vs_High)/sqrt(dist.HL_hr_ind_gat_intraInd%>%filter(comparison%in%c("LowHR","HighHR"))%>%nrow())

# low HR vs LowHR-HighHR
wt_HL_hr_ind.Low_vs_LowHigh<-coin::wilcox_test(dist ~ comparison | cohort, data=dist.HL_hr_ind_gat_intraInd%>%filter(comparison%in%c("LowHR","LowHR-HighHR")))
wt_HL_hr_ind.Low_vs_LowHigh

# High HR vs LowHR-HighHR
wt_HL_hr_ind.LowHigh_vs_High<-coin::wilcox_test(dist ~ comparison | cohort, data=dist.HL_hr_ind_gat_intraInd%>%filter(comparison%in%c("LowHR-HighHR","HighHR")))
wt_HL_hr_ind.LowHigh_vs_High
```

```{r}
# create boxplots 
dist.HL_hr_ind_gat_intraInd<-dist.HL_hr_ind_gat_intraInd%>%
  mutate(comparison2=gsub("HR","Iidx",comparison))
col_comp3<-col_comp2
names(col_comp3)<-gsub("HR","Iidx",names(col_comp2))
p_iiv_cohorts<-ggplot(dist.HL_hr_ind_gat_intraInd, aes(x=comparison2,y=dist))+
  geom_boxplot(aes(fill=comparison2),outlier.size = 0.1)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45,hjust=1))+
  scale_fill_manual(values=col_comp3, breaks=names(col_comp3),labels=c("flare-like","remission-like","flare-remission","flare-like","remission-like","flare-remission"))+
  ylab("Intra-individual BC dissimilarity")+
  facet_grid(. ~ cohort)
```

```{r}
# intra-individual dissimilarities, paired wilcoxon test with highHR and lowHR samples within individuals
dist.HL_hr_ind_gat_intraInd_highHR<-dist.HL_hr_ind_gat_intraInd%>%
  filter(comparison=="HighHR")%>%
  select(c(dist,comparison,IndividualID1))
dist.HL_hr_ind_gat_intraInd_lowHR<-dist.HL_hr_ind_gat_intraInd%>%
  filter(comparison=="LowHR")%>%
  select(c(dist,comparison,IndividualID1))
dist.HL_hr_ind_gat_intraInd_lowhighHR<-dist.HL_hr_ind_gat_intraInd%>%
  filter(comparison=="LowHR-HighHR")%>%
  select(c(dist,comparison,IndividualID1))

dist.HL_hr_ind_gat_intraInd_highlowpairs<-dist.HL_hr_ind_gat_intraInd_highHR%>%
  inner_join(dist.HL_hr_ind_gat_intraInd_lowHR, by="IndividualID1")

# calculate wilcoxon tests
wilcox.test(dist.HL_hr_ind_gat_intraInd_highlowpairs$dist.x,dist.HL_hr_ind_gat_intraInd_highlowpairs$dist.y, paired=TRUE)
wt_iiv_paired<-coin::wilcoxsign_test(dist.HL_hr_ind_gat_intraInd_highlowpairs$dist.x ~ dist.HL_hr_ind_gat_intraInd_highlowpairs$dist.y)
coin::statistic(wt_iiv_paired)/sqrt(length(dist.HL_hr_ind_gat_intraInd_highlowpairs$dist.x))
```

```{r}
# create plots
p_iiv_all<-ggplot(dist.HL_hr_ind_gat_intraInd, aes(x=comparison,y=dist))+
  geom_boxplot(aes(fill=comparison),outlier.size = 0.1)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45,hjust=1))+
  scale_fill_manual(values=col_comp2, breaks=names(col_comp2),labels=c("flare-like","remission-like","flare-remission","flare-like","remission-like","flare-remission"))+
  ylab("Intra-individual BC dissimilarity")

dist.HL_hr_ind_gat_intraInd_highlowpairs_gat<-dist.HL_hr_ind_gat_intraInd_highlowpairs%>%
  gather(pairs,dist,c(dist.x,dist.y))%>%
  distinct(IndividualID1,pairs,.keep_all = TRUE)%>%
  mutate(pairs=ifelse(pairs=="dist.x","flare-like","remission-like"))
p_iiv_paired<-ggplot(dist.HL_hr_ind_gat_intraInd_highlowpairs_gat, aes(x = pairs, y = dist)) +
  geom_boxplot(outlier.shape = NA, aes(fill = pairs)) +
  geom_jitter(width = 0.1, alpha = 0.6) +
  geom_line(aes(group = IndividualID1), alpha = 0.4) +
  scale_fill_manual(values=col_comp2)+
  theme_bw()+
  theme(axis.text.x=element_blank(),axis.title.x=element_blank(),axis.ticks.x = element_blank())+
  ylab("Intra-individual BC dissimilarity")

p_iiv<-plot_grid(p_iiv_all+theme(legend.position="none", axis.text.x=element_blank(), axis.title.x=element_blank(), axis.ticks.x=element_blank()),
          p_iiv_cohorts+theme(legend.position="top",axis.text.x=element_blank(), axis.text.y=element_blank(),axis.title=element_blank(), axis.ticks=element_blank()),
          nrow=1, align="h", axis="tb", rel_widths = c(1,4),labels=c("A","B"))
p_iiv

```

```{r}
# export plots
pdf(paste0(PATHanalNov2023,"Downstream/Results/Figures/p_iiv_hlHR.pdf"), width=8,height=3)
p_iiv
dev.off()

pdf(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_p_iiv_hlHR.pdf"), width=7,height=3)
p_iiv_cohorts+theme(legend.position="top",axis.text.x=element_blank(), axis.title.x=element_blank(), axis.ticks=element_blank())
dev.off()
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_p_iiv_hlHR.png"), p_iiv_cohorts+theme(legend.position="top",axis.text.x=element_blank(), axis.title.x=element_blank(), axis.ticks=element_blank())
,width=8,height=4)

ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_p_iiv-paired_hlHR.png"), p_iiv_paired, width=4,height=4)
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_p_iiv-paired_hlHR.pdf"), p_iiv_paired, width=4,height=4)

saveRDS(dist.HL_hr_ind_gat_intraInd,paste0(PATHanalNov2023,"humanReads/dist.HL_hr_ind_gat_intraInd.rds"))
```

```{r}
# create supplementary fig and export
p_iiv_A<-p_iiv_cohorts+theme(legend.position="none",legend.title = element_blank(),axis.text.x=element_blank(), axis.title.x=element_blank(), axis.ticks=element_blank())
p_iiv_B<-p_iiv_paired+theme(legend.position = "none")
p_iiv_leg<-get_legend(p_iiv_cohorts+theme(legend.title = element_blank()))
p_suppl_iiv_pairChrts<-plot_grid(p_iiv_A,
          p_iiv_B,
          p_iiv_leg,
          rel_widths = c(4,1,1),
          align = "h", axis = "tb",
          labels=c("A","B"), nrow=1)

ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_p_iiv-paired-cohorts_hlHR.png"), p_suppl_iiv_pairChrts, width=9,height=4)
ggsave(paste0(PATHanalNov2023,"Downstream/Results/Figures/suppl_p_iiv-paired-cohorts_hlHR.pdf"), p_suppl_iiv_pairChrts, width=9,height=4)

p_suppl_iiv_pairChrts
```


# Save workspace
```{r, message =F, warning =F}
save.image(paste0(PATHanalNov2023,"DA_analysis/November_2023_v3/R_analysis/RData/Temporal_variation.RData"))

```